---
title: "Aline's work"
format:
  html:
    embed-resources: true
editor: visual
toc: true
toc_depth: 4
theme:
  light: flatly
  dark: darkly
toc-location: left
execute:
  cache: true
  echo: false
  message: false
  warning: false
---

```{r}
#| cache: false
#| label: chargement-donnees
#| message: false
#| warning: false
#| include: false
source("~/Documents/POP_ALS_2025_02_03/1_codes/1_data_loading.R")

results_descriptive <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.1_results_descriptive.rds")
results_POPs_ALS_occurrence <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.2_results_POPs_ALS_occurrence.rds")
results_POPs_ALS_survival <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.3_results_POPs_ALS_survival.rds")
results_fattyacids_ALS_occurrence <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.4_results_fattyacids_ALS_occurrence.rds")
results_fattyacids_ALS_survival <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.5_results_fattyacids_ALS_survival.rds")
results_proteomic_ALS_occurrence <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.6.1_results_proteomic_ALS_occurrence.rds")
results_proteomic_ALS_occurrence_SL <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.6.2_results_proteomic_ALS_occurrence_SL.rds")
results_proteomic_ALS_occurrence_tidymodels <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.6.3_results_proteomic_ALS_occurrence_tidymodels.rds")
results_proteomic_ALS_occurrence_y_to_als <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.6.4_results_proteomic_ALS_occurrence_y_to_als.rds") 
results_proteomic_ALS_survival <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.7_results_proteomic_ALS_survival.rds")
results_POPs_proteomic <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/2.8_results_POPs_proteomic.rds")
```

# Danish data description
## Metadata

Not imputed for missing data

```{r}
#| label: danish-description-metadata
results_descriptive$danish$covar_danish |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Multiple cholesterol variables

```{r, fig.height=2}
#| label: danish-description-cholesterol
boxplot(data = bdd_danish, vars = c("fS_Kol", "cholesterol"))
descrip_num(data = bdd_danish, vars = c("fS_Kol", "cholesterol")) |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
cor.test(bdd_danish$fS_Kol, bdd_danish$cholesterol)
```

### ALS survival

```{r}
#| label: danish-description-als-survival
bdd_danish |> 
  filter(als == 1) |>
  ggplot() +
  aes(x = follow_up_death) +
  geom_density(fill = "lightgray") +
  theme_bw() +
  stat_overlay_normal_density(color = "red", linetype = "dashed") +
  theme(axis.title = element_blank())
descrip_num(data = bdd_danish, vars = "follow_up_death") |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Timing between baseline, ALS diagnosis and death

```{r}
#| label: danish-description-time
timing <- bdd |> 
  filter(study == "Danish") |>
  filter(als == 1) |>
  select(baseline_age, diagnosis_age, death_age, 
         time_baseline_diagnosis, time_baseline_death, time_diagnosis_death)

timing |> 
  tbl_summary() |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

descrip_num(data = timing, vars = colnames(timing))   |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
rm(timing)
```

### Other neurological diagnoses

```{r}
#| label: danish-description-other-neuro-diseases
results_descriptive$danish$table_other_diag_1
results_descriptive$danish$figure_other_diag
results_descriptive$danish$table_other_diag_2 |> autofit()
results_descriptive$danish$table_other_diag_3 |> autofit()
```

## POPs

### Distribution tables

```{r}
#| label: danish-description-POPs
results_descriptive$danish$POPs_table_danish |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Comparison of distribution cases/controls

```{r}
#| label: danish-description-comparison_case-controls
results_descriptive$danish$POPs_table_danish_by_als|>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

```{r fig.height=11, fig.width=11}
#| label: danish-description-POP_heatmap
corrplot(results_descriptive$danish$POPs_heatmap_danish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))
```

### Boxplots

```{r, fig.width=10, fig.height=10}
#| label: danish-description-POP-boxplot-1
results_descriptive$danish$POPs_boxplot_danish_by_als
```

```{r, fig.height=3}
#| label: danish-description-POP-boxplot-2
results_descriptive$danish$POPs_group_boxplot_danish_by_als
```

POPs distribution depending on alive/deceased status at the end of the follow-up (among cases)

```{r, fig.height=3}
#| label: danish-description-POP-boxplot-3
results_descriptive$danish$POPs_group_boxplot_danish_by_death
```

### Detailed distribution

#### Summed

```{r}
#| label: danish-description-POP-summed
densityplot(data = bdd_danish, vars = POPs_group)
boxplot(data = bdd_danish, vars = POPs_group)
heatmap_cor(cormat = bdd_danish[POPs_group], decimal = 1)
```

#### PCBs

```{r}
#| label: danish-description-PCBs-1
PCBs <- bdd_danish |> select(all_of(POPs)) |> select(contains("PCB")) |> colnames()
densityplot(data = bdd_danish, vars = PCBs[1:9])
```

```{r, fig.height=3.33}
#| label: danish-description-PCBs-2
densityplot(data = bdd_danish, vars = PCBs[10:13])
```

```{r}
#| label: danish-description-PCBs-3
boxplot(data = bdd_danish, vars = PCBs[1:10])
```

```{r, fig.height=2}
#| label: danish-description-PCBs-4
boxplot(data = bdd_danish, vars = PCBs[11:13])
```

```{r}
#| label: danish-description-PCBs-heatmap
heatmap_cor(cormat = bdd_danish[PCBs], decimal = 1)
rm(PCBs)
```

#### OCPs

```{r}
#| label: danish-description-OCPs
OCPs <- bdd_danish |> select(all_of(POPs)) |> select(contains("OCP")) |> colnames()
densityplot(data = bdd_danish, vars = OCPs)
boxplot(data = bdd_danish, vars = OCPs)
heatmap_cor(cormat = bdd_danish[OCPs], decimal = 1)
```

#### PBDEs

```{r,  fig.width=7, fig.height=2}
#| label: danish-description-PBDEs-1
PBDEs <- bdd_danish |> select(all_of(POPs)) |> select(contains("PBDE")) |> colnames()
densityplot(data = bdd_danish, vars = PBDEs)
boxplot(data = bdd_danish, vars = PBDEs)
```

```{r,  fig.width=7, fig.height=3}
#| label: danish-description-PBDEs-2
heatmap_cor(cormat = bdd_danish[PBDEs], decimal = 1)
```

## Fatty acids

### Distribution tables

```{r}
#| label: danish-description-fatty-acids-1
results_descriptive$danish$fattyacids_table_danish|>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
```

### Comparison of distribution cases/controls

```{r}
#| label: danish-description-fatty-acids-2
results_descriptive$danish$fattyacids_table_danish_by_als |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

```{r, fig.height=12, fig.width=12}
#| label: danish-description-fatty-acids-3
corrplot(results_descriptive$danish$fattyacids_heatmap_danish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))
```

### Boxplots

```{r, fig.height=10, fig.width=10}
#| label: danish-description-fatty-acids-4
results_descriptive$danish$fattyacids_boxplot_danish_by_als
```

### Detailed distribution

#### Density plots

```{r, fig.height=10, fig.width=10}
#| label: danish-description-fatty-acids-5
bdd_danish |> 
  select(sample, all_of(fattyacids_tot)) |>
  pivot_longer(cols = -sample, names_to = "fattyacids", values_to = "values") |>
  mutate(fattyacids = fct_recode(fattyacids, !!!fattyacids_tot_labels)) |>
  ggplot() +
  aes(x = values) +
  geom_density(fill = "#112446") +
  theme_minimal() +
  facet_wrap(~fattyacids, scales = "free", ncol = 4)
```

## Heatmap POPs/Fatty acids

```{r, fig.width=12, fig.height=12}
#| label: danish-description-fatty-acids-6
results_descriptive$danish$POPs_fattyacids_heatmap_danish
```

## Proteomic

### Distribution table

```{r}
#| label: danish-description-proteomic-1
results_descriptive$danish$proteomic_table_danish  |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Comparison of distribution cases/controls

```{r}
#| label: danish-description-proteomic-2
#| eval: false
#| include: false
results_descriptive$danish$proteomic_table_danish_by_als|>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

All

```{r fig.height=11, fig.width=11}
#| label: danish-description-proteomic-3
corrplot(results_descriptive$danish$proteomic_heatmap_danish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black',
         order = "hclust",         # regroupement des variables selon niveau de cor
         tl.srt = 45, 
         # addCoef.col = "black",
         # number.cex = 0.8,
         # number.digits = 1,
         tl.cex = 0.4,             # taille police des variables
         col = rev(COL2(diverging = "RdYlBu")))
```

```{r fig.height=11, fig.width=11}
#| label: danish-description-proteomic-4
vars <- colnames(results_descriptive$danish$proteomic_heatmap_danish)

col_labels <- ifelse(grepl("proteomic_metabolism_", vars), "#1b9e77",  # vert
                     ifelse(grepl("proteomic_immun_res_", vars), "#d95f02",    # orange
                            ifelse(grepl("proteomic_neuro_explo_", vars), "#7570b3",  # violet
                                   "black")))

corr_order <- corrMatOrder(results_descriptive$danish$proteomic_heatmap_danish, order = "hclust", hclust.method = "complete")
col_labels_ordered <- col_labels[corr_order]

corrplot(results_descriptive$danish$proteomic_heatmap_danish,
         method = "color",
         type = "lower",
         order = "hclust",
         hclust.method = "complete",
         # order = "original",
         tl.col = col_labels_ordered,  # couleurs alignées à l’ordre
         tl.srt = 45,
         tl.cex = 0.4,
         col = rev(COL2(diverging = "RdYlBu")))

rm(vars, col_labels, corr_order, col_labels_ordered)
```

Neuro explo

```{r fig.height=11, fig.width=11}
#| label: danish-description-proteomic-5
corrplot(results_descriptive$danish$proteomic_heatmap_danish_neuro_explo, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black',
         order = "hclust",         # regroupement des variables selon niveau de cor
         tl.srt = 45, 
         # addCoef.col = "black",
         # number.cex = 0.8,
         # number.digits = 1,
         tl.cex = 0.4,             # taille police des variables
         col = rev(COL2(diverging = "RdYlBu")))
```

Immune response

```{r fig.height=11, fig.width=11}
#| label: danish-description-proteomic-6
corrplot(results_descriptive$danish$proteomic_heatmap_danish_immun_res, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black',
         order = "hclust",         # regroupement des variables selon niveau de cor
         tl.srt = 45, 
         # addCoef.col = "black",
         # number.cex = 0.8,
         # number.digits = 1,
         tl.cex = 0.4,             # taille police des variables
         col = rev(COL2(diverging = "RdYlBu")))
```

Metabolism

```{r fig.height=11, fig.width=11}
#| label: danish-description-proteomic-7
corrplot(results_descriptive$danish$proteomic_heatmap_danish_metabolism, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black',
         order = "hclust",         # regroupement des variables selon niveau de cor
         tl.srt = 45, 
         # addCoef.col = "black",
         # number.cex = 0.8,
         # number.digits = 1,
         tl.cex = 0.4,             # taille police des variables
         col = rev(COL2(diverging = "RdYlBu")))

```

### Boxplots

```{r, fig.width=10, fig.height=40}
#| label: danish-description-proteomic-8
results_descriptive$danish$proteomic_boxplot_danish
```

Comparison between cases and controls

```{r, fig.width=12, fig.height=60}
#| label: danish-description-proteomic-9
results_descriptive$danish$proteomic_boxplot_danish_by_als
```

Comparison between alive/deceased status at the end of the follow-up (among cases)

```{r, fig.width=12, fig.height=60}
#| label: danish-description-proteomic-10
results_descriptive$danish$proteomic_boxplot_danish_by_death
```

### NfL

```{r, fig.height=7, fig.width=9}
#| label: danish-description-proteomic-nfl
results_descriptive$danish$figure_NEFL_1
```

## EVs

### Distribution tables

```{r}
#| label: danish-description-EVs-1
results_descriptive$danish$EVs_table_danish |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")|>
  autofit()
```

### Comparison of distribution cases/controls

```{r}
#| label: danish-description-EVs-2
results_descriptive$danish$EVs_table_danish_by_als |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

```{r, fig.height=12, fig.width=12}
#| label: danish-description-EVs-3
corrplot(results_descriptive$danish$EVs_heatmap_danish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))

```

### Boxplots

```{r, fig.height=7, fig.width=10}
#| label: danish-description-EVs-4
results_descriptive$danish$EVs_boxplot_danish
```

Comparison between cases and controls

```{r, fig.height=10, fig.width=10}
#| label: danish-description-EVs-5
results_descriptive$danish$EVs_boxplot_danish_by_als
```

Comparison between alive/deceased status at the end of the follow-up (among cases)

```{r, fig.width=10, fig.height=10}
#| label: danish-description-EVs-6
results_descriptive$danish$EVs_boxplot_danish_by_death
```

# Finnish data description

## ALS survival

```{r, fig.height=2, fig.width=3}
#| label: finnish-description-als-survival
bdd_finnish |> 
  filter(als == 1) |>
  select(follow_up_death, status_death, study) |>
  tbl_summary(by = "study") |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

densityplot(data = bdd_finnish, vars = "follow_up_death") 
descrip_num(data = bdd_finnish, vars = "follow_up_death") |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

## Metadata

```{r}
#| label: finnish-description-metadata
results_descriptive$finnish$covar_finnish |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

## POPs

### Distribution tables

```{r}
#| label: finnish-description-POPs-1
results_descriptive$finnish$POPs_table_finnish |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Comparison of distribution cases/controls

```{r}
#| label: finnish-description-POPs-2
results_descriptive$finnish$POPs_table_finnish_by_als |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

```{r fig.height=11, fig.width=11}
#| label: finnish-description-POPs-3
corrplot(results_descriptive$finnish$POPs_heatmap_finnish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))
```

### Boxplots

```{r, fig.width=10, fig.height=8}
#| label: finnish-description-POPs-4
# results_descriptive$finnish$POPs_boxplot_finnish
results_descriptive$finnish$POPs_boxplot_finnish_by_als
```

### Detailed distribution

#### Summed

```{r}
#| label: finnish-description-POPs-5
densityplot(data = bdd_finnish, vars = POPs_group)
boxplot(data = bdd_finnish, vars = POPs_group)
```

```{r, fig.height=4, fig.width=4}
#| label: finnish-description-POPs-6
heatmap_cor(cormat = bdd_finnish[POPs_group], decimal = 1)
```

#### PCBs

```{r}
#| label: finnish-description-POPs-7
PCBs <- bdd_finnish |> select(all_of(POPs_finnish)) |> select(contains("PCB")) |> colnames()
densityplot(data = bdd_finnish, vars = PCBs[1:9])
```

```{r, fig.height=3.33}
#| label: finnish-description-POPs-8
densityplot(data = bdd_finnish, vars = PCBs[10:13])
```

```{r}
#| label: finnish-description-POPs-9
boxplot(data = bdd_finnish, vars = PCBs[1:10])
```

```{r, fig.height=2}
#| label: finnish-description-POPs-10
boxplot(data = bdd_finnish, vars = PCBs[11:13])
```

```{r}
#| label: finnish-description-POPs-11
heatmap_cor(cormat = bdd_finnish[PCBs], decimal = 1)
rm(PCBs)
```

#### OCPs

```{r}
#| label: finnish-description-POPs-12
OCPs <- bdd_finnish |> select(all_of(POPs_finnish)) |> select(contains("OCP")) |> colnames()
densityplot(data = bdd_finnish, vars = OCPs)
boxplot(data = bdd_finnish, vars = OCPs)
# heatmap_cor(cormat = bdd_finnish[OCPs], decimal = 1)
```

#### PBDEs

```{r,  fig.width=7, fig.height=2}
#| label: finnish-description-POPs-13
PBDEs <- bdd_finnish |> select(all_of(POPs_finnish)) |> select(contains("PBDE")) |> colnames()
densityplot(data = bdd_finnish, vars = PBDEs)
boxplot(data = bdd_finnish, vars = PBDEs)
```

```{r,  fig.width=7, fig.height=4}
#| label: finnish-description-POPs-14
heatmap_cor(cormat = bdd_finnish[PBDEs], decimal = 1)
```

## Fatty acids

### Distribution tables

```{r}
#| label: finnish-description-fatty-acids-1
results_descriptive$finnish$fattyacids_table_finnish |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
```

### Comparison of distribution cases/controls

```{r}
#| label: finnish-description-fatty-acids-2
results_descriptive$finnish$fattyacids_table_finnish_by_als |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

```{r, fig.height=14, fig.width=14}
#| label: finnish-description-fatty-acids-3
corrplot(results_descriptive$finnish$fattyacids_heatmap_finnish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))
```

### Boxplots

```{r, fig.height=8, fig.width=10}
#| label: finnish-description-fatty-acids-4
results_descriptive$finnish$fattyacids_boxplot_finnish_by_als
```

### Detailed distribution

```{r, fig.height=10, fig.width=10}
#| label: finnish-description-fatty-acids-5
bdd_finnish |> 
  select(sample, all_of(fattyacids_tot)) |>
  pivot_longer(cols = -sample, names_to = "fattyacids", values_to = "values") |>
    mutate(fattyacids = fct_recode(fattyacids, !!!fattyacids_tot_labels)) |>
  ggplot() +
  aes(x = "", y = values) +
  geom_boxplot(shape = "circle", fill = "#112446") +
  coord_flip() +
  theme_bw() +
  theme(axis.title = element_blank()) +
  facet_wrap(~fattyacids, scales = "free", ncol = 3L, nrow = 14L)
```

## Heatmap POPs/Fatty acids

```{r, fig.width=12, fig.height=12}
#| label: finnish-description-POPs-fatty-acids
results_descriptive$finnish$POPs_fattyacids_heatmap_finnish
```

## Multiple cholesterol variables issue

Why do we have 3 cholesterol variables in the finnish data ?

```{r, fig.height=2}
#| label: finnish-description-cholesterol
boxplot(data = bdd_finnish, vars = c("fS_Kol", "chol", "cholesterol"))

descrip_num(data = bdd_finnish, vars = c("fS_Kol", "chol", "cholesterol")) |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") 
heatmap_cholesterol <- bdd_finnish |> select("fS_Kol", "chol", "cholesterol")
heatmap_cholesterol <- cor(heatmap_cholesterol, 
                    use = "pairwise.complete.obs", 
                    method = "pearson")

corrplot(heatmap_cholesterol, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))

```

## Timing between baseline, ALS diagnosis and death

### FMC

```{r}
#| label: finnish-description-timing-1
timing <- bdd |> 
  filter(study == "FMC") |>
  filter(als == 1) |>
  select(baseline_age, diagnosis_age, death_age) |>
  mutate(time_baseline_diagnosis = diagnosis_age - baseline_age, 
         time_baseline_death = death_age - baseline_age, 
         time_diagnosis_death = death_age - diagnosis_age) 

timing |> 
  tbl_summary() |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

descrip_num(data = timing, vars = colnames(timing)) |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
rm(timing)
```

### FMCF

```{r}
#| label: finnish-description-timing-2
timing <- bdd |> 
  filter(study == "FMCF") |>
  filter(als == 1) |>
  select(baseline_age, diagnosis_age, death_age, 
         time_baseline_diagnosis, time_baseline_death, time_diagnosis_death)

timing |> 
  tbl_summary() |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
   set_table_properties(align = "left")|>
  set_table_properties(align = "left")

descrip_num(data = timing, vars = colnames(timing)) |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
rm(timing)
```

### MFH

```{r}
#| label: finnish-description-timing-3
timing <- bdd |> 
  filter(study == "MFH") |>
  filter(als == 1) |>
  select(baseline_age, diagnosis_age, death_age, 
         time_baseline_diagnosis, time_baseline_death, time_diagnosis_death)

timing |> 
  tbl_summary() |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

descrip_num(data = timing, vars = colnames(timing)) |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
rm(timing)
```

# Comparison danish/finnish data

## On the total population

### Metadata

```{r}
#| label: comp-danish-finnish-description-metadata
results_descriptive$comp$covar_comp |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### POPs

```{r, fig.width=10, fig.height=10}
#| label: comp-danish-finnish-description-POPs
results_descriptive$comp$POPs_table_comp |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
results_descriptive$comp$POPs_boxplot_comp
```

### Fatty acids

```{r, fig.height=12, fig.width=12}
#| label: comp-danish-finnish-description-fatty-acids
results_descriptive$comp$fattyacids_table_comp |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

results_descriptive$comp$fattyacids_boxplot_comp 
```

## Among cases

(for ALS survival analyses)

### Metadata

```{r}
#| label: comp-danish-finnish-description-cases-metadata
results_descriptive$comp$covar_comp_cases |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### POPs

```{r, fig.width=9}
#| label: comp-danish-finnish-description-POPs-1
results_descriptive$comp$POPs_table_comp_cases |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

results_descriptive$comp$POPs_boxplot_comp_cases
```

```{r fig.height=12, fig.width=12}
#| label: comp-danish-finnish-description-POPs-2
corrplot(results_descriptive$comp$POPs_heatmap_cases, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))
```

# POPs and ALS occurence

## Danish cohort

### Effects of subject characteristics on ALS occurence

```{r}
#| label: danish-POPs-ALS-occurence-1
results_POPs_ALS_occurrence$main$covar |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Main analysis

#### Quartile models

In the copollutant models, the pollutant of interest is quartile while the others are s() in a GAM model

```{r fig.height=9, fig.width=9}
#| label: danish-POPs-ALS-occurence-2
results_POPs_ALS_occurrence$main$plot_quart
```

```{r}
#| label: danish-POPs-ALS-occurence-3
#| column: screen-inset-right
extra_rows <- results_POPs_ALS_occurrence$main$results_quart |>
  distinct(variable) |> 
  mutate(
    quartiles = "Quartile 1",
    "OR_base" = '-', "95%CI_base" = '-', "p.value_base" = '', "p.value_heterogeneity_base" = '', "p.value_trend_base" = '',
    "OR_adjusted" = '-', "95%CI_adjusted" = '-', "p.value_adjusted" = '', , "p.value_heterogeneity_adjusted" = '', "p.value_trend_adjusted" = '',
    "OR_copollutant" = '-', "95%CI_copollutant" ='-', "p.value_copollutant" = '', , "p.value_heterogeneity_copollutant" = '', "p.value_trend_copollutant" = '')

table_S2 <- 
  results_POPs_ALS_occurrence$main$main_results |>
  select(-lower_CI, -upper_CI, - p.value_raw) |>
  filter(model %in% c('base_quart', 'adjusted_quart', 'copollutant_quart')) |>
  pivot_wider(
    names_from = model,  
    values_from = c(OR, `95%CI`, p.value, p.value_heterogeneity, p.value_trend)) |>
  select(variable, 
         quartiles = df,
         contains("base_quart"), 
         contains("adjusted_quart"), 
         contains("copollutant_quart")) 
colnames(table_S2) <- gsub('_quart', '', colnames(table_S2))

table_S2 <- table_S2 |>
  mutate_if(is.numeric, as.character) |>
  bind_rows(extra_rows) |>
  mutate(
    variable = gsub("OCP_", "", variable), 
    variable = fct_relevel(variable, "PCB_4", "PCB_DL", "PCB_NDL",  "HCB", "ΣDDT", "β_HCH", "Σchlordane", "ΣPBDE" ),
    variable = fct_recode(variable, 
                          "Most prevalent PCBs" = "PCB_4",
                          "Dioxin-like PCBs" = "PCB_DL",
                          "Non-dioxin-like PCBs" = "PCB_NDL",
                          "β-HCH" = "β_HCH")) |>
  arrange(variable, quartiles) |>
  group_by(variable) |>
  mutate(p.value_heterogeneity_base = ifelse(quartiles == 'Quartile 1', p.value_heterogeneity_base[quartiles == 'Quartile 2'], ''), 
         p.value_trend_base = ifelse(quartiles == 'Quartile 1', p.value_trend_base[quartiles == 'Quartile 2'], ''),
         p.value_heterogeneity_adjusted = ifelse(quartiles == 'Quartile 1', p.value_heterogeneity_adjusted[quartiles == 'Quartile 2'], ''), 
         p.value_trend_adjusted = ifelse(quartiles == 'Quartile 1', p.value_trend_adjusted[quartiles == 'Quartile 2'], ''), 
         p.value_heterogeneity_copollutant = ifelse(quartiles == 'Quartile 1', p.value_heterogeneity_copollutant[quartiles == 'Quartile 2'], ''), 
         p.value_trend_copollutant = ifelse(quartiles == 'Quartile 1', p.value_trend_copollutant[quartiles == 'Quartile 2'], '')) |>
  ungroup() |>
  rename('OR' = 'OR_base', '95% CI' = '95%CI_base', 'p-value' = 'p.value_base', 
         'Heterogeneity test' = 'p.value_heterogeneity_base', 'Trend test' = 'p.value_trend_base',
         'OR ' = 'OR_adjusted', '95% CI ' = '95%CI_adjusted', 'p-value ' = 'p.value_adjusted', 
         'Heterogeneity test ' = 'p.value_heterogeneity_adjusted', 'Trend test ' = 'p.value_trend_adjusted',
         ' OR ' = 'OR_copollutant', ' 95% CI ' = '95%CI_copollutant', ' p-value ' = 'p.value_copollutant',
         ' Heterogeneity test ' = 'p.value_heterogeneity_copollutant', ' Trend test ' = 'p.value_trend_copollutant')
table_S2 <- table_S2 |> flextable() |>
  add_footer_lines(
    "1POPs were summed as follows: Dioxin-like PCBs corresponds to PCBs 118 and 156; non-dioxin-like PCBs corresponds to PCBs 28, 52, 74, 99, 101, 138, 153, 170, 180, 183, 187; most prevalent PCBs corresponds to PCBs 118, 138, 153, 180; ΣPBDE corresponds to PBDEs 47, 99, 153; ΣDDT corresponds to p,p’-DDT and p,p’-DDE and finally Σchlordane corresponds to trans-nonanchlor and oxychlordane.
  2All models are matched for sex and age. Adjusted models further account for smoking, BMI, serum total cholesterol, marital status, and education. Co-pollutant models further include all chemicals, where a GAM is used in which the chemical of interest is categorized into quartiles, while the other pollutants are included as smooth terms.
  3Estimated risk of ALS when exposures to POP are at quartiles 2, 3, and 4, compared to quartile 1.
  4CI: Confidence interval.
  5Heterogeneity tests in outcome value across POP quartiles, matched on age and sex.
  6Trend tests using continuous variables whose values corresponded to the quartile specific median POP levels, matched on age and sex.
  7Heterogeneity tests in outcome value across POP quartiles, matched on age and sex, and adjusted for smoking, BMI, serum total cholesterol, marital status, and education.
  8Trend tests using continuous variables whose values corresponded to the quartile specific median POP levels, matched on age and sex, and adjusted for smoking, BMI, serum total cholesterol, marital status, and education.") |>
  add_header(
    "variable" = "Exposures", "quartiles" = "Quartiles",
    "OR" = "Base Model", "95% CI" = "Base Model", "p-value" = "Base Model", "Heterogeneity test" = 'Base Model', "Trend test" = 'Base Model',
    "OR " = "Adjusted Model", "95% CI " = "Adjusted Model", "p-value " = "Adjusted Model", "Heterogeneity test " = 'Adjusted Model', "Trend test " = 'Adjusted Model',
    " OR " = "Co-pollutant model", " 95% CI " = "Co-pollutant model", " p-value " = "Co-pollutant model", " Heterogeneity test " = 'Co-pollutant model', " Trend test " = 'Co-pollutant model') |>
  merge_h(part = "header") |>
  merge_v(j = "variable") |>
  theme_vanilla() |>
  bold(j = "variable", part = "body") |>
  align(align = "center", part = "all") |>
  align(j = "variable", align = "left", part = "all") |> 
  merge_at(j = "variable", part = "header") |>
  merge_at(j = "quartiles", part = "header") |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all")|>
  set_table_properties(align = "left")

table_S2
rm(extra_rows, table_S2)
```

#### q-gcomp analysis

```{r, fig.height=3}
#| label: danish-POPs-ALS-occurence-4
results_POPs_ALS_occurrence$main$POPs_ALS_qgcomp_table_danish |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
results_POPs_ALS_occurrence$main$POPs_ALS_qgcomp_figure_danish
```

#### GAM models

Relationship between POPs exposure at baseline time and ALS occurrence estimated using a generalized additive model (gam) from the Danish Diet, Cancer and Health study cohort established between December 1993 and May 1997. The solid blue line represents the predicted ALS occurrence by POPs exposure given that all other covariates are at their respective means. The light blue represents the 95% confidence intervals. All models were adjusted for sex and age. Adjusted models further account for serum cholesterol, BMI, smoking status, marital status and education. The effective degrees of freedom (EDF) for the POPs smooth terms is a summary statistic of a gam that reflects the degree of nonlinearity (edf=1: linear, edf=2: quadratic, edf\>2: more complex than quadratic).

```{r, fig.height=3.5, fig.width=9}
#| label: danish-POPs-ALS-occurence-5
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$PCB_DL, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$PCB_DL, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$PCB_DL, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$PCB_NDL, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$PCB_NDL, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$PCB_NDL, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$PCB_4, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$PCB_4,  
  wrap_elements(grid::nullGrob()),
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$OCP_HCB, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$OCP_HCB, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$OCP_HCB, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$ΣDDT, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$ΣDDT, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$ΣDDT, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$OCP_β_HCH, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$OCP_β_HCH, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$OCP_β_HCH, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$Σchlordane, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$Σchlordane, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$Σchlordane, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$ΣPBDE, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$ΣPBDE, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$ΣPBDE, 
  nrow = 1)
```

### Sensitivity analysis total lipid instead of total cholesterol

```{r fig.height=8, fig.width=9}
#| label: danish-POPs-ALS-occurence-6
results_POPs_ALS_occurrence$main$plot_quart_sensi_lipid
```

```{r fig.height=9, fig.width=9}
#| column: screen-inset-right
#| label: danish-POPs-ALS-occurence-7
results_POPs_ALS_occurrence$main$table_quart_sensi_lipid

```

### Sensitivity analysis about diabetes

```{r fig.height=8, fig.width=9}
#| label: danish-POPs-ALS-occurence-8
results_POPs_ALS_occurrence$main$plot_quart_sensi_diabetes
```

```{r fig.height=9, fig.width=9}
#| column: screen-inset-right
#| label: danish-POPs-ALS-occurence-9
results_POPs_ALS_occurrence$main$table_quart_sensi_diabetes

```

### Sensitivity analysis POPs not summed

#### quartiles

```{r,  fig.height=15, fig.width=10}
#| label: danish-POPs-ALS-occurence-10
results_POPs_ALS_occurrence$sensitivity_not_summed$plot_quart_sensi_not_summed
```

#### GAMs

```{r, fig.height=6, fig.width=15}
#| label: danish-POPs-ALS-occurence-11
#| column: screen-inset-right
wrap_plots(results_POPs_ALS_occurrence$sensitivity_not_summed$plot_adjusted_gam_not_summed[1:8], ncol = 4)
wrap_plots(results_POPs_ALS_occurrence$sensitivity_not_summed$plot_adjusted_gam_not_summed[9:16], ncol = 4)
wrap_plots(results_POPs_ALS_occurrence$sensitivity_not_summed$plot_adjusted_gam_not_summed[17:22], ncol = 4)
```

## Metanalysis (quartiles)

```{r fig.height=7, fig.width=9}
#| label: danish-POPs-ALS-occurence-12
results_POPs_ALS_occurrence$metanalysis$plot_metanalysis_quart
```

```{r}
#| label: danish-POPs-ALS-occurence-13
#| column: screen-inset-right
extra_rows <- results_POPs_ALS_occurrence$metanalysis$metanalysis_quart |>
  distinct(explanatory) |> 
  mutate(
    term = "Quartile 1",
      "OR_base" = '-', "95%CI_base" = '-', "p-value_base" = '', "p-value_heterogeneity_base" = "", 
      "OR_adjusted" = '-', "95%CI_adjusted" = '-', "p-value_adjusted" = '', "p-value_heterogeneity_adjusted" = "", 
    "OR_copollutant" = '-', "95%CI_copollutant" = '-', "p-value_copollutant" = '', "p-value_heterogeneity_copollutant" = "")
    

table_S4 <- results_POPs_ALS_occurrence$metanalysis$metanalysis_quart |>
  select(model, explanatory, term, OR, `95%CI`, "p-value", 'p-value_heterogeneity') |>
  pivot_wider(
    names_from = model,  
    values_from = c("OR", "95%CI", "p-value", 'p-value_heterogeneity')) |>
  bind_rows(extra_rows) |>
  select(Exposures = explanatory, Quartiles = term,
         contains("base"), 
         contains("adjusted"), 
         contains("copollutant")) |>
  mutate(
    Exposures = gsub("OCP_", "", Exposures), 
    Exposures = fct_relevel(Exposures, "PCB_4", "PCB_DL", "PCB_NDL", "HCB", "ΣDDT", "β_HCH", "Σchlordane"),
    Exposures = fct_recode(Exposures, 
                          "Most prevalent PCBs" = "PCB_4",
                          "Dioxin-like PCBs" = "PCB_DL",
                          "Non-dioxin-like PCBs" = "PCB_NDL",
                          "β-HCH" = "β_HCH")) |>
  arrange(Exposures, Quartiles) 

table_S4 <- table_S4 |> 
  rename('OR' = 'OR_base', '95% CI' = '95%CI_base', 'p-value' = 'p-value_base', "Hetero-geneity test" = "p-value_heterogeneity_base",
         'OR ' = 'OR_adjusted', '95% CI ' = '95%CI_adjusted', 'p-value ' = 'p-value_adjusted', "Hetero-geneity test " = "p-value_heterogeneity_adjusted",
         ' OR ' = 'OR_copollutant', ' 95% CI ' = '95%CI_copollutant', ' p-value ' = 'p-value_copollutant', " Hetero-geneity test " = "p-value_heterogeneity_copollutant") |>
  flextable() |>
  add_footer_lines(
  "1POPs were summed as follows: Dioxin-like PCBs corresponds to PCBs 118 and 156; non-dioxin-like PCBs corresponds to PCBs 28, 52, 74, 99, 101, 138, 153, 170, 180, 183, 187; most prevalent PCBs corresponds to PCBs 118, 138, 153, 180; ΣDDT corresponds to p,p’-DDT and p,p’-DDE and finally Σchlordane corresponds to trans-nonanchlor and oxychlordane.
  2Base and adjusted models are conditional logistic regressions matched for sex and age, and for municipality and serum thawing history (Finnish cohorts). Adjusted models further account for smoking, BMI, serum total cholesterol, marital status, and education (except for FMC cohort for which education data was not available). Co-pollutant models further include all chemicals, where the chemical of interest is categorized into quartiles, while the others are treated as smooth terms in GAMs.
  3Estimated risk of ALS when exposures to POP are at quartiles 2, 3, and 4, compared to quartile 1.
  4CI: Confidence interval.
  5P-value of between-study heterogeneity (Cochran’s Q test).") |>
  add_header(
    "Exposures" = "Exposures", "Quartiles" = "Quartiles",
    "OR" = "Base model", "95% CI" = "Base model", "p-value" = "Base model", "Hetero-geneity test" = "Base model",
    "OR " = "Adjusted model", "95% CI " = "Adjusted model", "p-value " = "Adjusted model", "Hetero-geneity test " = "Adjusted model",
    " OR " = "Copollutant model", " 95% CI " = "Copollutant model", " p-value " = "Copollutant model", " Hetero-geneity test " = "Copollutant model") |>
  merge_h(part = "header") |>
  merge_v(j = "Exposures") |>
  theme_vanilla() |>
  bold(j = "Exposures", part = "body") |>
  align(align = "center", part = "all") |>
  align(j = "Exposures", align = "left", part = "all") |> 
  merge_at(j = "Exposures", part = "header") |>
  merge_at(j = "Quartiles", part = "header") |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all")|>
  set_table_properties(align = "left")
table_S4
rm(extra_rows, table_S4)
```

# POPs ans ALS survival

## Subjects characteristics (among cases only)

```{r}
#| label: danish-POPs-ALS-survival-1
#| paged-print: false
results_descriptive$comp$covar_comp_cases |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

## Danish cohort

### Effects of subject characteristics on ALS survival

```{r}
#| label: danish-POPs-ALS-survival-2
results_POPs_ALS_survival$main_analysis$covar_danish |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Effects of POPs on ALS survival

#### Continuous coding (standardized)

```{r}
#| label: danish-POPs-ALS-survival-3
results_POPs_ALS_survival$main_analysis$POPs_sd_ALS_table_danish  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.height=4}
#| label: danish-POPs-ALS-survival-4
results_POPs_ALS_survival$main_analysis$POPs_sd_ALS_figure_danish
```

#### Quartiles

```{r}
#| label: danish-POPs-ALS-survival-5
#| column: screen-inset-right
results_POPs_ALS_survival$main_analysis$POPs_quart_ALS_table_danish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r fig.height=9, fig.width=9}
#| label: danish-POPs-ALS-survival-6
results_POPs_ALS_survival$main_analysis$POPs_quart_ALS_figure_danish
```

#### Cox-gam

```{r fig.height=11, fig.width=8}
#| label: danish-POPs-ALS-survival-7
wrap_plots(list(results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$PCB_4, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_4, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$PCB_DL, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_DL, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$PCB_NDL, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_NDL, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$OCP_HCB, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$OCP_HCB), 
           ncol = 2)

wrap_plots(list(results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$ΣDDT, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$ΣDDT, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$OCP_β_HCH, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$OCP_β_HCH, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$Σchlordane, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$Σchlordane, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$ΣPBDE, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$ΣPBDE), 
           ncol = 2)

```

#### Q-gcomp analysis

Including all the pollutants

```{r, fig.height=3}
#| label: danish-POPs-ALS-survival-8
results_POPs_ALS_survival$main_analysis$qgcomp$POPs_ALS_qgcomp_table_danish  |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
results_POPs_ALS_survival$main_analysis$qgcomp$POPs_ALS_qgcomp_figure_danish
```

Selecting the pollutants with a positive weight and adjusting for the others

```{r, fig.height=2}
#| label: danish-POPs-ALS-survival-9
results_POPs_ALS_survival$main_analysis$qgcomp$POPs_ALS_qgcomp_positive_table_danish  |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
results_POPs_ALS_survival$main_analysis$qgcomp$POPs_ALS_qgcomp_positive_figure_danish
```

## Sensitivity analyses 1

Mixture models with POPs pre-selection

### Model forcing the covariates to be included

Using LASSO

```{r}
#| label: danish-POPs-ALS-survival-10
plot(results_POPs_ALS_survival$sensi1$sensi1_lasso_sd_danish)  
```

Using ridge

```{r}
#| label: danish-POPs-ALS-survival-11
plot(results_POPs_ALS_survival$sensi1$sensi1_ridge_sd_danish)  
```

Using elastic net

```{r}
#| label: danish-POPs-ALS-survival-12
plot(results_POPs_ALS_survival$sensi1$sensi1_elastic_net_sd_danish)
```

These plot the cross-validation curve (red dotted line) along with upper and lower standard deviation curves along the 𝜆sequence (error bars). Two special values along the 𝜆sequence are indicated by the vertical dotted lines.

lambda.min is the value of 𝜆that gives minimum mean cross-validated error

coefficients using lambda.min Using LASSO

```{r}
#| label: danish-POPs-ALS-survival-13
coef(results_POPs_ALS_survival$sensi1$sensi1_lasso_sd_danish, s = "lambda.min")  
```

Using ridge

```{r}
#| label: danish-POPs-ALS-survival-14
coef(results_POPs_ALS_survival$sensi1$sensi1_ridge_sd_danish, s = "lambda.min")  
```

Using elastic net

```{r}
#| label: danish-POPs-ALS-survival-15
coef(results_POPs_ALS_survival$sensi1$sensi1_elastic_net_sd_danish, s = "lambda.min")
```

while lambda.1se is the value of 𝜆that gives the most regularized model such that the cross-validated error is within one standard error of the minimum.

coefficients using lambda.1se Using LASSO

```{r}
#| label: danish-POPs-ALS-survival-16
coef(results_POPs_ALS_survival$sensi1$sensi1_lasso_sd_danish, s = "lambda.1se")  
```

Using ridge

```{r}
#| label: danish-POPs-ALS-survival-17
coef(results_POPs_ALS_survival$sensi1$sensi1_ridge_sd_danish, s = "lambda.1se")  
```

Using elastic net

```{r}
#| label: danish-POPs-ALS-survival-18
coef(results_POPs_ALS_survival$sensi1$sensi1_elastic_net_sd_danish, s = "lambda.1se")
```

### New co-pollutant model with selected POPs using elastic net (to get unbiaised coefficients and 95%CI and p-values)

#### Continuous coding

```{r, fig.height=2}
#| label: danish-POPs-ALS-survival-19
results_POPs_ALS_survival$sensi1$POPs_sd_ALS_figure_sensi1_danish
results_POPs_ALS_survival$sensi1$POPs_sd_ALS_table_sensi1_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

#### Quartile coding

```{r, fig.height=3}
#| label: danish-POPs-ALS-survival-20
results_POPs_ALS_survival$sensi1$POPs_quart_ALS_figure_sensi1_danish
results_POPs_ALS_survival$sensi1$POPs_quart_ALS_table_sensi1_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

### ERS model with the POPs selestec with elastic net

#### Continuous coding (standardized)

```{r, fig.height=2}
#| label: danish-POPs-ALS-survival-21
results_POPs_ALS_survival$sensi1$POPs_sd_ALS_figure_sensi1_ERS_danish
results_POPs_ALS_survival$sensi1$POPs_sd_ALS_table_sensi1_ERS_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

#### Quartile coding

```{r, fig.height=3}
#| label: danish-POPs-ALS-survival-22
results_POPs_ALS_survival$sensi1$POPs_quart_ALS_figure_sensi1_ERS_danish
results_POPs_ALS_survival$sensi1$POPs_quart_ALS_table_sensi1_ERS_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

## Sensitivity analyses 2

### Model forcing the covariates to be included

coefficients using lambda.min (lambda.min is the value of 𝜆that gives minimum mean cross-validated error)

Using LASSO

```{r}
#| label: danish-POPs-ALS-survival-23
coef(results_POPs_ALS_survival$sensi2$sensi2_lasso_sd_danish, s = "lambda.min")  
```

Using elastic net

```{r}
#| label: danish-POPs-ALS-survival-24
coef(results_POPs_ALS_survival$sensi2$sensi2_elastic_net_sd_danish, s = "lambda.min")
```

### New co-pollutant model with selected POPs using elastic net (to get unbiaised coefficients and 95%CI and p-values)

#### Continuous coding (standardized)

```{r, fig.height=3}
#| label: danish-POPs-ALS-survival-25
results_POPs_ALS_survival$sensi2$POPs_sd_ALS_figure_sensi2_danish
results_POPs_ALS_survival$sensi2$POPs_sd_ALS_table_sensi2_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

#### Quartile coding

```{r}
#| label: danish-POPs-ALS-survival-26
results_POPs_ALS_survival$sensi2$POPs_quart_ALS_figure_sensi2_danish
results_POPs_ALS_survival$sensi2$POPs_quart_ALS_table_sensi2_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

### ERS model with the POPs selestec with elastic net

#### Continuous coding (standardized)

```{r, fig.height=2}
#| label: danish-POPs-ALS-survival-27
results_POPs_ALS_survival$sensi2$POPs_sd_ALS_figure_sensi2_ERS_danish
results_POPs_ALS_survival$sensi2$POPs_sd_ALS_table_sensi2_ERS_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

#### Quartile coding

```{r, fig.height=3, fig.width=8}
#| label: danish-POPs-ALS-survival-28
results_POPs_ALS_survival$sensi2$POPs_quart_ALS_figure_sensi2_ERS_danish
results_POPs_ALS_survival$sensi2$POPs_quart_ALS_table_sensi2_ERS_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

## Sensitivity analyses 3

```{r, fig.height=3, fig.width=8}
#| label: danish-POPs-ALS-survival-29
results_POPs_ALS_survival$sensi3$sensi3_figure
```

```{r}
#| label: danish-POPs-ALS-survival-30
results_POPs_ALS_survival$sensi3$sensi3_table
```

## Sensitivity analysis 4

### Cox models (sd)

```{r, fig.height=4}
#| label: danish-POPs-ALS-survival-31
results_POPs_ALS_survival$sensi4$POPs_sd_ALS_table_danish_sensi_4 |>
  flextable::set_table_properties(align = "left", layout = "autofit")
results_POPs_ALS_survival$sensi4$POPs_sd_ALS_figure_danish_sensi_4
```

### Cox-GAMs models

```{r fig.height=11, fig.width=8}
#| label: danish-POPs-ALS-survival-32
wrap_plots(list(results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_4,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$PCB_4,
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_DL,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$PCB_DL,
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_NDL,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$PCB_NDL,
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$OCP_HCB,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$OCP_HCB), 
           ncol = 2)

wrap_plots(list(
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$ΣDDT,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$ΣDDT, 
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$OCP_β_HCH,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$OCP_β_HCH,
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$Σchlordane,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$Σchlordane,
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$ΣPBDE,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$ΣPBDE), 
           ncol = 2)

```

## Sensitivity analysis 5: time trend effect?
```{r}
results_POPs_ALS_survival$sensi5$POPs_sd_ALS_figure_danish_sensi_5
```

## Sensitivity analysis 6: sex effect?
```{r}
results_POPs_ALS_survival$sensi6$POPs_sd_ALS_figure_danish_sensi_6
```

## Sensitivity analysis 7: age at diagnosis effect effect?
```{r}
results_POPs_ALS_survival$sensi7$POPs_sd_ALS_figure_danish_sensi_7
```

# Fatty acids and ALS occurence

## Danish cohort

### Effects of subject characterist on ALS occurence in the Danish cohort

```{r}
#| label: danish-fatty-acids-ALS-occurence-1
results_fattyacids_ALS_occurrence$danish$covar_als_table_danish  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### Effects of fatty acids on ALS occurence in the Danish cohort

#### Continuous coding (standardized)

```{r}
#| label: danish-fatty-acids-ALS-occurence-2
results_fattyacids_ALS_occurrence$danish$fattyacids_sd_als_table_danish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r}
#| label: danish-fatty-acids-ALS-occurence-3
results_fattyacids_ALS_occurrence$danish$fattyacids_sd_als_figure_danish
```

#### Quartiles

```{r}
#| label: danish-fatty-acids-ALS-occurence-4
results_fattyacids_ALS_occurrence$danish$fattyacids_quart_als_table_danish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.height=12}
#| label: danish-fatty-acids-ALS-occurence-5
results_fattyacids_ALS_occurrence$danish$fattyacids_quart_als_figure_danish
```

## Finnish cohorts

### Effects of fatty acids on ALS occurence in the Finnish cohorts

#### Continuous coding (standardized)

```{r}
#| label: finnish-fatty-acids-ALS-occurence-1
results_fattyacids_ALS_occurrence$finnish$fattyacids_sd_als_table_finnish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r}
#| label: finnish-fatty-acids-ALS-occurence-2
results_fattyacids_ALS_occurrence$finnish$fattyacids_sd_als_figure_finnish
```

#### Quartiles

```{r}
#| label: finnish-fatty-acids-ALS-occurence-3
results_fattyacids_ALS_occurrence$finnish$fattyacids_quart_als_table_finnish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.height=12}
#| label: finnish-fatty-acids-ALS-occurence-4
results_fattyacids_ALS_occurrence$finnish$fattyacids_quart_als_figure_finnish
```

### Sensitivity Ca

#### Continuous coding (standardized)

```{r}
#| label: finnish-fatty-acids-ALS-occurence-5
results_fattyacids_ALS_occurrence$sensi_ca$fattyacids_sd_als_table_sensi_ca_finnish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
results_fattyacids_ALS_occurrence$sensi_ca$fattyacids_sd_als_figure_sensi_ca_finnish
```

#### Quartiles

```{r}
#| label: finnish-fatty-acids-ALS-occurence-6
results_fattyacids_ALS_occurrence$sensi_ca$fattyacids_quart_als_table_sensi_ca_finnish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

```{r, fig.height=13}
#| label: finnish-fatty-acids-ALS-occurence-7
results_fattyacids_ALS_occurrence$sensi_ca$fattyacids_quart_als_figure_sensi_ca_finnish
```

# Fatty acids ans ALS survival

## Danish cohort

### Effects of covariates on ALS survival

```{r}
#| label: danish-fatty-acids-ALS-survival-1
results_fattyacids_ALS_survival$danish$covar_danish |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Continuous coding (standardized)

```{r}
#| label: danish-fatty-acids-ALS-survival-2
results_fattyacids_ALS_survival$danish$fattyacids_sd_als_table_danish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
results_fattyacids_ALS_survival$danish$fattyacids_sd_als_figure_danish
```

### Quartiles

```{r}
#| label: danish-fatty-acids-ALS-survival-3
results_fattyacids_ALS_survival$danish$fattyacids_quart_als_table_danish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.height=12}
#| label: danish-fatty-acids-ALS-survival-4
results_fattyacids_ALS_survival$danish$fattyacids_quart_als_figure_danish
```

```{r fig.height=12, fig.width=12}
#| label: danish-fatty-acids-ALS-survival-5
#| eval: false
#| column: screen-inset-right
#| include: false
arrange_ggsurvplots(
  results_fattyacids_ALS_survival$danish$survival_plots_danish[1:4],
  ncol = 2,              
  nrow = 2)
```

# NfL and ALS risk

## Main analysis

### Continuous coding (standardized)

```{r, fig.height=2}
#| label: danish-nfl-ALS-occurence-1
results_proteomic_ALS_occurrence$NfL_results$NfL_sd_ALS_figure_sensi_1 
```

```{r}
#| label: danish-nfl-ALS-occurence-2
results_proteomic_ALS_occurrence$NfL_results$NfL_sd_ALS_table_sensi_1 |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

### Quartiles

```{r}
#| label: danish-nfl-ALS-occurence-3
#| column: screen-inset-right
results_proteomic_ALS_occurrence$NfL_results$NfL_quart_ALS_table_sensi_1 |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```



## Sensitivity analysis: effect of follow_up duration

### Continuous (standardized)

```{r, fig.height=3}
#| label: danish-nfl-ALS-occurence-4
results_proteomic_ALS_occurrence$NfL_results$NfL_sd_ALS_figure_sensi_follow_up_base_adj
```

```{r}
#| label: danish-nfl-ALS-occurence-5
#| column: screen-inset-right
results_proteomic_ALS_occurrence$NfL_results$NfL_sd_ALS_table_sensi_follow_up_base_adj |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```


## Sensitivity analysis: effect of sex

### Continuous (standardized)

```{r}
#| label: danish-nfl-ALS-occurence-6
results_proteomic_ALS_occurrence$NfL_results$NfL_sd_ALS_table_sensi_sex_adj 
```



## Additionnal analysis: NfL over time

The locally estimated scatterplot smoothing (LOESS) curve illustrates the changes in the ratio of NEFL levels in each case–control pair/triplet over time. Each point corresponds to 1 case–control pair/triplet.

```{r}
#| label: danish-nfl-ALS-occurence-7
results_proteomic_ALS_occurrence$NfL_results$NfL_over_time$figure_NfL_over_time
```

After removing the NEFL outlier

```{r}
#| label: danish-nfl-ALS-occurence-8
results_proteomic_ALS_occurrence$NfL_results$NfL_over_time$figure_NfL_over_time_sensi_1
```

## Additional analysis: AUC

```{r, fig.height=8, fig.width=18}
#| label: danish-nfl-ALS-occurence-9
results_proteomic_ALS_occurrence$NfL_results$AUC_NfL$AUC_figure_adjusted_pattern +
  theme(legend.direction = "vertical", legend.position = "right")
results_proteomic_ALS_occurrence$NfL_results$AUC_NfL$AUC_figure_unadjusted_pattern +
  theme(legend.direction = "vertical", legend.position = "right")
```

## Additional analyses: removing individuals with other neurological conditions

```{r}
#| label: danish-nfl-ALS-occurence-10
results_proteomic_ALS_occurrence$NfL_results$other_disease$table_without_other_diseases
```

# NfL and ALS survival

## Main analyses

### Continuous coding (standardized)

```{r}
#| label: danish-nfl-ALS-survival-1
results_proteomic_ALS_survival$NfL_results$NfL_sd_ALS_table |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

### Quartiles

```{r}
#| label: danish-nfl-ALS-survival-2
#| column: screen-inset-right
results_proteomic_ALS_survival$NfL_results$NfL_quart_ALS_table 
```


## Sensitivity analysis: effect of follow-up duration

```{r, fig.height=4}
#| label: danish-nfl-ALS-survival-3
results_proteomic_ALS_survival$NfL_results$NfL_sd_ALS_figure_sensi_follow_up
```

```{r}
#| label: danish-nfl-ALS-survival-4
results_proteomic_ALS_survival$NfL_results$NfL_sd_ALS_table_sensi_follow_up |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

## Sensitivity analysis: effect of sex

```{r}
#| label: danish-nfl-ALS-survival-5
results_proteomic_ALS_survival$NfL_results$NfL_sd_ALS_table_sensi_sex
```


# Proteomic and ALS occurrence

## Effects of subject characterist on ALS occurence in the Danish cohort

```{r}
#| label: danish-proteomic-ALS-occurence-1
results_proteomic_ALS_occurrence$main$covar |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

## Main analysis

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-2
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_table |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-3
#| column: screen-inset-right
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_base_figure +  theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_adjusted_figure + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

Conditional logistic regressions (quartiles), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r}
#| label: danish-proteomic-ALS-occurence-4
#| column: screen-inset-right
results_proteomic_ALS_occurrence$main$proteomic_quart_ALS_table  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

Generalized additive models modeling proteins as smooth terms (thin plate splines), base models adjusted on sex and age at baseline. Adjusted models further account for smoking and body mass index

```{r, fig.width=9, fig.height=60}
#| label: danish-proteomic-ALS-occurence-5

all_vars_labels <- set_names(
  str_replace(proteomic,
              "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
              ""),
  proteomic)

plot_base_gam <- map(proteomic, make_gam_plot_base, data = bdd_danish)
names(plot_base_gam) <- proteomic
plot_base_gam <- list(
  signif = purrr::keep(plot_base_gam, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_base_gam, ~ .x$p_value > 0.05) |> map("plot"))


plot_adjusted_gam <- map(proteomic, make_gam_plot_adjusted, data = bdd_danish)
names(plot_adjusted_gam) <- proteomic
plot_adjusted_gam <- list(
  signif = purrr::keep(plot_adjusted_gam, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_adjusted_gam, ~ .x$p_value > 0.05) |> map("plot"))


plots_lists <- list(
  sensi_1_base = plot_base_gam$signif,
  sensi_1_adj  = plot_adjusted_gam$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(all_vars_labels, plots_lists, all_proteins, plot_grid)

```

## Sensitivity 1: Removing NEFL outlier

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-6
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_table_sensi_1 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-7
#| column: screen-inset-right
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_base_figure + labs(title = "Main analysis - Base model") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + labs(title = "NEFL outlier removed - Base model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
  
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_adjusted_figure + labs(title = "Main analysis - adjusted model") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + labs(title = "NEFL outlier removed - Adjusted model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

Conditional logistic regressions (quartiles), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r}
#| label: danish-proteomic-ALS-occurence-8
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_quart_ALS_table_sensi_1 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

Generalized additive models modeling proteins as smooth terms (thin plate splines), base models adjusted on sex and age at baseline. Adjusted models further account for smoking and body mass index

```{r, fig.width=10, fig.height=5}
#| label: danish-proteomic-ALS-occurence-9
proteomic_sensi_1 <-
  proteomic |> 
  str_replace("proteomic_neuro_explo_NEFL", "proteomic_neuro_explo_NEFL_sensi_1")

bdd_danish_sensi_1 <- 
  bdd_danish  |> 
  mutate(
    proteomic_neuro_explo_NEFL_sensi_1 = ifelse(match == 159, NA, proteomic_neuro_explo_NEFL))

all_vars_labels <- set_names(
  str_replace(proteomic_sensi_1,
              "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
              ""),
  proteomic_sensi_1)

plot_base_gam_sensi_1 <- map(proteomic_sensi_1, make_gam_plot_base, data = bdd_danish_sensi_1)
names(plot_base_gam_sensi_1) <- proteomic_sensi_1
plot_base_gam_sensi_1 <- list(
  signif = purrr::keep(plot_base_gam_sensi_1, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_base_gam_sensi_1, ~ .x$p_value > 0.05) |> map("plot"))


plot_adjusted_gam_sensi_1 <- map(proteomic_sensi_1, make_gam_plot_adjusted, data = bdd_danish_sensi_1)
names(plot_adjusted_gam_sensi_1) <- proteomic_sensi_1
plot_adjusted_gam_sensi_1 <- list(
  signif = purrr::keep(plot_adjusted_gam_sensi_1, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_adjusted_gam_sensi_1, ~ .x$p_value > 0.05) |> map("plot"))



wrap_plots(plot_base_gam$signif$proteomic_neuro_explo_NEFL,
           plot_base_gam_sensi_1$signif$proteomic_neuro_explo_NEFL_sensi_1, 
           ncol = 2)

wrap_plots(plot_adjusted_gam$signif$proteomic_neuro_explo_NEFL,
           plot_adjusted_gam_sensi_1$signif$proteomic_neuro_explo_NEFL_sensi_1, 
           ncol = 2)

rm(all_vars_labels, 
   bdd_danish_sensi_1, 
   proteomic_sensi_1)

```



## Sensitivity 2: Filtering to cases and their controls with follow-up \< 5 years

### Logistic regressions

```{r}
#| label: danish-proteomic-ALS-occurence-10
results_proteomic_ALS_occurrence$sensi_2$proteomic_sd_ALS_table_sensi_2 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r fig.height=6, fig.width=15}
#| label: danish-proteomic-ALS-occurence-11
#| column: screen-inset-right

results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_base_figure + labs(title = "Main analysis - Base model") + theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_2$proteomic_sd_ALS_base_figure_sensi_2 + labs(title = "Filtered to follow-up < 5 years - Base model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)

results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_adjusted_figure + labs(title = "Main analysis - Adjusted model") + theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_2$proteomic_sd_ALS_adjusted_figure_sensi_2 + labs(title = "Filtered to follow-up < 5 years - Adjusted model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

```{r}
#| label: danish-proteomic-ALS-occurence-12
#| column: screen-inset-right

results_proteomic_ALS_occurrence$sensi_3$proteomic_quart_ALS_table_sensi_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=60, fig.width=12}
#| label: danish-proteomic-ALS-occurence-13
#| column: screen-inset-right

bdd_danish_sensi_2 <- 
  bdd_danish |>
  group_by(match) |>                                                             
  filter(any(als == 1 & follow_up < 60)) |>                                     # sensi 2 : we remove cases (and their controls) with follow-up > 60 months 
  ungroup()


all_vars_labels <- set_names(
  str_replace(proteomic,
              "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
              ""),
  proteomic)

plot_base_gam_sensi_2 <- map(proteomic, make_gam_plot_base, data = bdd_danish_sensi_2)
names(plot_base_gam_sensi_2) <- proteomic
plot_base_gam_sensi_2 <- list(
  signif = purrr::keep(plot_base_gam_sensi_2, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_base_gam_sensi_2, ~ .x$p_value > 0.05) |> map("plot"))


plot_adjusted_gam_sensi_2 <- map(proteomic, make_gam_plot_adjusted, data = bdd_danish_sensi_2)
names(plot_adjusted_gam_sensi_2) <- proteomic
plot_adjusted_gam_sensi_2 <- list(
  signif = purrr::keep(plot_adjusted_gam_sensi_2, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_adjusted_gam_sensi_2, ~ .x$p_value > 0.05) |> map("plot"))



plots_lists <- list(
  sensi_1_base = plot_base_gam_sensi_1$signif,
  sensi_1_adj  = plot_adjusted_gam_sensi_1$signif,
  sensi_2_base = plot_base_gam_sensi_2$signif,
  sensi_2_adj  = plot_adjusted_gam_sensi_2$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 4) 

rm(plots_lists, all_proteins, plot_grid, 
   bdd_danish_sensi_2, all_vars_labels)

```

## Sensitivity 3: Filtering to cases and their controls with follow-up \> 5 years

### Logistic regressions

```{r}
#| label: danish-proteomic-ALS-occurence-14
results_proteomic_ALS_occurrence$sensi_3$proteomic_sd_ALS_table_sensi_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r fig.height=6, fig.width=15}
#| label: danish-proteomic-ALS-occurence-15
#| column: screen-inset-right

results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_base_figure + labs(title = "Main analysis - Base model") + theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_3$proteomic_sd_ALS_base_figure_sensi_3 + labs(title = "Filtered to follow-up > 5 years - Base model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)

results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_adjusted_figure + labs(title = "Main analysis - Adjusted model") + theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_3$proteomic_sd_ALS_adjusted_figure_sensi_3 + labs(title = "Filtered to follow-up > 5years - Adjusted model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

```{r}
#| label: danish-proteomic-ALS-occurence-16
#| column: screen-inset-right

results_proteomic_ALS_occurrence$sensi_3$proteomic_quart_ALS_table_sensi_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=60, fig.width=12}
#| label: danish-proteomic-ALS-occurence-17
#| column: screen-inset-right

bdd_danish_sensi_3 <- 
  bdd_danish |>
  group_by(match) |>
  filter(any(als == 1 & follow_up > 60)) |>                                     # had to remove 17 cases with follow-up<60 months (and their controls) -> 51 people
  ungroup() 

all_vars_labels <- set_names(
  str_replace(proteomic,
              "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
              ""),
  proteomic)

plot_base_gam_sensi_3 <- map(proteomic, make_gam_plot_base, data = bdd_danish_sensi_3)
names(plot_base_gam_sensi_3) <- proteomic
plot_base_gam_sensi_3 <- list(
  signif = purrr::keep(plot_base_gam_sensi_3, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_base_gam_sensi_3, ~ .x$p_value > 0.05) |> map("plot"))



plot_adjusted_gam_sensi_3 <- map(proteomic, make_gam_plot_adjusted, data = bdd_danish_sensi_3)
names(plot_adjusted_gam_sensi_3) <- proteomic
plot_adjusted_gam_sensi_3 <- list(
  signif = purrr::keep(plot_adjusted_gam_sensi_3, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_adjusted_gam_sensi_3, ~ .x$p_value > 0.05) |> map("plot"))


plots_lists <- list(
  sensi_1_base = plot_base_gam_sensi_1$signif,
  sensi_1_adj  = plot_adjusted_gam_sensi_1$signif,
  sensi_3_base = plot_base_gam_sensi_3$signif,
  sensi_3_adj  = plot_adjusted_gam_sensi_3$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 4) 

rm(plots_lists, all_proteins, plot_grid, 
   bdd_danish_sensi_3, all_vars_labels)

```

## Sensitivity 1 + 3: Removing NEFL outlier + filtering cases and their controls with follow-up \> 5 years

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-18
results_proteomic_ALS_occurrence$sensi_1_3$proteomic_sd_ALS_table_sensi_1_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-19
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Base model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3$proteomic_sd_ALS_base_figure_sensi_1_3 + 
  labs(title = "Filtered to follow-up>5years\n+ NEFL outlier excluded \n Base model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
  
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Adjusted model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3$proteomic_sd_ALS_adjusted_figure_sensi_1_3 + 
  labs(title = "Filtered to follow-up>5years\n+ NEFL outlier excluded \n Adjusted model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

Conditional logistic regressions (quartiles), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r}
#| label: danish-proteomic-ALS-occurence-20
#| column: screen-inset-right

results_proteomic_ALS_occurrence$sensi_1_3$proteomic_quart_ALS_table_sensi_1_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

Generalized additive models modeling proteins as smooth terms (thin plate splines), base models adjusted on sex and age at baseline. Adjusted models further account for smoking and body mass index

```{r fig.width=12, fig.height=60}
#| label: danish-proteomic-ALS-occurence-21
#| column: screen-inset-right

proteomic_sensi_1_3 <-
  proteomic |> 
  str_replace("proteomic_neuro_explo_NEFL", "proteomic_neuro_explo_NEFL_sensi_1_3")

bdd_danish_sensi_1_3 <- 
  bdd_danish |>
  group_by(match) |>                                                            # sensi 3 : we remove cases (and their controls) with follow-up <60 months to see if the association with NfL really happens a long time before ALS diagnosis 
  filter(any(als == 1 & follow_up > 60)) |>
  ungroup()|>
  mutate(                                                                       # sensi 1 : we remove NfL in match 159 because it's an outlier
    proteomic_neuro_explo_NEFL_sensi_1_3 =
      ifelse(match == 159, NA, proteomic_neuro_explo_NEFL))


all_vars_labels <- set_names(
  str_replace(proteomic_sensi_1_3,
              "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
              ""),
  proteomic_sensi_1_3)

plot_base_gam_sensi_1_3 <- map(proteomic_sensi_1_3, make_gam_plot_base, data = bdd_danish_sensi_1_3)
names(plot_base_gam_sensi_1_3) <- proteomic_sensi_1_3
plot_base_gam_sensi_1_3 <- list(
  signif = purrr::keep(plot_base_gam_sensi_1_3, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_base_gam_sensi_1_3, ~ .x$p_value > 0.05) |> map("plot"))



plot_adjusted_gam_sensi_1_3 <- map(proteomic_sensi_1_3, make_gam_plot_adjusted, data = bdd_danish_sensi_1_3)
names(plot_adjusted_gam_sensi_1_3) <- proteomic_sensi_1_3
plot_adjusted_gam_sensi_1_3 <- list(
  signif = purrr::keep(plot_adjusted_gam_sensi_1_3, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_adjusted_gam_sensi_1_3, ~ .x$p_value > 0.05) |> map("plot"))


plots_lists <- list(
  sensi_1_base = plot_base_gam_sensi_1$signif,
  sensi_1_adj  = plot_adjusted_gam_sensi_1$signif,
  sensi_1_3_base = plot_base_gam_sensi_1_3$signif,
  sensi_1_3_adj  = plot_adjusted_gam_sensi_1_3$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 4) 

rm(plots_lists, all_proteins, plot_grid, 
   bdd_danish_sensi_1_3, 
   all_vars_labels, 
   proteomic_sensi_1_3)

```

## Sensitivity 1 + 3 + 4: Removing NEFL outlier + filtering cases and their controls with follow-up \> 5 years + filtering follow-up \< 50%

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-22

results_proteomic_ALS_occurrence$sensi_1_3_4$proteomic_sd_ALS_table_sensi_1_3_4 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-23
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Base model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3_4$proteomic_sd_ALS_base_figure_sensi_1_3_4 + 
  labs(title = "NEFL outlier excluded\n+Filtered to follow-up>5years\n+Filtered to follow-up<50%\nBase model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
  
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Adjusted model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3_4$proteomic_sd_ALS_adjusted_figure_sensi_1_3_4 + 
  labs(title = "NEFL outlier excluded\n+Filtered to follow-up>5years\n+Filtered to follow-up<50%\nAdjusted model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

Conditional logistic regressions (quartiles), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r}
#| label: danish-proteomic-ALS-occurence-24
#| column: screen-inset-right

results_proteomic_ALS_occurrence$sensi_1_3_4$proteomic_quart_ALS_table_sensi_1_3_4 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

Generalized additive models modeling proteins as smooth terms (thin plate splines), base models adjusted on sex and age at baseline. Adjusted models further account for smoking and body mass index

```{r fig.width=12, fig.height=60}
#| label: danish-proteomic-ALS-occurence-25
#| column: screen-inset-right

proteomic_sensi_1_3_4 <-
  proteomic |> 
  str_replace("proteomic_neuro_explo_NEFL", "proteomic_neuro_explo_NEFL_sensi_1_3_4")

bdd_danish_sensi_1_3_4 <- 
  bdd_danish |>
  group_by(match) |>                                                              
  filter(any(als == 1 & follow_up > 60)) |>                                     # sensi 3 : we remove cases (and their controls) with follow-up < 60 months
  ungroup() |>
  mutate(                                                                       # sensi 1 : we remove NfL in match 159 because it's an outlier
    proteomic_neuro_explo_NEFL_sensi_1_3_4 = 
      ifelse(match == 159, NA, proteomic_neuro_explo_NEFL)) |>
  mutate(seuil = quantile(follow_up, 0.5, na.rm = TRUE)) |>                     # sensi 4 : we remove 50% of the highest values of follow-up (after removing the follow-up<5years)
  group_by(match) |>
  filter(any(als == 1 & follow_up <= seuil)) |>
  ungroup() |>
  select(-seuil) 


all_vars_labels <- set_names(
  str_replace(proteomic_sensi_1_3_4,
              "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
              ""),
  proteomic_sensi_1_3_4)

plot_base_gam_sensi_1_3_4 <- map(proteomic_sensi_1_3_4, make_gam_plot_base, data = bdd_danish_sensi_1_3_4)
names(plot_base_gam_sensi_1_3_4) <- proteomic_sensi_1_3_4
plot_base_gam_sensi_1_3_4 <- list(
  signif = purrr::keep(plot_base_gam_sensi_1_3_4, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_base_gam_sensi_1_3_4, ~ .x$p_value > 0.05) |> map("plot"))


plot_adjusted_gam_sensi_1_3_4 <- map(proteomic_sensi_1_3_4, make_gam_plot_adjusted, data = bdd_danish_sensi_1_3_4)
names(plot_adjusted_gam_sensi_1_3_4) <- proteomic_sensi_1_3_4
plot_adjusted_gam_sensi_1_3_4 <- list(
  signif = purrr::keep(plot_adjusted_gam_sensi_1_3_4, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_adjusted_gam_sensi_1_3_4, ~ .x$p_value > 0.05) |> map("plot"))


plots_lists <- list(
  sensi_1_base = plot_base_gam_sensi_1$signif,
  sensi_1_adj  = plot_adjusted_gam_sensi_1$signif,
  sensi_1_3_4_base = plot_base_gam_sensi_1_3_4$signif,
  sensi_1_3_4_adj  = plot_adjusted_gam_sensi_1_3_4$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 4) 

rm(plots_lists, all_proteins, plot_grid, 
   all_vars_labels, 
   proteomic_sensi_1_3_4,
   bdd_danish_sensi_1_3_4)

```

## Sensitivity 1 + 3 + 5: Removing NEFL outlier + filtering cases and their controls with follow-up \> 5 years + filtering follow-up \> 50%

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-26

results_proteomic_ALS_occurrence$sensi_1_3_5$proteomic_sd_ALS_table_sensi_1_3_5 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-27
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Base model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3_5$proteomic_sd_ALS_base_figure_sensi_1_3_5 + 
  labs(title = "NEFL outlier excluded\n+Filtered to follow-up>5years\n+Filtered to follow-up>50%\nBase model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
  
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Adjusted model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3_5$proteomic_sd_ALS_adjusted_figure_sensi_1_3_5 + 
  labs(title = "NEFL outlier excluded\n+Filtered to follow-up>5years\n+Filtered to follow-up>50%\nAdjusted model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

Conditional logistic regressions (quartiles), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r}
#| label: danish-proteomic-ALS-occurence-28
#| column: screen-inset-right

results_proteomic_ALS_occurrence$sensi_1_3_5$proteomic_quart_ALS_table_sensi_1_3_5 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

Generalized additive models modeling proteins as smooth terms (thin plate splines), base models adjusted on sex and age at baseline. Adjusted models further account for smoking and body mass index

```{r fig.width=12, fig.height=60}
#| label: danish-proteomic-ALS-occurence-29
#| column: screen-inset-right


proteomic_sensi_1_3_5 <-
  proteomic |> 
  str_replace("proteomic_neuro_explo_NEFL", "proteomic_neuro_explo_NEFL_sensi_1_3_5")


bdd_danish_sensi_1_3_5 <- 
  bdd_danish |>
  group_by(match) |>                                                            # sensi 3 : we remove cases (and their controls) with follow-up < 60 months  
  filter(any(als == 1 & follow_up > 60)) |>
  ungroup()|>
  mutate(                                                                       # sensi 1 : we remove NfL in match 159 because it's an outlier
    proteomic_neuro_explo_NEFL_sensi_1_3_5 = 
      ifelse(match == 159, NA, proteomic_neuro_explo_NEFL)) |>
  mutate(seuil = quantile(follow_up, 0.5, na.rm = TRUE)) |>                     # sensi 5 : we remove 50% of the lowest values of follow-up (after removing the follow-up<5years)
  group_by(match) |>
  filter(any(als == 1 & follow_up > seuil)) |>
  ungroup() |>
  select(-seuil)

all_vars_labels <- set_names(
  str_replace(proteomic_sensi_1_3_5,
              "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
              ""),
  proteomic_sensi_1_3_5)

plot_base_gam_sensi_1_3_5 <- map(proteomic_sensi_1_3_5, make_gam_plot_base, data = bdd_danish_sensi_1_3_5)
names(plot_base_gam_sensi_1_3_5) <- proteomic_sensi_1_3_5
plot_base_gam_sensi_1_3_5 <- list(
  signif = purrr::keep(plot_base_gam_sensi_1_3_5, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_base_gam_sensi_1_3_5, ~ .x$p_value > 0.05) |> map("plot"))


plot_adjusted_gam_sensi_1_3_5 <- map(proteomic_sensi_1_3_5, make_gam_plot_adjusted, data = bdd_danish_sensi_1_3_5)
names(plot_adjusted_gam_sensi_1_3_5) <- proteomic_sensi_1_3_5
plot_adjusted_gam_sensi_1_3_5 <- list(
  signif = purrr::keep(plot_adjusted_gam_sensi_1_3_5, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_adjusted_gam_sensi_1_3_5, ~ .x$p_value > 0.05) |> map("plot"))


plots_lists <- list(
  sensi_1_base = plot_base_gam_sensi_1$signif,
  sensi_1_adj  = plot_adjusted_gam_sensi_1$signif,
  sensi_1_3_5_base = plot_base_gam_sensi_1_3_5$signif,
  sensi_1_3_5_adj  = plot_adjusted_gam_sensi_1_3_5$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 4) 

rm(plots_lists, all_proteins, plot_grid, 
   all_vars_labels, 
   proteomic_sensi_1_3_5,
   bdd_danish_sensi_1_3_5)

```

## Sensitivity 6: stratifiyng the analysis in tertiles of follow-up duration

Data not shown because of merging sensi 1 and sensi 6

```{r fig.height=12, fig.width=15}
#| label: danish-proteomic-ALS-occurence-30
#| eval: false
#| include: false

### Logistic regressions
# Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_base_figure + labs(title = "Main analysis - full sample size") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_base_figure_sensi_6_T1 + labs(title = "Filtered to follow-up T1 (<11 years)") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_base_figure_sensi_6_T2 + labs(title = "Filtered to follow-up T2 (11-16 years)") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_base_figure_sensi_6_T3 + labs(title = "Filtered to follow-up T3 (>16 years)") + theme(legend.position = "bottom") 
  
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_adjusted_figure + labs(title = "Main analysis - full sample size") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_adjusted_figure_sensi_6_T1 + labs(title = "Filtered to follow-up T1 (<11 years)") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_adjusted_figure_sensi_6_T2 + labs(title = "Filtered to follow-up T2 (11-16 years)") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_adjusted_figure_sensi_6_T3 + labs(title = "Filtered to follow-up T3 (>16 years)") + theme(legend.position = "bottom") 
```

## Sensitivity 1 + 6: Removing NEFL outlier + stratifiyng the analysis in tertiles of follow-up duration

```{r}
#| label: danish-proteomic-ALS-occurence-31
results_proteomic_ALS_occurrence$sensi_6$sensi_6_table_follow_up
```

```{r, fig.height=3, fig.width=4}
#| label: danish-proteomic-ALS-occurence-32
results_proteomic_ALS_occurrence$sensi_6$sensi_6_densityplot_follow_up
```

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r fig.width=15, fig.height=12}
#| label: danish-proteomic-ALS-occurence-33
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + labs(title = "NEFL outlier removed \n Base model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_base_figure_sensi_1_6_T1 + labs(title = "Filtered to follow-up T1 (<11 years)\n+ NEFL outlier excluded \n Base model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_base_figure_sensi_1_6_T2 + labs(title = "Filtered to follow-up T2 (11-16 years)\n+ NEFL outlier excluded \n Base model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_base_figure_sensi_1_6_T3 + labs(title = "Filtered to follow-up T3 (>16 years)\n+ NEFL outlier excluded \n Base model") + theme(legend.position = "bottom")  
  
  
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + labs(title = "NEFL outlier removed \n Adjusted model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_adjusted_figure_sensi_1_6_T1 + labs(title = "Filtered to follow-up T1 (<11 years)\n+ NEFL outlier excluded \n Adjusted model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_adjusted_figure_sensi_1_6_T2 + labs(title = "Filtered to follow-up T2 (11-16 years)\n+ NEFL outlier excluded \n Adjusted model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_adjusted_figure_sensi_1_6_T3 + labs(title = "Filtered to follow-up T3 (>16 years)\n+ NEFL outlier excluded \n Adjusted model") + theme(legend.position = "bottom")  
```

NEFL is not significantly associated with ALS occurence when the follow-up duration is \>16 years. We don't see many associations in the follow-up group T1: I suspect that it is because of high heterogeneity in follow-up duration among this sub-group (0-11 years). --\> then I trust more the sensi 2 results where we just remove the cases (and their controls) with follow-up \<5 years.

## Sensitivity 1 + 7: Removing NEFL outlier + stratifiyng the analysis on sex

### Continuous coding (standardized)

```{r}
#| label: danish-proteomic-ALS-occurence-34
results_proteomic_ALS_occurrence$sensi_1_7$proteomic_sd_ALS_table_sensi_1_7
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-occurence-35
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Base model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_7$proteomic_sd_ALS_base_figure_sensi_1_7  + 
  labs(title = "NEFL outlier excluded\n+Filtered on sex") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
  
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Adjusted model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_7$proteomic_sd_ALS_adjusted_figure_sensi_1_7 + 
  labs(title = "NEFL outlier excluded\n+Filtered on sex") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)

```

### Quartiles

```{r}
#| label: danish-proteomic-ALS-occurence-36
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1_7$proteomic_quart_ALS_table_sensi_1_7
```

### GAMs

```{r fig.width=9, fig.height=70}
#| label: danish-proteomic-ALS-occurence-37
#| column: screen-inset-right

proteomic_sensi_1_7_female <-
  proteomic |> 
  str_replace("proteomic_neuro_explo_NEFL", "proteomic_neuro_explo_NEFL_sensi_1_7_female")

bdd_danish_sensi_1_7_female <- 
  bdd_danish |>
  group_by(match) |>                                                            # sensi 7 : we purrr::keep cases (and their controls) that are females 
  filter(any(als == 1 & sex == "Female")) |>
  ungroup()|>
  mutate(                                                                       # sensi 1 : we remove NfL in match 159 because it's an outlier
    proteomic_neuro_explo_NEFL_sensi_1_7_female =
      ifelse(match == 159, NA, proteomic_neuro_explo_NEFL))



all_vars_labels <- set_names(
  str_replace(proteomic_sensi_1_7_female,
              "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
              ""),
  proteomic_sensi_1_7_female)


plot_base_gam_sensi_1_7_female <- map(proteomic_sensi_1_7_female, make_gam_plot_base_sex, data = bdd_danish_sensi_1_7_female)
names(plot_base_gam_sensi_1_7_female) <- proteomic_sensi_1_7_female
plot_base_gam_sensi_1_7_female <- list(
  signif = purrr::keep(plot_base_gam_sensi_1_7_female, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_base_gam_sensi_1_7_female, ~ .x$p_value > 0.05) |> map("plot"))


plot_adjusted_gam_sensi_1_7_female <- map(proteomic_sensi_1_7_female, make_gam_plot_adjusted_sex, data = bdd_danish_sensi_1_7_female)
names(plot_adjusted_gam_sensi_1_7_female) <- proteomic_sensi_1_7_female
plot_adjusted_gam_sensi_1_7_female <- list(
  signif = purrr::keep(plot_adjusted_gam_sensi_1_7_female, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = purrr::keep(plot_adjusted_gam_sensi_1_7_female, ~ .x$p_value > 0.05) |> map("plot"))


proteomic_sensi_1_7_male <-
  proteomic |> 
  str_replace("proteomic_neuro_explo_NEFL", "proteomic_neuro_explo_NEFL_sensi_1_7_male")

bdd_danish_sensi_1_7_male <- 
  bdd_danish |>
  group_by(match) |>                                                            # sensi 7 : we keep cases (and their controls) that are male 
  filter(any(als == 1 & sex == "Male")) |>
  ungroup()|>
  mutate(                                                                       # sensi 1 : we remove NfL in match 159 because it's an outlier
    proteomic_neuro_explo_NEFL_sensi_1_7_male =
      ifelse(match == 159, NA, proteomic_neuro_explo_NEFL))


all_vars_labels <- set_names(
  str_replace(proteomic_sensi_1_7_male,
              "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
              ""),
  proteomic_sensi_1_7_male)

plot_base_gam_sensi_1_7_male <- map(proteomic_sensi_1_7_male, make_gam_plot_base_sex, data = bdd_danish_sensi_1_7_male)
names(plot_base_gam_sensi_1_7_male) <- proteomic_sensi_1_7_male
plot_base_gam_sensi_1_7_male <- list(
  signif = keep(plot_base_gam_sensi_1_7_male, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = keep(plot_base_gam_sensi_1_7_male, ~ .x$p_value > 0.05) |> map("plot"))


plot_adjusted_gam_sensi_1_7_male <- map(proteomic_sensi_1_7_male, make_gam_plot_adjusted_sex, data = bdd_danish_sensi_1_7_male)
names(plot_adjusted_gam_sensi_1_7_male) <- proteomic_sensi_1_7_male
plot_adjusted_gam_sensi_1_7_male <- list(
  signif = keep(plot_adjusted_gam_sensi_1_7_male, ~ .x$p_value <= 0.05) |> map("plot"),
  not_signif = keep(plot_adjusted_gam_sensi_1_7_male, ~ .x$p_value > 0.05) |> map("plot"))


plots_lists <- list(
  sensi_1_adj  = plot_adjusted_gam_sensi_1$signif,
  sensi_1_7_female = plot_base_gam_sensi_1_7_female$signif,
  sensi_1_7_male  = plot_base_gam_sensi_1_7_male$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 3) 

rm(plots_lists, all_proteins, plot_grid, 
   all_vars_labels, 
   proteomic_sensi_1_7_female,
   proteomic_sensi_1_7_male,
   bdd_danish_sensi_1_7_female,
   bdd_danish_sensi_1_7_male)


```

```{r}
#| label: danish-proteomic-ALS-occurence-38
rm(   make_gam_plot_base, 
      make_gam_plot_adjusted,
      make_gam_plot_base_sex, 
      make_gam_plot_adjusted_sex, 
      plot_base_gam, 
      plot_adjusted_gam, 
      plot_base_gam_sensi_1, 
      plot_adjusted_gam_sensi_1, 
      plot_base_gam_sensi_2, 
      plot_adjusted_gam_sensi_2, 
      plot_base_gam_sensi_3, 
      plot_adjusted_gam_sensi_3, 
      plot_base_gam_sensi_1_3, 
      plot_adjusted_gam_sensi_1_3, 
      plot_base_gam_sensi_1_3_4, 
      plot_adjusted_gam_sensi_1_3_4, 
      plot_base_gam_sensi_1_3_5, 
      plot_adjusted_gam_sensi_1_3_5, 
      plot_base_gam_sensi_1_7_female, 
      plot_base_gam_sensi_1_7_male, 
      plot_adjusted_gam_sensi_1_7_female, 
      plot_adjusted_gam_sensi_1_7_male)
```

## Additionnal analysis 1

```{r, fig.height=3, fig.width=3}
#| label: danish-proteomic-ALS-occurence-39
shapiro.test(results_proteomic_ALS_occurrence$additional_analysis_1$OR_distribution$OR_log)

results_proteomic_ALS_occurrence$additional_analysis_1$boxplot_OR
results_proteomic_ALS_occurrence$additional_analysis_1$densityplot_OR
```

## Additionnal analysis 2

Testing the distribution of the p-values depending on OR\>=1 to the distribution of the p-values depending on OR\<1 using a Fasano-Franceschini test

The Fasano–Franceschini test is a multivariate extension of the Kolmogorov–Smirnov test. While the classical KS test only compares one-dimensional distributions, the Fasano–Franceschini test can compare two-dimensional or multidimensional distributions.

It works by evaluating, for every point of one dataset, how the cumulative distribution differs from the other dataset in the four quadrants around that point. The test statistic D represents the largest difference between the empirical multivariate distributions of the two datasets. A large value of D indicates that the two datasets occupy different regions of the multidimensional space (e.g., different shapes, locations, or dispersions).

The null hypothesis (H₀) is: H₀: The two datasets come from the same multivariate distribution. A small p-value indicates a significant deviation between the two distributions.

### all proteins

```{r, fig.height=3, fig.width=9}
#| label: danish-proteomic-ALS-occurence-40
#| column: screen-inset-right
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_all_figure
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_all_results

results_proteomic_ALS_occurrence$additional_analysis_2$jackknife_all_results |> 
  select(d_stat_percent) |> 
  mutate(d_stat_percent = round(d_stat_percent, 1)) |>
  tbl_summary()


```

Because the p-value is 0.00587, which is well below the standard 0.05 threshold, we reject the null hypothesis that the two sets of results (OR ≥ 1 vs. OR \< 1) come from the same distribution.

This means that the two groups of proteins, those associated with increased risk and those associated with decreased risk, show significantly different joint distributions of odds ratios and p-values. Therefore, the asymmetric patterns we observe are unlikely to be due to random fluctuation caused by multiple testing. The results support the idea that the proteins associated with increased disease risk and those associated with decreased disease risk represent genuinely different biological signals, rather than artifacts of high-dimensional multiple testing.

### Immune response

```{r, fig.height=3, fig.width=9}
#| label: danish-proteomic-ALS-occurence-41
#| column: screen-inset-right
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_immune_figure
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_immune_results

results_proteomic_ALS_occurrence$additional_analysis_2$jackknife_immune_results |> 
  select(d_stat_percent) |> 
  mutate(d_stat_percent = round(d_stat_percent, 1)) |>
  tbl_summary()
```

### Metbolism

```{r, fig.height=3, fig.width=9}
#| label: danish-proteomic-ALS-occurence-42
#| column: screen-inset-right
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_metabolism_figure
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_metabolism_results

results_proteomic_ALS_occurrence$additional_analysis_2$jackknife_metabolism_results |> 
  select(d_stat_percent) |> 
  mutate(d_stat_percent = round(d_stat_percent, 1)) |>
  tbl_summary()

```

### Neuro explo

```{r, fig.height=3, fig.width=9}
#| label: danish-proteomic-ALS-occurence-43
#| column: screen-inset-right
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_neuro_figure
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_neuro_results

results_proteomic_ALS_occurrence$additional_analysis_2$jackknife_neuro_results |> 
  select(d_stat_percent) |> 
  mutate(d_stat_percent = round(d_stat_percent, 1)) |>
  tbl_summary()
```

## Additional analysis 3 - superlearner

### Test 1 a : method AUC, many algorithms, no pre screening, default hyperparameters

```{r}
#| label: danish-proteomic-ALS-occurence-44
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_a)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_a) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1a)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

Leeklasso is doing the best job --\> means that the relations are probably linear and for only a few proteins (because this algorithm has a internal prescreening)

Trees methodologies are not doing a great job but there are supposed to handle correctly high dimension: maybe not a lot of interactions between the proteins? But should investigate with tunned parameters

Algorithms such as SL.bayesglm, SL.stepAIC, SL.ksvm, SL.gam, SL.polymars and SL.earth should do better wit pre screening

### Defining the best hyperparameter combinaison for each algorithm

#### SL.glmnet

```{r}
#| label: danish-proteomic-ALS-occurence-45
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_glmnet)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_glmnet) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.glmnet)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.ranger

```{r}
#| label: danish-proteomic-ALS-occurence-46
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_ranger)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_ranger) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.ranger)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.xgboost

```{r}
#| label: danish-proteomic-ALS-occurence-47
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_xgboost)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_xgboost) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.xgboost)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.ksvm

```{r}
#| label: danish-proteomic-ALS-occurence-48
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_ksvm)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_ksvm) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.ksvm)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.earth

```{r}
#| label: danish-proteomic-ALS-occurence-49
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_earth)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_earth) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.earth)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

### Test 1 b : method AUC, many algorithms, pre screening, optimized hyperparameters

```{r}
#| label: danish-proteomic-ALS-occurence-50
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_b_opt_20)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_b_opt_20) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1b top 20)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

```{r}
#| label: danish-proteomic-ALS-occurence-51
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_b_opt_10)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_b_opt_10) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1b top 10)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

### Test 1 c : method AUC, selected algorithms, pre screening, optimized hyperparameters

```{r}
#| label: danish-proteomic-ALS-occurence-52
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_20)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_20) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1c top 20)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

```{r}
#| label: danish-proteomic-ALS-occurence-53
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_10)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_10) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1c top 10)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

### Weights of the algorithms (sl_fit_CV_1_c_opt_20)

```{r}
#| label: danish-proteomic-ALS-occurence-54
# Calcul des poids moyens en pourcentage
all_weights <- lapply(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_20$AllSL, function(x) x$coef)
weights_matrix <- do.call(rbind, all_weights)
colnames(weights_matrix) <- results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_20$libraryNames

df_weights <- data.frame(
  Algorithms = colnames(weights_matrix),
  Mean_weights = colMeans(weights_matrix) * 100) |> 
  dplyr::arrange(desc(Mean_weights))

df_weights |>
  flextable() |>
  set_header_labels(Mean_weights = "Mean weights (%)") |>
  colformat_double(j = "Mean_weights", digits = 1) |>
  flextable::font(fontname = "Calibri", part = "all") |>
  fontsize(size = 10, part = "all") |>
  autofit()

rm(all_weights, weights_matrix, df_weights)
```

### Selected proteins by leeklasso (sl_fit_CV_1_c_All)

```{r}
#| label: danish-proteomic-ALS-occurence-55
# Extraction des protéines et coefficients pour Leekasso
leek_results_list <- lapply(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_20$AllSL, function(fold_fit) {
  obj <- fold_fit$fitLibrary$SL.leekasso_All$object
  if(is.null(obj)) return(NULL)
  
  coefs <- coef(obj)
  coefs <- coefs[names(coefs) != "(Intercept)"]
  
  data.frame(
    Protein = names(coefs),
    Log_Odds = as.numeric(coefs)
  )
})

df_leek_final <- do.call(rbind, leek_results_list) |>
  dplyr::group_by(Protein) |>
  dplyr::summarise(
    Consistency = n() / 10 * 100,
    Mean_OR = exp(mean(Log_Odds)) ) |>
  dplyr::arrange(desc(Consistency))


df_leek_final |>
  flextable() |>
  set_header_labels(
    Protein = "Selected proteins",
    Consistency = "Consistency across folds (%)", 
    Mean_OR = "Mean ORs across folds") |>
  colformat_double(j = "Mean_OR", digits = 2) |>
  colformat_double(j = "Consistency", digits = 1) |>
  flextable::font(fontname = "Calibri", part = "all") |>
  fontsize(size = 10, part = "all") |>
  autofit()

rm(leek_results_list, df_leek_final)
```

### Selected proteins by GLMNet (sl_fit_CV_1_c_All)

```{r}
#| label: danish-proteomic-ALS-occurence-56
# Extraction des protéines selectionnées par GLMNet dans chaque pli
glm_results_list <- lapply(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_20$AllSL, function(fold_fit) {
  fit_glmnet <- fold_fit$fitLibrary$SL.glmnet.opt_All$object
  coefs <- as.matrix(coef(fit_glmnet, s = "lambda.min"))
  
  data.frame(
    Protein = rownames(coefs),
    Log_Odds = as.numeric(coefs)
  ) |>
    dplyr::filter(Protein != "(Intercept)")
})

df_glm_final <- do.call(rbind, glm_results_list) |>
  dplyr::group_by(Protein) |>
  dplyr::summarise(
    Consistency = sum(Log_Odds != 0) / 10 * 100,
    Mean_OR = exp(mean(Log_Odds[Log_Odds != 0]))) |>
  dplyr::filter(Consistency > 0) |>
  dplyr::arrange(desc(Consistency))

df_glm_final |>
  flextable() |>
  set_header_labels(
    Protein = "Proteine",
    Consistency = "Consistency across folds (%)", 
    Mean_OR = "Mean ORs across folds") |>
  colformat_double(j = "Mean_OR", digits = 2) |>
  colformat_double(j = "Consistency", digits = 1) |>
  flextable::font(fontname = "Calibri", part = "all") |>
  fontsize(size = 10, part = "all") |>
  autofit()

rm(glm_results_list, df_glm_final)
```

APEX1: DNA repair enzyme (Apurinic/Apyrimidinic Endonuclease 1)

NPTXR: Neuronal Pentraxin Receptor (involved in synaptic remodeling and glutamatergic signaling)

NfL: Marker of neuroaxonal damage

ARG1 (Arginase-1): Marker of alternative (M2) macrophage/microglial polarization

CD33: Myeloid cell surface antigen (Microglial marker)

PIK3AP1 (Phosphoinositide 3-kinase adapter protein 1): link between B-cell receptors and the PI3K signaling pathway

PSIP1 (PC4 and SFRS1-interacting protein 1): transcriptional coactivator that promotes cell survival (protecting cells against oxidative stress)

GLRX (Glutaredoxin-1): small redox enzyme that uses glutathione to reduce oxidized proteins (disulfides) (maintaining redox homeostasis)

IL-32 (Interleukin-32): Pro-inflammatory cytokine

NPM1 (Nucleophosmin): multifunctional "chaperone" protein found mainly in the nucleolus. it is involved in ribosome biogenesis, DNA repair, and protein folding.

GHRL (Ghrelin): metabolic peptide produced mainly in the stomach that regulates appetite and energy balance.

CD302: C-type lectin receptor involved in cell adhesion and migration, primarily expressed on myeloid cells like macrophages and dendritic cells.

PADI2: Protein-arginine deiminase type-2

### Test 2: + time as a predictor

### Defining the best hyperparameter combinaison for each algorithm

#### SL.glmnet

```{r}
#| label: danish-proteomic-ALS-occurence-57
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_glmnet)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_glmnet) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.glmnet)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.ranger

```{r}
#| label: danish-proteomic-ALS-occurence-58
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_ranger)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_ranger) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.ranger)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.xgboost

```{r}
#| label: danish-proteomic-ALS-occurence-59
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_xgboost)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_xgboost) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.xgboost)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.ksvm

```{r}
#| label: danish-proteomic-ALS-occurence-60
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_ksvm)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_ksvm) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.ksvm)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.earth

```{r}
#| label: danish-proteomic-ALS-occurence-61
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_earth)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_earth) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.earth)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

### Test 2 b : method AUC, many algorithms, pre screening, optimized hyperparameters

```{r}
#| label: danish-proteomic-ALS-occurence-62
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_b_opt_20)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_b_opt_20) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1b top 20)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

```{r}
#| label: danish-proteomic-ALS-occurence-63
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_b_opt_10)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_b_opt_10) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1b top 10)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

### Test 2 c : method AUC, selected algorithms, pre screening, optimized hyperparameters

```{r}
#| label: danish-proteomic-ALS-occurence-64
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_20)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_20) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1c top 20)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

```{r}
#| label: danish-proteomic-ALS-occurence-65
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_10)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_10) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1c top 10)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```


### Weights of the algorithms ()

```{r}
#| label: danish-proteomic-ALS-occurence-66
# Calcul des poids moyens en pourcentage
all_weights <- lapply(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_10$AllSL, function(x) x$coef)
weights_matrix <- do.call(rbind, all_weights)
colnames(weights_matrix) <- results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_10$libraryNames

df_weights <- data.frame(
  Algorithms = colnames(weights_matrix),
  Mean_weights = colMeans(weights_matrix) * 100) |> 
  dplyr::arrange(desc(Mean_weights))

df_weights |>
  flextable() |>
  set_header_labels(Mean_weights = "Mean weights (%)") |>
  colformat_double(j = "Mean_weights", digits = 1) |>
  flextable::font(fontname = "Calibri", part = "all") |>
  fontsize(size = 10, part = "all") |>
  autofit()

rm(all_weights, weights_matrix, df_weights)
```

### Selected proteins by leeklasso

```{r}
#| label: danish-proteomic-ALS-occurence-67
# Extraction des protéines et coefficients pour Leekasso
leek_results_list <- lapply(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_10$AllSL, function(fold_fit) {
  obj <- fold_fit$fitLibrary$SL.leekasso_All$object
  if(is.null(obj)) return(NULL)
  
  coefs <- coef(obj)
  coefs <- coefs[names(coefs) != "(Intercept)"]
  
  data.frame(
    Protein = names(coefs),
    Log_Odds = as.numeric(coefs)
  )
})

df_leek_final <- do.call(rbind, leek_results_list) |>
  dplyr::group_by(Protein) |>
  dplyr::summarise(
    Consistency = n() / 10 * 100,
    Mean_OR = exp(mean(Log_Odds)) ) |>
  dplyr::arrange(desc(Consistency))


df_leek_final |>
  flextable() |>
  set_header_labels(
    Protein = "Selected proteins",
    Consistency = "Consistency across folds (%)", 
    Mean_OR = "Mean ORs across folds") |>
  colformat_double(j = "Mean_OR", digits = 2) |>
  colformat_double(j = "Consistency", digits = 1) |>
  flextable::font(fontname = "Calibri", part = "all") |>
  fontsize(size = 10, part = "all") |>
  autofit()

rm(leek_results_list, df_leek_final)
```

### Selected proteins by GLMNet

```{r}
#| label: danish-proteomic-ALS-occurence-68
# Extraction des protéines selectionnées par GLMNet dans chaque pli
glm_results_list <- lapply(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_10$AllSL, function(fold_fit) {
  fit_glmnet <- fold_fit$fitLibrary$SL.glmnet.a0.9_Al$object
  coefs <- as.matrix(coef(fit_glmnet, s = "lambda.min"))
  
  data.frame(
    Protein = rownames(coefs),
    Log_Odds = as.numeric(coefs)
  ) |>
    dplyr::filter(Protein != "(Intercept)")
})

df_glm_final <- do.call(rbind, glm_results_list) |>
  dplyr::group_by(Protein) |>
  dplyr::summarise(
    Consistency = sum(Log_Odds != 0) / 10 * 100,
    Mean_OR = exp(mean(Log_Odds[Log_Odds != 0]))) |>
  dplyr::filter(Consistency > 0) |>
  dplyr::arrange(desc(Consistency))

df_glm_final |>
  flextable() |>
  set_header_labels(
    Protein = "Proteine",
    Consistency = "Consistency across folds (%)", 
    Mean_OR = "Mean ORs across folds") |>
  colformat_double(j = "Mean_OR", digits = 2) |>
  colformat_double(j = "Consistency", digits = 1) |>
  flextable::font(fontname = "Calibri", part = "all") |>
  fontsize(size = 10, part = "all") |>
  autofit()

rm(glm_results_list, df_glm_final)
```

## Additional analysis 4 - tidymodels 

### Test 1: Can plasma proteins predict ALS? (proteins + covariates: birth year, sex, BMI, smoking)

For each test, the pipeline follows two phases:

-   Hyperparameter tuning: each algorithm is independently optimized via cross-validation.

-   Model comparison: the best-tuned version of each algorithm is evaluated on the same folds and ranked by AUC.

All models use a grouped 10-cross-validation respecting the matching structure: all 3 individuals from a same triplet (1 case + 2 controls) are always kept on the same side — either all in the training set or all in the test set. Without this grouping, the model could learn patterns tied to the matching variables (birth year and sex) rather than the true proteomic signal, artificially inflating AUC.

#### Penalized Logistic Regression (glmnet) tuning
Standard logistic regression would fail here because p \> n (\~280 predictors, 166 cases). glmnet solves this by adding a penalty term to the loss function that shrinks coefficients toward zero, effectively performing implicit variable selection.

Two hyperparameters are tuned:

-   penalty (lambda): Strength of regularization. Higher lambda = more coefficients forced to zero.

-   mixture (alpha): Type of penalty.

    -   alpha = 1 = Lasso (exact zeros, variable selection).

    -   alpha = 0 = Ridge (all coefficients shrunk but non-zero).

    -   Between 0 and 1 = Elastic Net.

Tuning method: tune_bayes(). Instead of exhaustively testing all combinations on a fixed grid, Bayesian optimization builds a probabilistic model of the performance surface and intelligently focuses evaluations on promising regions. 40 iterations explore the penalty x mixture space far more efficiently than a 7 x 20 = 140-point regular grid.

```{r}
#| label: test1-glmnet
autoplot(results_proteomic_ALS_occurrence_tidymodels$test_1$tune_glmnet, type = "performance") +
  theme_minimal() + 
  ggtitle("Test 1 – glmnet (tune_bayes)")

show_best(results_proteomic_ALS_occurrence_tidymodels$test_1$tune_glmnet, 
          metric = "roc_auc",
          n = 5)|>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

cat("Best glmnet T1 — mixture (alpha):", 
    round(results_proteomic_ALS_occurrence_tidymodels$test_1$best_glmnet$mixture, 3),
    "| penalty (lambda):", 
    round(results_proteomic_ALS_occurrence_tidymodels$test_1$best_glmnet$penalty, 5), "\n")
```

#### Random forest tuning

Random Forest trains 1,000 decision trees independently, each on a random bootstrap sample of the data and using only a random subset of predictors at each split. The final prediction is the average across all 1,000 trees.This double randomization makes Random Forest highly resistant to overfitting.

Three hyperparameters are tuned:

-   mtry: Number of predictors randomly considered at each split. Explored between 10% and 30% of total predictors.

-   min_n: Minimum number of observations in a terminal node. It controls tree depth and complexity.

-   splitrule: Criterion used to choose the best split. "gini" = standard Gini impurity. "extratrees" = thresholds are drawn randomly, adding more randomization, often better in high dimensions. Both are tested and the best is kept.

Tuning method: tune_bayes() for mtry and min_n (continuous parameters). splitrule is discrete so the code loops over both values and selects the best.

```{r}
#| label: test1-rf
cat("Best RF T1 — splitrule:", results_proteomic_ALS_occurrence_tidymodels$test_1$best_rf_t1$splitrule,
    "| mtry:", results_proteomic_ALS_occurrence_tidymodels$test_1$best_rf_t1$best$mtry,
    "| min_n:", results_proteomic_ALS_occurrence_tidymodels$test_1$best_rf_t1$best$min_n, "\n")
```

#### Gradient Boosting (XGBoost) tuning

XGBoost builds trees sequentially: each tree corrects the residual errors of the previous one. This sequential method generally yields higher accuracy than random forest but makes XGBoost more prone to overfitting and sensitive to more hyperparameters.

Tuning was done in two steps:

-   Step 1 — Structural hyperparameters:

    -   tree_depth: Maximum depth of each tree (1–4). Kept shallow to prevent overfitting with small data.

    -   learn_rate (eta): Step size for each correction (0.001–0.1). Smaller = more conservative learning.

    -   mtry (colsample_bytree): Fraction of predictors used per tree.

    -   min_n (min_child_weight): Minimum complexity required to create a new node.

    -   loss_reduction (gamma): Minimum gain required to make an additional split — acts as an extra regularizer.

    -   sample_size (subsample): Fixed at 0.8: 80% of rows sampled per tree. Not tuned (standard value).

-   Step 2 — Number of trees via early stopping (xgb.cv, not available with tidymodels):

    -   Once the best structural hyperparameters are fixed, xgb.cv() runs up to 5,000 boosting rounds but stops automatically (early_stopping_rounds = 50) if the CV AUC has not improved for 50 consecutive rounds. This prevents both underfitting (too few trees) and overfitting (too many trees).

    -   The same fold indices as tidymodels are passed to xgb.cv() to ensure the CV is strictly comparable.

```{r}
#| label: test1-xgboost
autoplot(results_proteomic_ALS_occurrence_tidymodels$test_1$tune_xgb_t1) + 
  theme_minimal() + 
  ggtitle("Test 1 – XGBoost")

show_best(results_proteomic_ALS_occurrence_tidymodels$test_1$tune_xgb_t1, 
          metric = "roc_auc", 
          n = 5) |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  flextable::fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

cat("Best XGB structural params T1:\n"); results_proteomic_ALS_occurrence_tidymodels$test_1$best_xgb_params_t1 |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  flextable::fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

# 
# cat("AUC CV (mean ± SD):", 
#     round(results_proteomic_ALS_occurrence_tidymodels$test_1$xgb_cv_es_t1$evaluation_log$test_auc_mean[results_proteomic_ALS_occurrence_tidymodels$test_1$best_xgb_trees_t1], 3),
#     "±", 
#     round(results_proteomic_ALS_occurrence_tidymodels$test_1$xgb_cv_es_t1$evaluation_log$test_auc_std[results_proteomic_ALS_occurrence_tidymodels$test_1$best_xgb_trees_t1], 3), "\n")
```

```{r fig.height=5}
#| eval: false
#| include: false
# Courbe d'apprentissage
results_proteomic_ALS_occurrence_tidymodels$test_1$xgb_cv_es_t1$evaluation_log |>
  ggplot(aes(x = iter)) +
  geom_line(aes(y = results_proteomic_ALS_occurrence_tidymodels$test_1$xgb_cv_es_t1$evaluation_log$train_auc_mean, color = "Train")) +
  geom_line(aes(y = results_proteomic_ALS_occurrence_tidymodels$test_1$xgb_cv_es_t1$evaluation_log$test_auc_mean,  color = "CV Test")) +
  geom_ribbon(aes(ymin = results_proteomic_ALS_occurrence_tidymodels$test_1$xgb_cv_es_t1$evaluation_log$test_auc_mean - results_proteomic_ALS_occurrence_tidymodels$test_1$xgb_cv_es_t1$evaluation_log$test_auc_std,
                  ymax = results_proteomic_ALS_occurrence_tidymodels$test_1$xgb_cv_es_t1$evaluation_log$test_auc_mean + results_proteomic_ALS_occurrence_tidymodels$test_1$xgb_cv_es_t1$evaluation_log$test_auc_std), alpha = 0.15) +
  geom_vline(xintercept = results_proteomic_ALS_occurrence_tidymodels$test_1$best_xgb_trees_t1, linetype = "dashed", color = "grey40") +
  labs(title = "Learning curve XGBoost – Test 1",
       x = "Nb of trees", y = "AUC", color = "") +
  theme_minimal()
```

#### Linear Support Vector Machine (SMV) tuning

The SVM finds the hyperplane that best separates cases from controls by maximizing the margin between the two groups.

"Linear" means the decision boundary is a linear combination of predictors.

One hyperparameter is tuned:

-   cost: Trade-off between maximizing the margin and penalizing misclassified points.

    -   Low cost = wide margin, more errors tolerated (strong regularization).

    -   High cost = narrow margin, fewer errors tolerated.

Applied on the top-20 filtered recipe because SVM is numerically unstable in p \> n settings.

```{r}
#| label: test1-smv
show_best(results_proteomic_ALS_occurrence_tidymodels$test_1$tune_svm_t1, metric = "roc_auc", n = 5) |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

cat("Best SVM T1 — cost:", round(results_proteomic_ALS_occurrence_tidymodels$test_1$best_svm_t1$cost, 4), "\n")
```

#### Multivariate Adaptive Regression Splines (MARS) tuning

MARS is a non-linear model that automatically discovers breakpoints in predictor-outcome relationships. For example, it can learn "if protein X \> threshold, the effect is strong; otherwise it is negligible". More flexible than linear models but more interpretable than tree ensembles.

Two hyperparameters are tuned:

-   num_terms: Number of terms in the model — controls complexity (5, 10, 20, 50 tested).

-   prod_degree: Degree of interactions: 1 = additive model only, 2 = pairwise interactions allowed.

Applied on the top-20 filtered recipe because MARS is numerically unstable in p \> n settings.

```{r}
#| label: test1-mars
show_best(results_proteomic_ALS_occurrence_tidymodels$test_1$tune_mars_t1, metric = "roc_auc", n = 5) |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
```

#### Comparaison of the best combinaisons Test 1

Once the best hyperparameters are identified for each algorithm, all models are evaluated on the same 10 cross-validation folds. This ensures a fair comparison: same data splits, same metric (AUC).

```{r}
#| label: test1-comparison-best-combinaisons
results_proteomic_ALS_occurrence_tidymodels$test_1$summary_t1 |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

autoplot(results_proteomic_ALS_occurrence_tidymodels$test_1$results_t1, metric = "roc_auc") +
  theme_minimal() +
  geom_text(aes(label = round(mean, 3)), hjust = -0.1, size = 3) +
  labs(title = "Test 1 – Comparaison of the models (AUC, grouped 10-fold)",
       x = "Mean AUC (CV)", y = "")
```

#### Best model Test 1

```{r}
#| label: test1-best-model
cat("Best model Test 1:", results_proteomic_ALS_occurrence_tidymodels$test_1$best_wf_id_t1, "\n")
cat("   Mean AUC:", round(results_proteomic_ALS_occurrence_tidymodels$test_1$summary_t1$mean[1], 3),
    "± SE:", round(results_proteomic_ALS_occurrence_tidymodels$test_1$summary_t1$std_err[1], 3), "\n")
```

#### Interpretation : glmnet (winner) + XGBoost (comparison)

##### Glmnet

Top 89 predictors (by \|coefficient\|) selected by glmnet elasticnet

```{r, fig.height=15}
results_proteomic_ALS_occurrence_tidymodels$test_1$t_glmnet_t1
results_proteomic_ALS_occurrence_tidymodels$test_1$f_glmnet_t1
```

##### XGboost

Top 20 variables by SHAP importance

```{r, fig.height=8, fig.width=8}
results_proteomic_ALS_occurrence_tidymodels$test_1$t_xgboost_t1
```

SHAP beeswarm plot

```{r, fig.height=12, fig.width=12}
results_proteomic_ALS_occurrence_tidymodels$test_1$f_xgb_beeswarm_t1
```

SHAP bar plot

```{r}
results_proteomic_ALS_occurrence_tidymodels$test_1$f_xgb_barplot_t1
```

Dependence plots (top 6)

```{r, fig.height=25, fig.width=8}
results_proteomic_ALS_occurrence_tidymodels$test_1$f_xgb_dependanceplot_t1 |> 
  wrap_plots(ncol = 1, nrow = 6)
```

#### Comparison: glmnet vs XGBoost across both tests

```{r}
cat("COMPARISON: glmnet vs XGBoost COEFFICIENTS/SHAP\n")
cat("  Common top 20 proteins:", length(results_proteomic_ALS_occurrence_tidymodels$test_1$overlap_t1), "\n")
if (length(results_proteomic_ALS_occurrence_tidymodels$test_1$overlap_t1) > 0) {
  cat("  Proteins:", paste(results_proteomic_ALS_occurrence_tidymodels$test_1$overlap_t1, collapse = ", "), "\n")
}
```


### Test 2: Can plasma proteins predict ALS? (proteins + follow-up + covariates: birth year, sex, BMI, smoking) 
#### Penalized Logistic Regression (glmnet) tuning

Standard logistic regression would fail here because p \> n (\~280 predictors, 166 cases). glmnet solves this by adding a penalty term to the loss function that shrinks coefficients toward zero, effectively performing implicit variable selection.

Two hyperparameters are tuned:

-   penalty (lambda): Strength of regularization. Higher lambda = more coefficients forced to zero.

-   mixture (alpha): Type of penalty.

    -   alpha = 1 = Lasso (exact zeros, variable selection).

    -   alpha = 0 = Ridge (all coefficients shrunk but non-zero).

    -   Between 0 and 1 = Elastic Net.

Tuning method: tune_bayes(). Instead of exhaustively testing all combinations on a fixed grid, Bayesian optimization builds a probabilistic model of the performance surface and intelligently focuses evaluations on promising regions. 40 iterations explore the penalty x mixture space far more efficiently than a 7 x 20 = 140-point regular grid.

```{r}
#| label: test2-glmnet
autoplot(results_proteomic_ALS_occurrence_tidymodels$test_2$tune_glmnet, type = "performance") +
  theme_minimal() + 
  ggtitle("Test 2 – glmnet (tune_bayes)")

show_best(results_proteomic_ALS_occurrence_tidymodels$test_2$tune_glmnet, 
          metric = "roc_auc",
          n = 5) |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

cat("Best glmnet T2 — mixture (alpha):", 
    round(results_proteomic_ALS_occurrence_tidymodels$test_2$best_glmnet$mixture, 3),
    "| penalty (lambda):", 
    round(results_proteomic_ALS_occurrence_tidymodels$test_2$best_glmnet$penalty, 5), "\n")
```

#### Random forest tuning

```{r}
#| label: test2-rf
cat("Best RF T2 — splitrule:", results_proteomic_ALS_occurrence_tidymodels$test_2$best_rf_t2$splitrule,
    "| mtry:", results_proteomic_ALS_occurrence_tidymodels$test_2$best_rf_t2$best$mtry,
    "| min_n:", results_proteomic_ALS_occurrence_tidymodels$test_2$best_rf_t2$best$min_n, "\n")
```

#### Gradient Boosting (XGBoost) tuning

XGBoost builds trees sequentially: each tree corrects the residual errors of the previous one. This sequential method generally yields higher accuracy than random forest but makes XGBoost more prone to overfitting and sensitive to more hyperparameters.

Tuning was done in two steps:

-   Step 1 — Structural hyperparameters:

    -   tree_depth: Maximum depth of each tree (1–4). Kept shallow to prevent overfitting with small data.

    -   learn_rate (eta): Step size for each correction (0.001–0.1). Smaller = more conservative learning.

    -   mtry (colsample_bytree): Fraction of predictors used per tree.

    -   min_n (min_child_weight): Minimum complexity required to create a new node.

    -   loss_reduction (gamma): Minimum gain required to make an additional split — acts as an extra regularizer.

    -   sample_size (subsample): Fixed at 0.8: 80% of rows sampled per tree. Not tuned (standard value).

-   Step 2 — Number of trees via early stopping (xgb.cv, not available with tidymodels):

    -   Once the best structural hyperparameters are fixed, xgb.cv() runs up to 5,000 boosting rounds but stops automatically (early_stopping_rounds = 50) if the CV AUC has not improved for 50 consecutive rounds. This prevents both underfitting (too few trees) and overfitting (too many trees).

    -   The same fold indices as tidymodels are passed to xgb.cv() to ensure the CV is strictly comparable.

```{r}
#| label: test2-xgboost
autoplot(results_proteomic_ALS_occurrence_tidymodels$test_2$tune_xgb_t2) + 
  theme_minimal() + 
  ggtitle("Test 2 – XGBoost")

show_best(results_proteomic_ALS_occurrence_tidymodels$test_2$tune_xgb_t2, 
          metric = "roc_auc", 
          n = 5)|>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  flextable::fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

cat("Best XGB structural params T2:\n"); results_proteomic_ALS_occurrence_tidymodels$test_2$best_xgb_params_t2 |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  flextable::fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

# 
# cat("CV AUC (mean ± SD):", 
#     round(results_proteomic_ALS_occurrence_tidymodels$test_2$xgb_cv_es_t2$evaluation_log$test_auc_mean[results_proteomic_ALS_occurrence_tidymodels$test_2$best_xgb_trees_t2], 3),
#     "±", 
#     round(results_proteomic_ALS_occurrence_tidymodels$test_2$xgb_cv_es_t2$evaluation_log$test_auc_std[results_proteomic_ALS_occurrence_tidymodels$test_2$best_xgb_trees_t2], 3), "\n")
```

```{r fig.height=5}
#| eval: false
#| include: false
# Courbe d'apprentissage
results_proteomic_ALS_occurrence_tidymodels$test_2$xgb_cv_es_t2$evaluation_log |>
  ggplot(aes(x = iter)) +
  geom_line(aes(y = results_proteomic_ALS_occurrence_tidymodels$test_2$xgb_cv_es_t2$evaluation_log$train_auc_mean, color = "Train")) +
  geom_line(aes(y = results_proteomic_ALS_occurrence_tidymodels$test_2$xgb_cv_es_t2$evaluation_log$test_auc_mean,  color = "CV Test")) +
  geom_ribbon(aes(ymin = results_proteomic_ALS_occurrence_tidymodels$test_2$xgb_cv_es_t2$evaluation_log$test_auc_mean - results_proteomic_ALS_occurrence_tidymodels$test_2$xgb_cv_es_t2$evaluation_log$test_auc_std,
                  ymax = results_proteomic_ALS_occurrence_tidymodels$test_2$xgb_cv_es_t2$evaluation_log$test_auc_mean + results_proteomic_ALS_occurrence_tidymodels$test_2$xgb_cv_es_t2$evaluation_log$test_auc_std), 
              alpha = 0.15) +
  geom_vline(xintercept = results_proteomic_ALS_occurrence_tidymodels$test_2$best_xgb_trees_t2, linetype = "dashed", color = "grey40") +
  labs(title = "Learning curve XGBoost – Test 2",
       x = "Nb of trees", y = "AUC", color = "") +
  theme_minimal()
```

#### Linear Support Vector Machine (SMV) tuning

The SVM finds the hyperplane that best separates cases from controls by maximizing the margin between the two groups.

"Linear" means the decision boundary is a linear combination of predictors.

One hyperparameter is tuned:

-   cost: Trade-off between maximizing the margin and penalizing misclassified points.

    -   Low cost = wide margin, more errors tolerated (strong regularization).

    -   High cost = narrow margin, fewer errors tolerated.

Applied on the top-20 filtered recipe because SVM is numerically unstable in p \> n settings.

```{r}
#| label: test2-smv
show_best(results_proteomic_ALS_occurrence_tidymodels$test_2$tune_svm_t2, metric = "roc_auc", n = 5) |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

cat("Best SVM T2 — cost:", round(results_proteomic_ALS_occurrence_tidymodels$test_2$best_svm_t2$cost, 4), "\n")
```

#### Multivariate Adaptive Regression Splines (MARS) tuning

MARS is a non-linear model that automatically discovers breakpoints in predictor-outcome relationships. For example, it can learn "if protein X \> threshold, the effect is strong; otherwise it is negligible". More flexible than linear models but more interpretable than tree ensembles.

Two hyperparameters are tuned:

-   num_terms: Number of terms in the model — controls complexity (5, 10, 20, 50 tested).

-   prod_degree: Degree of interactions: 1 = additive model only, 2 = pairwise interactions allowed.

Applied on the top-20 filtered recipe because MARS is numerically unstable in p \> n settings.

```{r}
#| label: test2-mars
show_best(results_proteomic_ALS_occurrence_tidymodels$test_2$tune_mars_t2, metric = "roc_auc", n = 5) |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
```

#### Comparaison of the best combinaisons Test 2

Once the best hyperparameters are identified for each algorithm, all models are evaluated on the same 10 cross-validation folds. This ensures a fair comparison: same data splits, same metric (AUC).

```{r}
#| label: test2-comparison-best-combinaisons
results_proteomic_ALS_occurrence_tidymodels$test_2$summary_t2 |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

autoplot(results_proteomic_ALS_occurrence_tidymodels$test_2$results_t2, metric = "roc_auc") +
  theme_minimal() +
  geom_text(aes(label = round(mean, 3)), hjust = -0.1, size = 3) +
  labs(title = "Test 2 – Model comparaison (AUC, grouped 10-fold)", 
       x = "Mean AUC", y = "")
```

#### Best model Test 2

```{r}
#| label: test2-best-model
cat("Best model Test 2 :", results_proteomic_ALS_occurrence_tidymodels$test_2$best_wf_id_t2, "\n")
cat("   Mean AUC:", round(results_proteomic_ALS_occurrence_tidymodels$test_2$summary_t2$mean[1], 3),
    "± SE:", round(results_proteomic_ALS_occurrence_tidymodels$test_2$summary_t2$std_err[1], 3), "\n")
```

#### Interpreation : glmnet (winner) + XGBoost (comparison)

##### Glmnet

Top 110 predictors (by \|coefficient\|)

```{r, fig.height=18}
#| label: Interpreation:glmnet(winner)+XGBoost(comparison)-1
results_proteomic_ALS_occurrence_tidymodels$test_2$t_glmnet_t2
results_proteomic_ALS_occurrence_tidymodels$test_2$p_glmnet_t2
```

##### XGboost

Top 20 variables by SHAP importance

```{r, fig.height=8, fig.width=8}
#| label: Interpreation:glmnet(winner)+XGBoost(comparison)-2
results_proteomic_ALS_occurrence_tidymodels$test_2$t_xgboost_t2
```

SHAP beeswarm plot

```{r, fig.height=12, fig.width=12}
#| label: Interpreation:glmnet(winner)+XGBoost(comparison)-3
results_proteomic_ALS_occurrence_tidymodels$test_2$f_xgb_beeswarm_t2
```

SHAP bar plot

```{r}
#| label: Interpreation:glmnet(winner)+XGBoost(comparison)-4
results_proteomic_ALS_occurrence_tidymodels$test_2$f_xgb_barplot_t2
```

Dependence plots (top 6)

```{r, fig.height=25, fig.width=8}
#| label: Interpreation:glmnet(winner)+XGBoost(comparison)-5
results_proteomic_ALS_occurrence_tidymodels$test_2$f_xgb_dependanceplot_t2 |> 
  wrap_plots(ncol = 1, nrow = 6)
```

Dependence plots by follow-up duration (top 6)

```{r, fig.height=25, fig.width=8}
#| label: Interpreation:glmnet(winner)+XGBoost(comparison)-6
results_proteomic_ALS_occurrence_tidymodels$test_2$f_xgb_dependanceplot_by_followup_t2 |> 
  wrap_plots(ncol = 1, nrow = 6)
```

Interaction follow_up × protéines

```{r, fig.height=9, fig.width=8}
#| label: Interpreation:glmnet(winner)+XGBoost(comparison)-7
#| column: screen-inset-right
cat("\nSHAP |value| by follow-up tertile:\n")
results_proteomic_ALS_occurrence_tidymodels$test_2$f_fu_shap
```

Trajectory analysis

```{r}
#| label: Interpreation:glmnet(winner)+XGBoost(comparison)-8
cat("\nProtein signal trajectory:\n")
results_proteomic_ALS_occurrence_tidymodels$test_2$shap_fu_trend 
```

#### Comparison: glmnet vs XGBoost across both tests

```{r}
#| label: Interpreation:glmnet(winner)+XGBoost(comparison)-9
cat("COMPARISON: glmnet vs XGBoost COEFFICIENTS/SHAP\n")
cat("  Common top 20 proteins:", length(results_proteomic_ALS_occurrence_tidymodels$test_2$overlap_t2), "\n")
if (length(results_proteomic_ALS_occurrence_tidymodels$test_2$overlap_t2) > 0) {
  cat("  Proteins:", paste(results_proteomic_ALS_occurrence_tidymodels$test_2$overlap_t2, collapse = ", "), "\n")
}
```


### Test 3: interaction with follow-up duration?
#### Comparison of glmnet models 
```{r}
#| label: tidymodel-test-3-1
autoplot(results_proteomic_ALS_occurrence_tidymodels$test_3$tune_glmnet, type = "performance") +
  theme_minimal() + 
  ggtitle("Test 3 – glmnet (tune_bayes)")

show_best(results_proteomic_ALS_occurrence_tidymodels$test_3$tune_glmnet, 
          metric = "roc_auc",
          n = 5) |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

cat("Best glmnet T3 — mixture (alpha):", 
    round(results_proteomic_ALS_occurrence_tidymodels$test_3$best_glmnet$mixture, 3),
    "| penalty (lambda):", 
    round(results_proteomic_ALS_occurrence_tidymodels$test_3$best_glmnet$penalty, 5), "\n")
```

#### Interpretation of the best glmnet model
```{r}
#| label: tidymodel-test-3-2
# Extraction de la valeur numérique moyenne du meilleur AUC
best_auc_value <- show_best(
  results_proteomic_ALS_occurrence_tidymodels$test_3$tune_glmnet_t3, 
  metric = "roc_auc", 
  n = 1)$mean

# Récupération des prédictions hors-échantillon
best_predictions_t3 <- 
  results_proteomic_ALS_occurrence_tidymodels$test_3$tune_glmnet_t3 |> 
  collect_predictions(
    parameters = results_proteomic_ALS_occurrence_tidymodels$test_3$best_glmnet_t3)

# Calcul des points de la courbe ROC
roc_curve_data_t3 <- 
  best_predictions_t3 |>
  roc_curve(truth = als, .pred_case)

# plot
roc_curve_data_t3 |> 
  autoplot() +
  labs(
    title = "AUC curve - Test 3 (Elastic Net with follow-up interactions)",
    subtitle = paste0("CV 10-folds — Mean AUC: ", round(best_auc_value, 3))) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(color = "gray40", size = 10))


rm(best_auc_value, best_predictions_t3, roc_curve_data_t3)
```

#### Extraction of the terms selected by elastic net 
```{r}
#| label: tidymodel-test-3-3
# if(results_proteomic_ALS_occurrence_tidymodels$test_3$glmnet_t3_result$success) {
#   cat("\n--- Test 3 results ---")
#   cat("\nBest hyperparameters:")
#   cat("\n  - Alpha (mixture):", round(results_proteomic_ALS_occurrence_tidymodels$test_3$best_glmnet_t3$mixture, 3))
#   cat("\n  - Lambda (penalty):", round(results_proteomic_ALS_occurrence_tidymodels$test_3$best_glmnet_t3$penalty, 5))
#   cat("\nNumber of selected terms by elastic net:", results_proteomic_ALS_occurrence_tidymodels$test_3$glmnet_t3_result$n_selected, "\n")
#   
#   print(head(results_proteomic_ALS_occurrence_tidymodels$test_3$glmnet_t3_result$coefs_all, 143))
# } else {
#   cat("Erreur lors de l'extraction des coefficients :", results_proteomic_ALS_occurrence_tidymodels$test_3$glmnet_t3_result$error, "\n")
# }


# Top 143 predictors (by |coefficient|)
results_proteomic_ALS_occurrence_tidymodels$test_3$glmnet_t3_result$coefs_protein |> 
  slice_head(n = 143) |>
  as.data.frame() |>
  mutate(Feature = str_remove(Feature, "proteomic_neuro_explo_|proteomic_immun_res_|proteomic_metabolism_"), 
         Feature = str_replace(Feature, "_x_follow_up_no_na_y", " x follow-up")) |>
  select(-abs_coef, -direction, - coefficient_raw) |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  flextable::fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

```



# Proteomic and years to ALS

## Cox models among cases

Event = ALS, time to event = from baseline to ALS diagnostic (cases)


### continuous standardized

```{r, fig.width=15, fig.height=5}
#| label: danish-proteomic-y-to-ALS-1
results_proteomic_ALS_occurrence_y_to_als$main_results_cox$proteomic_sd_ALS_table_cox |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
#| label: danish-proteomic-y-to-ALS-2
wrap_plots(results_proteomic_ALS_occurrence_y_to_als$main_results_cox$proteomic_sd_ALS_base_figure_cox, 
           results_proteomic_ALS_occurrence_y_to_als$main_results_cox$proteomic_sd_ALS_adjusted_figure_cox, 
           ncol = 2)
```

### quartiles

```{r}
#| label: danish-proteomic-y-to-ALS-3
#| column: screen-inset-right
results_proteomic_ALS_occurrence_y_to_als$main_results_cox$proteomic_quart_ALS_table_cox |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r}
#| label: danish-proteomic-y-to-ALS-4
bdd_danish_cases <- bdd_danish |> filter(als == 1)

surv_obj <- Surv(time = bdd_danish_cases$follow_up, event = bdd_danish_cases$als)

vars_labels <- 
  set_names(str_replace(proteomic, 
                        "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_", 
                        ""), 
            proteomic)


#### Base ----
cox_gam_results_base <- 
  fit_and_plot_cox_gam(data = bdd_danish_cases, 
                       vars_labels = vars_labels, 
                       covariates = c("sex", "baseline_age"), 
                       title = "Base model")

#### Adjusted ----
cox_gam_results_adjusted <- 
  fit_and_plot_cox_gam(data = bdd_danish_cases, 
                       vars_labels = vars_labels, 
                       covariates = c("sex", "baseline_age", "smoking_2cat_i", "bmi"), 
                       title = "Adjusted model")

plot_base_cox_gam_danish <- list(
  signif = map(keep(cox_gam_results_base, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(keep(cox_gam_results_base, ~ .x$pval_raw > 0.05), "plot"))

plot_adjusted_cox_gam_danish <- list(
  signif = map(keep(cox_gam_results_adjusted, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(keep(cox_gam_results_adjusted, ~ .x$pval_raw > 0.05), "plot"))

rm(vars_labels, surv_obj, bdd_danish_cases, cox_gam_results_base, cox_gam_results_adjusted)
```

```{r fig.height=20, fig.width=9}
#| label: danish-proteomic-y-to-ALS-5
plots_lists <- list(
  base  = plot_base_cox_gam_danish$signif, 
  adj  = plot_adjusted_cox_gam_danish$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(plots_lists, all_proteins, plot_grid, plot_base_cox_gam_danish, plot_adjusted_cox_gam_danish)

```

## Fixed-horizon logistic analyses

### Proteome wide associations (logistic regressions)

#### Continuous coding (standardized)

```{r}
#| label: danish-proteomic-y-to-ALS-6
results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$proteome_wide$proteomic_sd_ALS_table_y_10_20
```

```{r}
#| label: danish-proteomic-y-to-ALS-7
results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$proteome_wide$proteomic_sd_ALS_adjusted_figure_y_10ans
```

```{r}
#| label: danish-proteomic-y-to-ALS-8
results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$proteome_wide$proteomic_sd_ALS_adjusted_figure_y_20ans
```

#### Quartiles

```{r}
#| label: danish-proteomic-y-to-ALS-9
#| column: screen-inset-right
results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$proteome_wide$proteomic_quart_ALS_table_y_10_20
```

### Machine learning (Xboost)

#### 10 years horizon

Number of optimal trees

```{r}
#| label: danish-proteomic-y-to-ALS-10
cat("---> Optimal number of trees (10 years horizon) :", results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$best_nrounds_10, "\n")
```

Maximum AUC (Cross-validated)

```{r}
#| label: danish-proteomic-y-to-ALS-11
cat("---> Maximum AUC (cross validated, 10 years horizon) :", max(results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$cv_10ans$evaluation_log$test_auc_mean), "\n")
```

Importance top 20 variables (10 years horizon)

```{r}
#| label: danish-proteomic-y-to-ALS-12
cat("\nTop 20 predictors - 10 years horizon :\n")
head(results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$importance_10, 20) |>
    flextable() |>
  merge_h(part = "header") |>
  theme_vanilla() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |> 
  set_table_properties(align = "left")
```

```{r}
#| label: danish-proteomic-y-to-ALS-13
xgb.plot.importance(
  importance_matrix = results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$importance_10[1:20, ], 
  main = "Importance of the top 20 predictors (10 years horizon)")
```

#### 20 years horizon

Number of optimal trees

```{r}
#| label: danish-proteomic-y-to-ALS-14
cat("---> Optimal number of trees (20 years horizon) :", results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$best_nrounds_20, "\n")
```

Maximum AUC (Cross-validated)

```{r}
#| label: danish-proteomic-y-to-ALS-15
cat("---> Maximum AUC (cross validated, 20 years horizon) :", max(results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$cv_20ans$evaluation_log$test_auc_mean), "\n")
```

Importance top 20 variables (20 year horizon)

```{r}
#| label: danish-proteomic-y-to-ALS-16
cat("\nTop 20 predictors - 20 years horizon :\n")
head(results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$importance_20, 20) |>
    flextable() |>
  merge_h(part = "header") |>
  theme_vanilla() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all")|> 
  set_table_properties(align = "left")
```

```{r}
#| label: danish-proteomic-y-to-ALS-17
xgb.plot.importance(
  importance_matrix = results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$importance_20[1:20, ], 
  main = "Importance of the top 20 predictors (20 years horizon)")
```

## Logistic models including time

### Proteome wide associations

```{r}
#| label: danish-proteomic-y-to-ALS-18
results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$proteome_wide$proteomic_interaction_table
```

### Machine learning (XGboost)

#### Etape 1 : choisir les meilleurs hyperparameters avec super learner (excepté nb de trees) 
```{r}
#| label: danish-proteomic-y-to-ALS-19
summary(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_1_AUC)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
plot(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_1_AUC) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1 to choose the best hyperparameters except nb of trees)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")

summary(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_1_NNLS)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
plot(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_1_NNLS)  + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1 to choose the best hyperparameters except nb of trees)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "V-fold CV Risk Estimate (NNLS)") +
  geom_hline(yintercept = 0.222, linetype = "dashed", color = "red")
```

$$MSE_{hasard} = p(1-p) = \frac{1}{3} \times \frac{2}{3} = \frac{2}{9} \approx 0.222$$


#### Etape 2 : choisir le nb de trees avec SL (grace a l'argument early_stopping_rounds ) et les meilleurs hyperparameters choisis en 1
```{r}
#| label: danish-proteomic-y-to-ALS-20
summary(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_2_AUC)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
plot(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_2_AUC) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 2 to choose the best nb of trees)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")

summary(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_2_NNLS)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
plot(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_2_NNLS) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 2 to choose the best nb of trees)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "V-fold CV Risk Estimate (NNLS)") +
  geom_hline(yintercept = 0.222, linetype = "dashed", color = "red")
```

#### Etapes 1 et 2 mixées ensembles
```{r}
#| label: danish-proteomic-y-to-ALS-21
summary(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_1_2_AUC)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
plot(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_1_2_AUC) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1 and 2 mixed to choose the best hyperparameters and nb of trees)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")

summary(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_1_2_NNLS)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
plot(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_1_2_NNLS)  + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1 and 2 mixed to choose the best hyperparameters and nb of trees)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "V-fold CV Risk Estimate (NNLS)") +
  geom_hline(yintercept = 0.222, linetype = "dashed", color = "red")
```

## Linear models 
### Proteome wide associations 
```{r}
#| label: danish-proteomic-y-to-ALS-22
results_proteomic_ALS_occurrence_y_to_als$main_results_linear$proteome_wide$proteomic_sd_ALS_table_linear 
results_proteomic_ALS_occurrence_y_to_als$main_results_linear$proteome_wide$proteomic_sd_ALS_base_figure_linear
results_proteomic_ALS_occurrence_y_to_als$main_results_linear$proteome_wide$proteomic_sd_ALS_adjusted_figure_linear 
```

### Machine learning (tidymodels)
```{r}
#| label: danish-proteomic-y-to-ALS-23
results_proteomic_ALS_occurrence_y_to_als$main_results_linear$machine_learning$t_glmnet_reg

results_proteomic_ALS_occurrence_y_to_als$main_results_linear$machine_learning$f_glmnet_reg 
```

# Proteomic and ALS survival

## Effects of subject characteristics on ALS survival

```{r}
#| label: danish-proteomic-ALS-survival-1
results_proteomic_ALS_survival$main$covar |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

## Main analyses

### continuous standardized

```{r, fig.width=15, fig.height=5}
#| label: danish-proteomic-ALS-survival-2
results_proteomic_ALS_survival$main$proteomic_sd_ALS_table |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-survival-3
#| column: screen-inset-right

wrap_plots(results_proteomic_ALS_survival$main$proteomic_sd_ALS_base_figure, 
           results_proteomic_ALS_survival$main$proteomic_sd_ALS_adjusted_figure, 
           ncol = 2)
```

### quartiles

```{r}
#| label: danish-proteomic-ALS-survival-4
#| column: screen-inset-right
results_proteomic_ALS_survival$main$proteomic_quart_ALS_table  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=55, fig.width=9}
#| label: danish-proteomic-ALS-survival-5
fit_and_plot_cox_gam <- function(
    data,
    vars = proteomic,
    time = "follow_up_death",
    status = "status_death", 
    covariates = c("sex", "diagnosis_age"),
    vars_labels = NULL,
    y_limits = c(-12, 12),
    title = "Base model"
) {
  
  if (is.null(vars_labels)) {
    vars_labels <- set_names(vars, vars)
  }
  
  fit_one <- function(var) {
    
    outcome <- cbind(data[[time]], data[[status]])
    formula_str <- paste0(
      "outcome ~ s(", var, ") + ",
      paste(covariates, collapse = " + "))
    
    model <- mgcv::gam(as.formula(formula_str),
                       data = data,
                       family = cox.ph())
  
    smry <- summary(model)
    edf_val <- smry$s.table[1, "edf"]
    pval_raw <- smry$s.table[1, "p-value"]
    
    edf_label <- format(edf_val, nsmall = 1, digits = 1)
    pval_label <- case_when(
      pval_raw < 0.01 ~ "< 0.01",
      pval_raw > 0.99 ~ "> 0.99",
      TRUE ~ format(pval_raw, nsmall = 2, digits = 2))
    
    pdf(NULL)
    plot_data_raw <- plot(model, select = 1, seWithMean = TRUE, rug = FALSE)
    dev.off()
    
    if (length(plot_data_raw) == 0) return(NULL)
    
    plot_data <- plot_data_raw[[1]]
    smooth_df <- data.frame(
      x = plot_data$x,
      fit = plot_data$fit,
      se = plot_data$se)
    
    p1 <- ggplot(smooth_df, aes(x = x, y = fit)) +
      geom_line(linewidth = 1.2, color = "steelblue") +
      geom_ribbon(aes(ymin = fit - 2 * se, 
                      ymax = fit + 2 * se),
                  alpha = 0.2, fill = "steelblue") +
      labs(
        title = title,
        y = "LogHR (smooth estimate)",
        x = NULL) +
      annotate(
        "text",
        x = -Inf, y = Inf,
        hjust = -0.1, vjust = 1.5,
        label = paste0("EDF: ", edf_label, "\np-value: ", pval_label),
        size = 4.2) +
      scale_y_continuous(limits = y_limits) +
      theme_minimal() +
      theme(
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank())
    
    p2 <- ggplot(data, aes(x = .data[[var]], y = 1)) +
      geom_boxplot(width = 0.3, fill = "steelblue", color = "black") +
      labs(x = vars_labels[[var]]) +
      theme_minimal() +
      theme(
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())
  
    list(
      model = model,
      pval_raw = pval_raw,
      plot = p1 / p2 + plot_layout(heights = c(10, 1)))
  }
  
  results <- map(vars, fit_one)
  names(results) <- vars_labels[vars]
  
  return(results)
}

# Creation of cases specific datasets ----
bdd_cases_danish <- 
  bdd_danish |>
  filter(als == 1) |>                                                           # remove controls 
  mutate(                                                                       # remove NEFL outlier
    proteomic_neuro_explo_NEFL = ifelse(match == 159, NA, proteomic_neuro_explo_NEFL)) |>   
  mutate(across(all_of(proteomic),                                              # creation of proteomic standardized variables (cohort and cases specific)  
                ~as.numeric(scale(.x)),
                .names = "{.col}_sd")) 

surv_obj <- Surv(time = bdd_cases_danish$follow_up_death,                       # set the outcomes
                 event = bdd_cases_danish$status_death)

var_label(bdd_cases_danish[proteomic]) <- NULL

vars_labels <- 
  set_names(str_replace(proteomic, 
                        "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_", 
                        ""), 
            proteomic)


### Base ----
cox_gam_results_base <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish, 
                       vars_labels = vars_labels)

### Adjusted ----
cox_gam_results_adjusted <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish, 
                       vars_labels = vars_labels, 
                       covariates = c("sex", "diagnosis_age", "smoking_2cat_i", "bmi"), 
                       title = "Adjusted model")

plot_base_cox_gam_danish <- list(
  signif = map(purrr::keep(cox_gam_results_base, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_base, ~ .x$pval_raw > 0.05), "plot"))
plot_adjusted_cox_gam_danish <- list(
  signif = map(purrr::keep(cox_gam_results_adjusted, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_adjusted, ~ .x$pval_raw > 0.05), "plot"))


plots_lists <- list(
  main_base  = plot_base_cox_gam_danish$signif, 
  main_adj  = plot_adjusted_cox_gam_danish$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(bdd_cases_danish, surv_obj, vars_labels,  
   cox_gam_results_base, cox_gam_results_adjusted,  plots_lists, all_proteins, plot_grid, 
   plot_base_cox_gam_danish)

```

## Sensi 1 + 3 : Removing NEFL outlier + filtering cases with follow-up \> 5 years

### continuous standardized

```{r, fig.width=15, fig.height=5}
#| label: danish-proteomic-ALS-survival-6
results_proteomic_ALS_survival$sensi_1_3$proteomic_sd_ALS_table_sensi_1_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-survival-7
#| column: screen-inset-right
wrap_plots(results_proteomic_ALS_survival$sensi_1_3$proteomic_sd_ALS_base_figure_sensi_1_3, 
           results_proteomic_ALS_survival$sensi_1_3$proteomic_sd_ALS_adjusted_figure_sensi_1_3, 
           ncol = 2)

```

### quartiles

```{r}
#| label: danish-proteomic-ALS-survival-8
#| column: screen-inset-right
results_proteomic_ALS_survival$sensi_1_3$proteomic_quart_ALS_table_sensi_1_3  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=60, fig.width=9}
#| label: danish-proteomic-ALS-survival-9
bdd_cases_danish_sensi_1_3 <- 
  bdd_danish |>
  filter (als == 1 & follow_up > 60) |>                                         # filtering cases with follow_up > 5 years
  mutate(                                                                       # remove NEFL outlier
    proteomic_neuro_explo_NEFL = ifelse(match == 159, NA, proteomic_neuro_explo_NEFL)) 

surv_obj_sensi_1_3 <- Surv(time = bdd_cases_danish_sensi_1_3$follow_up_death,   # set the outcomes
                           event = bdd_cases_danish_sensi_1_3$status_death)

vars_labels <- set_names(str_replace(proteomic, 
                                     "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_", 
                                     ""), 
                         proteomic)


### Base  ----
cox_gam_results_base_sensi_1_3 <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish_sensi_1_3, 
                       vars_labels = vars_labels)


### Adjusted  ----
cox_gam_results_adjusted_sensi_1_3 <- 
    fit_and_plot_cox_gam(data = bdd_cases_danish_sensi_1_3, 
                         vars_labels = vars_labels, 
                         covariates = c("sex", "diagnosis_age", "smoking_2cat_i", "bmi"), 
                         title = "Adjusted model")


plot_base_cox_gam_danish_sensi_1_3 <- list(
  signif = map(purrr::keep(cox_gam_results_base_sensi_1_3, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_base_sensi_1_3, ~ .x$pval_raw > 0.05), "plot"))
plot_adjusted_cox_gam_danish_sensi_1_3 <- list(
  signif = map(purrr::keep(cox_gam_results_adjusted_sensi_1_3, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_adjusted_sensi_1_3, ~ .x$pval_raw > 0.05), "plot"))


plots_lists <- list(
  main = plot_adjusted_cox_gam_danish$signif,
  sensi_1_3 = plot_adjusted_cox_gam_danish_sensi_1_3$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(bdd_cases_danish_sensi_1_3, surv_obj_sensi_1_3, vars_labels,  
   cox_gam_results_base_sensi_1_3, cox_gam_results_adjusted_sensi_1_3, 
   plot_base_cox_gam_danish_sensi_1_3, plot_adjusted_cox_gam_danish_sensi_1_3, 
   plots_lists, all_proteins, plot_grid)
```

## Sensi 1 + 3 + 4 : Removing NEFL outlier + filtering cases with follow-up \> 5 years + filtering follow-up \< 50%

### continuous standardized

```{r, fig.width=9, fig.height=3}
#| label: danish-proteomic-ALS-survival-10
#| column: screen-inset-right

results_proteomic_ALS_survival$sensi_1_3_4$proteomic_sd_ALS_table_sensi_1_3_4 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-survival-11
#| column: screen-inset-right
wrap_plots(results_proteomic_ALS_survival$sensi_1_3_4$proteomic_sd_ALS_base_figure_sensi_1_3_4,
           results_proteomic_ALS_survival$sensi_1_3_4$proteomic_sd_ALS_adjusted_figure_sensi_1_3_4, 
           ncol = 2)

```

### quartiles

```{r}
#| label: danish-proteomic-ALS-survival-12
#| column: screen-inset-right
results_proteomic_ALS_survival$sensi_1_3_4$proteomic_quart_ALS_table_sensi_1_3_4  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=100, fig.width=9}
#| label: danish-proteomic-ALS-survival-13
bdd_cases_danish_sensi_1_3_4 <- bdd_danish |>
  filter(als == 1 & follow_up > 60) |>                                          # remove controls and remove cases with follow-up<5years 
  mutate(                                                                       # remove NEFL outlier
    proteomic_neuro_explo_NEFL = ifelse(match == 159, NA, proteomic_neuro_explo_NEFL)) |>   
  mutate(seuil = quantile(follow_up, 0.5, na.rm = TRUE)) |>                     # sensi 4 : we remove 50% of the highest values of follow-up (after removing the follow-up<5years)
  filter(follow_up <= seuil) 

surv_obj_sensi_1_3_4 <- Surv(time = bdd_cases_danish_sensi_1_3_4$follow_up_death,           # set the outcomes
                             event = bdd_cases_danish_sensi_1_3_4$status_death)


vars_labels <- set_names(str_replace(proteomic, 
                                     "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_", 
                                     ""), 
                         proteomic)


### Base  ----
cox_gam_results_base_sensi_1_3_4 <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish_sensi_1_3_4, 
                       vars_labels = vars_labels)


### Adjusted  ----
cox_gam_results_adjusted_sensi_1_3_4 <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish_sensi_1_3_4, 
                       vars_labels = vars_labels, 
                       covariates = c("sex", "diagnosis_age", "smoking_2cat_i", "bmi"), 
                       title = "Adjusted model")

plot_base_cox_gam_danish_sensi_1_3_4 <- list(
  signif = map(purrr::keep(cox_gam_results_base_sensi_1_3_4, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_base_sensi_1_3_4, ~ .x$pval_raw > 0.05), "plot"))
plot_adjusted_cox_gam_danish_sensi_1_3_4 <- list(
  signif = map(purrr::keep(cox_gam_results_adjusted_sensi_1_3_4, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_adjusted_sensi_1_3_4, ~ .x$pval_raw > 0.05), "plot"))


plots_lists <- list(
  main = plot_adjusted_cox_gam_danish$signif,
  sensi_1_3 = plot_adjusted_cox_gam_danish_sensi_1_3_4$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(bdd_cases_danish_sensi_1_3_4, surv_obj_sensi_1_3_4, vars_labels, 
   cox_gam_results_base_sensi_1_3_4, cox_gam_results_adjusted_sensi_1_3_4, 
   plot_base_cox_gam_danish_sensi_1_3_4, plot_adjusted_cox_gam_danish_sensi_1_3_4, 
   plots_lists, all_proteins, plot_grid)
```

## Sensi 1 + 3 + 5 : Removing NEFL outlier + filtering cases with follow-up \> 5 years + filtering follow-up \> 50%

### continuous standardized

```{r, fig.width=9, fig.height=3}
#| label: danish-proteomic-ALS-survival-14
#| column: screen-inset-right

results_proteomic_ALS_survival$sensi_1_3_5$proteomic_sd_ALS_table_sensi_1_3_5 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-survival-15
#| column: screen-inset-right
wrap_plots(results_proteomic_ALS_survival$sensi_1_3_5$proteomic_sd_ALS_base_figure_sensi_1_3_5,
           results_proteomic_ALS_survival$sensi_1_3_5$proteomic_sd_ALS_adjusted_figure_sensi_1_3_5, 
           ncol = 2)

```

### quartiles

```{r}
#| label: danish-proteomic-ALS-survival-16
#| column: screen-inset-right
results_proteomic_ALS_survival$sensi_1_3_5$proteomic_quart_ALS_table_sensi_1_3_5  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=80, fig.width=9}
#| label: danish-proteomic-ALS-survival-17
bdd_cases_danish_sensi_1_3_5 <- bdd_danish |>
  filter(als == 1 & follow_up > 60) |>                                          # remove controls and remove cases with follow-up<5years 
  mutate(                                                                       # remove NEFL outlier
    proteomic_neuro_explo_NEFL = ifelse(match == 159, NA, proteomic_neuro_explo_NEFL)) |>   
  mutate(seuil = quantile(follow_up, 0.5, na.rm = TRUE)) |>                     # sensi 4 : we remove 50% of the highest values of follow-up (after removing the follow-up<5years)
  filter(follow_up > seuil) 

surv_obj_sensi_1_3_5 <- Surv(time = bdd_cases_danish_sensi_1_3_5$follow_up_death,           # set the outcomes
                             event = bdd_cases_danish_sensi_1_3_5$status_death)


vars_labels <- set_names(str_replace(proteomic, 
                                     "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_", 
                                     ""), 
                         proteomic)


### Base  ----
cox_gam_results_base_sensi_1_3_5 <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish_sensi_1_3_5, 
                       vars_labels = vars_labels)


### Adjusted  ----
cox_gam_results_adjusted_sensi_1_3_5 <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish_sensi_1_3_5, 
                       vars_labels = vars_labels, 
                       covariates = c("sex", "diagnosis_age", "smoking_2cat_i", "bmi"), 
                       title = "Adjusted model")

plot_base_cox_gam_danish_sensi_1_3_5 <- list(
  signif = map(purrr::keep(cox_gam_results_base_sensi_1_3_5, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_base_sensi_1_3_5, ~ .x$pval_raw > 0.05), "plot"))
plot_adjusted_cox_gam_danish_sensi_1_3_5 <- list(
  signif = map(purrr::keep(cox_gam_results_adjusted_sensi_1_3_5, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_adjusted_sensi_1_3_5, ~ .x$pval_raw > 0.05), "plot"))

plots_lists <- list(
  main = plot_adjusted_cox_gam_danish$signif,
  sensi_1_3_5 = plot_adjusted_cox_gam_danish_sensi_1_3_5$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(bdd_cases_danish_sensi_1_3_5, surv_obj_sensi_1_3_5, vars_labels, 
   cox_gam_results_base_sensi_1_3_5, cox_gam_results_adjusted_sensi_1_3_5, 
   plot_base_cox_gam_danish_sensi_1_3_5, plot_adjusted_cox_gam_danish_sensi_1_3_5, 
   plots_lists, all_proteins, plot_grid)

```

## Sensi 1 + 7 females : Removing NEFL outlier + filtering female cases

### continuous standardized

```{r, fig.width=9, fig.height=3}
#| label: danish-proteomic-ALS-survival-18
#| column: screen-inset-right

results_proteomic_ALS_survival$sensi_1_7_female$proteomic_sd_ALS_table_sensi_1_7_female |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-survival-19
#| column: screen-inset-right
wrap_plots(results_proteomic_ALS_survival$sensi_1_7_female$proteomic_sd_ALS_base_figure_sensi_1_7_female,
           results_proteomic_ALS_survival$sensi_1_7_female$proteomic_sd_ALS_adjusted_figure_sensi_1_7_female, 
           ncol = 2)

```

### quartiles

```{r}
#| label: danish-proteomic-ALS-survival-20
#| column: screen-inset-right
results_proteomic_ALS_survival$sensi_1_7_female$proteomic_quart_ALS_table_sensi_1_7_female  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=80, fig.width=9}
#| label: danish-proteomic-ALS-survival-21
bdd_cases_danish_sensi_1_7_female <- 
  bdd_danish |>
  filter(als == 1) |>                                                           # remove controls
  mutate(                                                                       # remove NEFL outlier
    proteomic_neuro_explo_NEFL = ifelse(match == 159, NA, proteomic_neuro_explo_NEFL)) |>   
  filter(sex == "Female")                                                    # sensi 7 female : we purrr::keep only female cases
  

surv_obj_sensi_1_7_female <- Surv(time = bdd_cases_danish_sensi_1_7_female$follow_up_death,           # set the outcomes
                                  event = bdd_cases_danish_sensi_1_7_female$status_death)


vars_labels <- set_names(str_replace(proteomic,
                                     "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
                                     ""),
                         proteomic)


### Base  ----
cox_gam_results_base_sensi_1_7_female <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish_sensi_1_7_female, 
                       vars_labels = vars_labels, 
                       covariates = "diagnosis_age")


### Adjusted  ----
cox_gam_results_adjusted_sensi_1_7_female <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish_sensi_1_7_female, 
                       vars_labels = vars_labels, 
                       covariates = c("diagnosis_age", "smoking_2cat_i", "bmi"), 
                       title = "Adjusted model")

plot_base_cox_gam_danish_sensi_1_7_female <- list(
  signif = map(purrr::keep(cox_gam_results_base_sensi_1_7_female, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_base_sensi_1_7_female, ~ .x$pval_raw > 0.05), "plot"))
plot_adjusted_cox_gam_danish_sensi_1_7_female <- list(
  signif = map(purrr::keep(cox_gam_results_adjusted_sensi_1_7_female, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_adjusted_sensi_1_7_female, ~ .x$pval_raw > 0.05), "plot"))


plots_lists <- list(
  main = plot_adjusted_cox_gam_danish$signif,
  sensi_1_7_female = plot_adjusted_cox_gam_danish_sensi_1_7_female$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(bdd_cases_danish_sensi_1_7_female, surv_obj_sensi_1_7_female, vars_labels, 
   cox_gam_results_base_sensi_1_7_female, cox_gam_results_adjusted_sensi_1_7_female, 
   plot_base_cox_gam_danish_sensi_1_7_female, plot_adjusted_cox_gam_danish_sensi_1_7_female, 
   plots_lists, all_proteins, plot_grid, vars_labels)
```

## Sensi 1 + 7 males : Removing NEFL outlier + male cases

### continuous standardized

```{r, fig.width=9, fig.height=3}
#| label: danish-proteomic-ALS-survival-22
#| column: screen-inset-right

results_proteomic_ALS_survival$sensi_1_7_male$proteomic_sd_ALS_table_sensi_1_7_male |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| label: danish-proteomic-ALS-survival-23
#| column: screen-inset-right
wrap_plots(results_proteomic_ALS_survival$sensi_1_7_male$proteomic_sd_ALS_base_figure_sensi_1_7_male,
           results_proteomic_ALS_survival$sensi_1_7_male$proteomic_sd_ALS_adjusted_figure_sensi_1_7_male, 
           ncol = 2)

```

### quartiles

```{r}
#| label: danish-proteomic-ALS-survival-24
#| column: screen-inset-right

results_proteomic_ALS_survival$sensi_1_7_male$proteomic_quart_ALS_table_sensi_1_7_male  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=80, fig.width=9}
#| label: danish-proteomic-ALS-survival-25
bdd_cases_danish_sensi_1_7_male <- bdd_danish |>
  filter(als == 1) |>                                                           # remove controls
  mutate(                                                                       # remove NEFL outlier
    proteomic_neuro_explo_NEFL = ifelse(match == 159, NA, proteomic_neuro_explo_NEFL)) |>   
  filter(sex == "Male")                                                         # sensi 7 male : we purrr::keep only male cases


surv_obj_sensi_1_7_male <- Surv(time = bdd_cases_danish_sensi_1_7_male$follow_up_death,           # set the outcomes
                                event = bdd_cases_danish_sensi_1_7_male$status_death)

vars_labels <- set_names(str_replace(proteomic,
                                     "proteomic_immun_res_|proteomic_neuro_explo_|proteomic_metabolism_",
                                     ""),
                         proteomic)


### Base  ----
cox_gam_results_base_sensi_1_7_male <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish_sensi_1_7_male, 
                       vars_labels = vars_labels, 
                       covariates = "diagnosis_age")


### Adjusted  ----
cox_gam_results_adjusted_sensi_1_7_male <- 
  fit_and_plot_cox_gam(data = bdd_cases_danish_sensi_1_7_male, 
                       vars_labels = vars_labels, 
                       covariates = c("diagnosis_age", "smoking_2cat_i", "bmi"), 
                       title = "Adjusted model")

plot_base_cox_gam_danish_sensi_1_7_male <- list(
  signif = map(purrr::keep(cox_gam_results_base_sensi_1_7_male, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_base_sensi_1_7_male, ~ .x$pval_raw > 0.05), "plot"))
plot_adjusted_cox_gam_danish_sensi_1_7_male <- list(
  signif = map(purrr::keep(cox_gam_results_adjusted_sensi_1_7_male, ~ .x$pval_raw <= 0.05), "plot"),
  not_signif = map(purrr::keep(cox_gam_results_adjusted_sensi_1_7_male, ~ .x$pval_raw > 0.05), "plot"))

plots_lists <- list(
  main = plot_adjusted_cox_gam_danish$signif,
  sensi_1_7_male = plot_adjusted_cox_gam_danish_sensi_1_7_male$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(plot_base_cox_gam_danish, plot_adjusted_cox_gam_danish, 
   bdd_cases_danish_sensi_1_7_male, surv_obj_sensi_1_7_male, vars_labels, 
   cox_gam_results_base_sensi_1_7_male, cox_gam_results_adjusted_sensi_1_7_male, 
   plot_base_cox_gam_danish_sensi_1_7_male, plot_adjusted_cox_gam_danish_sensi_1_7_male, 
   plots_lists, all_proteins, plot_grid, 
   plot_adjusted_cox_gam_danish)

```


# POPs and proteomic 
```{r, fig.height=6}
#| label: danish-POPs-proteomic-1
#| column: screen-inset-right
results_POPs_proteomic$main$POPs_sd_proteomic_table
results_POPs_proteomic$main$POPs_sd_proteomic_figure
```


```{r}
#| label: danish-POPs-proteomic-2
results_POPs_proteomic$main$POPs_sd_NfL_table
```

