---
title: "Aline's work"
format:
  html:
    embed-resources: true
editor: visual
toc: true
toc_depth: 4
theme:
  light: flatly
  dark: darkly
toc-location: left
execute:
  echo: false
  message: false
  warning: false
---

```{r}
#| message: false
#| warning: false
#| include: false
source("~/Documents/POP_ALS_2025_02_03/1_codes/2.7_analyses_proteomic_ALS_survival.R") |>
  suppressMessages() |> 
  suppressWarnings()

results_descriptive <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/results_descriptive.rds")
results_POPs_ALS_occurrence <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/results_POPs_ALS_occurrence.rds")
results_POPs_ALS_survival <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/results_POPs_ALS_survival.rds")
results_fattyacids_ALS_occurrence <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/results_fattyacids_ALS_occurrence.rds")
results_fattyacids_ALS_survival <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/results_fattyacids_ALS_survival.rds")
results_proteomic_ALS_occurrence_SL <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/results_proteomic_ALS_occurrence_SL.rds")
results_proteomic_ALS_occurrence_y_to_als <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/results_proteomic_ALS_occurrence_y_to_als.rds") 
# results_proteomic_ALS_occurrence <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/results_proteomic_ALS_occurrence.rds")
# results_proteomic_ALS_survival <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/results_proteomic_ALS_survival.rds")
```

# Danish data description

## Metadata

Not imputed for missing data

```{r}
results_descriptive$danish$covar_danish |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Multiple cholesterol variables

```{r, fig.height=2}
boxplot(data = bdd_danish, vars = c("fS_Kol", "cholesterol"))
descrip_num(data = bdd_danish, vars = c("fS_Kol", "cholesterol")) |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
cor.test(bdd_danish$fS_Kol, bdd_danish$cholesterol)
```

### ALS survival

```{r}
bdd_danish |> 
  filter(als == 1) |>
  ggplot() +
  aes(x = follow_up_death) +
  geom_density(fill = "lightgray") +
  theme_bw() +
  stat_overlay_normal_density(color = "red", linetype = "dashed") +
  theme(axis.title = element_blank())
descrip_num(data = bdd_danish, vars = "follow_up_death") |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Timing between baseline, ALS diagnosis and death

```{r}
timing <- bdd |> 
  filter(study == "Danish") |>
  filter(als == 1) |>
  select(baseline_age, diagnosis_age, death_age, 
         time_baseline_diagnosis, time_baseline_death, time_diagnosis_death)

timing |> 
  tbl_summary() |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

descrip_num(data = timing, vars = colnames(timing))   |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
rm(timing)
```

### Other neuorlogical diagnoses

```{r}
results_descriptive$danish$table_other_diag_1
results_descriptive$danish$figure_other_diag
results_descriptive$danish$table_other_diag_2
results_descriptive$danish$table_other_diag_3
```

## POPs

### Distribution tables

```{r}
results_descriptive$danish$POPs_table_danish |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Comparison of distribution cases/controls

```{r}
results_descriptive$danish$POPs_table_danish_by_als|>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

```{r fig.height=11, fig.width=11}
corrplot(results_descriptive$danish$POPs_heatmap_danish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))
```

### Boxplots

```{r, fig.width=10, fig.height=10}
results_descriptive$danish$POPs_boxplot_danish_by_als
```

```{r, fig.height=3}
results_descriptive$danish$POPs_group_boxplot_danish_by_als
```

POPs distribution depending on alive/deceased status at the end of the follow-up (among cases)

```{r, fig.height=3}
results_descriptive$danish$POPs_group_boxplot_danish_by_death
```

### Detailed distribution

#### Summed

```{r}
densityplot(data = bdd_danish, vars = POPs_group)
boxplot(data = bdd_danish, vars = POPs_group)
heatmap_cor(cormat = bdd_danish[POPs_group], decimal = 1)
```

#### PCBs

```{r}
PCBs <- bdd_danish |> select(all_of(POPs)) |> select(contains("PCB")) |> colnames()
densityplot(data = bdd_danish, vars = PCBs[1:9])
```

```{r, fig.height=3.33}
densityplot(data = bdd_danish, vars = PCBs[10:13])
```

```{r}
boxplot(data = bdd_danish, vars = PCBs[1:10])
```

```{r, fig.height=2}
boxplot(data = bdd_danish, vars = PCBs[11:13])
```

```{r}
heatmap_cor(cormat = bdd_danish[PCBs], decimal = 1)
rm(PCBs)
```

#### OCPs

```{r}
OCPs <- bdd_danish |> select(all_of(POPs)) |> select(contains("OCP")) |> colnames()
densityplot(data = bdd_danish, vars = OCPs)
boxplot(data = bdd_danish, vars = OCPs)
heatmap_cor(cormat = bdd_danish[OCPs], decimal = 1)
```

#### PBDEs

```{r,  fig.width=7, fig.height=2}
PBDEs <- bdd_danish |> select(all_of(POPs)) |> select(contains("PBDE")) |> colnames()
densityplot(data = bdd_danish, vars = PBDEs)
boxplot(data = bdd_danish, vars = PBDEs)
```

```{r,  fig.width=7, fig.height=3}
heatmap_cor(cormat = bdd_danish[PBDEs], decimal = 1)
```

## Fatty acids

### Distribution tables

```{r}
results_descriptive$danish$fattyacids_table_danish|>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
```

### Comparison of distribution cases/controls

```{r}
results_descriptive$danish$fattyacids_table_danish_by_als |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

```{r, fig.height=12, fig.width=12}
corrplot(results_descriptive$danish$fattyacids_heatmap_danish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))
```

### Boxplots

```{r, fig.height=10, fig.width=10}
results_descriptive$danish$fattyacids_boxplot_danish_by_als
```

### Detailed distribution

#### Density plots

```{r, fig.height=10, fig.width=10}
bdd_danish |> 
  select(sample, all_of(fattyacids_tot)) |>
  pivot_longer(cols = -sample, names_to = "fattyacids", values_to = "values") |>
  mutate(fattyacids = fct_recode(fattyacids, !!!fattyacids_tot_labels)) |>
  ggplot() +
  aes(x = values) +
  geom_density(fill = "#112446") +
  theme_minimal() +
  facet_wrap(~fattyacids, scales = "free", ncol = 4)
```

## Heatmap POPs/Fatty acids

```{r, fig.width=12, fig.height=12}
results_descriptive$danish$POPs_fattyacids_heatmap_danish
```

## Proteomic

### Distribution table

```{r}
results_descriptive$danish$proteomic_table_danish  |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Comparison of distribution cases/controls

```{r}
#| eval: false
#| message: false
#| warning: false
#| include: false
results_descriptive$danish$proteomic_table_danish_by_als|>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

All

```{r fig.height=11, fig.width=11}
#| eval: false
#| message: false
#| warning: false
#| include: false
corrplot(results_descriptive$danish$proteomic_heatmap_danish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black',
         order = "hclust",         # regroupement des variables selon niveau de cor
         tl.srt = 45, 
         # addCoef.col = "black",
         # number.cex = 0.8,
         # number.digits = 1,
         tl.cex = 0.4,             # taille police des variables
         col = rev(COL2(diverging = "RdYlBu")))
```

```{r fig.height=11, fig.width=11}

vars <- colnames(results_descriptive$danish$proteomic_heatmap_danish)

col_labels <- ifelse(grepl("proteomic_metabolism_", vars), "#1b9e77",  # vert
                     ifelse(grepl("proteomic_immun_res_", vars), "#d95f02",    # orange
                            ifelse(grepl("proteomic_neuro_explo_", vars), "#7570b3",  # violet
                                   "black")))

corr_order <- corrMatOrder(results_descriptive$danish$proteomic_heatmap_danish, order = "hclust", hclust.method = "complete")
col_labels_ordered <- col_labels[corr_order]

corrplot(results_descriptive$danish$proteomic_heatmap_danish,
         method = "color",
         type = "lower",
         order = "hclust",
         hclust.method = "complete",
         # order = "original",
         tl.col = col_labels_ordered,  # couleurs alignées à l’ordre
         tl.srt = 45,
         tl.cex = 0.4,
         col = rev(COL2(diverging = "RdYlBu")))

rm(vars, col_labels, corr_order, col_labels_ordered)
```

Neuro explo

```{r fig.height=11, fig.width=11}

corrplot(results_descriptive$danish$proteomic_heatmap_danish_neuro_explo, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black',
         order = "hclust",         # regroupement des variables selon niveau de cor
         tl.srt = 45, 
         # addCoef.col = "black",
         # number.cex = 0.8,
         # number.digits = 1,
         tl.cex = 0.4,             # taille police des variables
         col = rev(COL2(diverging = "RdYlBu")))
```

Immune response

```{r fig.height=11, fig.width=11}

corrplot(results_descriptive$danish$proteomic_heatmap_danish_immun_res, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black',
         order = "hclust",         # regroupement des variables selon niveau de cor
         tl.srt = 45, 
         # addCoef.col = "black",
         # number.cex = 0.8,
         # number.digits = 1,
         tl.cex = 0.4,             # taille police des variables
         col = rev(COL2(diverging = "RdYlBu")))
```

Metabolism

```{r fig.height=11, fig.width=11}

corrplot(results_descriptive$danish$proteomic_heatmap_danish_metabolism, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black',
         order = "hclust",         # regroupement des variables selon niveau de cor
         tl.srt = 45, 
         # addCoef.col = "black",
         # number.cex = 0.8,
         # number.digits = 1,
         tl.cex = 0.4,             # taille police des variables
         col = rev(COL2(diverging = "RdYlBu")))

```

### Boxplots

```{r, fig.width=10, fig.height=40}
results_descriptive$danish$proteomic_boxplot_danish
```

Comparison between cases and controls

```{r, fig.width=12, fig.height=60}
results_descriptive$danish$proteomic_boxplot_danish_by_als
```

Comparison between alive/deceased status at the end of the follow-up (among cases)

```{r, fig.width=12, fig.height=60}
results_descriptive$danish$proteomic_boxplot_danish_by_death
```

### NEFL

```{r, fig.height=7, fig.width=9}
results_descriptive$danish$figure_NEFL_1
```

## EVs

### Distribution tables

```{r}
results_descriptive$danish$EVs_table_danish |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")|>
  autofit()
```

### Comparison of distribution cases/controls

```{r}
results_descriptive$danish$EVs_table_danish_by_als |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

```{r, fig.height=12, fig.width=12}
corrplot(results_descriptive$danish$EVs_heatmap_danish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))

```

### Boxplots

```{r, fig.height=7, fig.width=10}
results_descriptive$danish$EVs_boxplot_danish
```

Comparison between cases and controls

```{r, fig.height=10, fig.width=10}
results_descriptive$danish$EVs_boxplot_danish_by_als
```

Comparison between alive/deceased status at the end of the follow-up (among cases)

```{r, fig.width=10, fig.height=10}
results_descriptive$danish$EVs_boxplot_danish_by_death
```

# Finnish data description

## ALS survival

```{r, fig.height=2, fig.width=3}
bdd_finnish |> 
  filter(als == 1) |>
  select(follow_up_death, status_death, study) |>
  tbl_summary(by = "study") |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

densityplot(data = bdd_finnish, vars = "follow_up_death") 
descrip_num(data = bdd_finnish, vars = "follow_up_death") |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

## Metadata

```{r}
results_descriptive$finnish$covar_finnish |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

## POPs

### Distribution tables

```{r}
results_descriptive$finnish$POPs_table_finnish |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Comparison of distribution cases/controls

```{r}
results_descriptive$finnish$POPs_table_finnish_by_als |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

```{r fig.height=11, fig.width=11}
corrplot(results_descriptive$finnish$POPs_heatmap_finnish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))
```

### Boxplots

```{r, fig.width=10, fig.height=8}
# results_descriptive$finnish$POPs_boxplot_finnish
results_descriptive$finnish$POPs_boxplot_finnish_by_als
```

### Detailed distribution

#### Summed

```{r}
densityplot(data = bdd_finnish, vars = POPs_group)
boxplot(data = bdd_finnish, vars = POPs_group)
```

```{r, fig.height=4, fig.width=4}
heatmap_cor(cormat = bdd_finnish[POPs_group], decimal = 1)
```

#### PCBs

```{r}
PCBs <- bdd_finnish |> select(all_of(POPs_finnish)) |> select(contains("PCB")) |> colnames()
densityplot(data = bdd_finnish, vars = PCBs[1:9])
```

```{r, fig.height=3.33}
densityplot(data = bdd_finnish, vars = PCBs[10:13])
```

```{r}
boxplot(data = bdd_finnish, vars = PCBs[1:10])
```

```{r, fig.height=2}
boxplot(data = bdd_finnish, vars = PCBs[11:13])
```

```{r}
heatmap_cor(cormat = bdd_finnish[PCBs], decimal = 1)
rm(PCBs)
```

#### OCPs

```{r}
OCPs <- bdd_finnish |> select(all_of(POPs_finnish)) |> select(contains("OCP")) |> colnames()
densityplot(data = bdd_finnish, vars = OCPs)
boxplot(data = bdd_finnish, vars = OCPs)
# heatmap_cor(cormat = bdd_finnish[OCPs], decimal = 1)
```

#### PBDEs

```{r,  fig.width=7, fig.height=2}
PBDEs <- bdd_finnish |> select(all_of(POPs_finnish)) |> select(contains("PBDE")) |> colnames()
densityplot(data = bdd_finnish, vars = PBDEs)
boxplot(data = bdd_finnish, vars = PBDEs)
```

```{r,  fig.width=7, fig.height=4}
heatmap_cor(cormat = bdd_finnish[PBDEs], decimal = 1)
```

## Fatty acids

### Distribution tables

```{r}
results_descriptive$finnish$fattyacids_table_finnish |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()
```

### Comparison of distribution cases/controls

```{r}
results_descriptive$finnish$fattyacids_table_finnish_by_als |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Heatmap of correlation

```{r, fig.height=14, fig.width=14}
corrplot(results_descriptive$finnish$fattyacids_heatmap_finnish, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))
```

### Boxplots

```{r, fig.height=8, fig.width=10}
results_descriptive$finnish$fattyacids_boxplot_finnish_by_als
```

### Detailed distribution

```{r, fig.height=10, fig.width=10}
bdd_finnish |> 
  select(sample, all_of(fattyacids_tot)) |>
  pivot_longer(cols = -sample, names_to = "fattyacids", values_to = "values") |>
    mutate(fattyacids = fct_recode(fattyacids, !!!fattyacids_tot_labels)) |>
  ggplot() +
  aes(x = "", y = values) +
  geom_boxplot(shape = "circle", fill = "#112446") +
  coord_flip() +
  theme_bw() +
  theme(axis.title = element_blank()) +
  facet_wrap(~fattyacids, scales = "free", ncol = 3L, nrow = 14L)
```

## Heatmap POPs/Fatty acids

```{r, fig.width=12, fig.height=12}
results_descriptive$finnish$POPs_fattyacids_heatmap_finnish
```

## Multiple cholesterol variables issue

Why do we have 3 cholesterol variables in the finnish data ?

```{r, fig.height=2}
boxplot(data = bdd_finnish, vars = c("fS_Kol", "chol", "cholesterol"))

descrip_num(data = bdd_finnish, vars = c("fS_Kol", "chol", "cholesterol")) |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") 
heatmap_cholesterol <- bdd_finnish |> select("fS_Kol", "chol", "cholesterol")
heatmap_cholesterol <- cor(heatmap_cholesterol, 
                    use = "pairwise.complete.obs", 
                    method = "pearson")

corrplot(heatmap_cholesterol, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))

```

## Timing between baseline, ALS diagnosis and death

### FMC

```{r}
timing <- bdd |> 
  filter(study == "FMC") |>
  filter(als == 1) |>
  select(baseline_age, diagnosis_age, death_age) |>
  mutate(time_baseline_diagnosis = diagnosis_age - baseline_age, 
         time_baseline_death = death_age - baseline_age, 
         time_diagnosis_death = death_age - diagnosis_age) 

timing |> 
  tbl_summary() |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

descrip_num(data = timing, vars = colnames(timing)) |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
rm(timing)
```

### FMCF

```{r}
timing <- bdd |> 
  filter(study == "FMCF") |>
  filter(als == 1) |>
  select(baseline_age, diagnosis_age, death_age, 
         time_baseline_diagnosis, time_baseline_death, time_diagnosis_death)

timing |> 
  tbl_summary() |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
   set_table_properties(align = "left")|>
  set_table_properties(align = "left")

descrip_num(data = timing, vars = colnames(timing)) |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
rm(timing)
```

### MFH

```{r}
timing <- bdd |> 
  filter(study == "MFH") |>
  filter(als == 1) |>
  select(baseline_age, diagnosis_age, death_age, 
         time_baseline_diagnosis, time_baseline_death, time_diagnosis_death)

timing |> 
  tbl_summary() |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

descrip_num(data = timing, vars = colnames(timing)) |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
rm(timing)
```

# Comparison danish/finnish data

## On the total population

### Metadata

```{r}
results_descriptive$comp$covar_comp |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### POPs

```{r, fig.width=10, fig.height=10}
results_descriptive$comp$POPs_table_comp |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
results_descriptive$comp$POPs_boxplot_comp
```

### Fatty acids

```{r, fig.height=12, fig.width=12}
results_descriptive$comp$fattyacids_table_comp |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

results_descriptive$comp$fattyacids_boxplot_comp 
```

## Among cases

(for ALS survival analyses)

### Metadata

```{r}
results_descriptive$comp$covar_comp_cases |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### POPs

```{r, fig.width=9}
results_descriptive$comp$POPs_table_comp_cases |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")

results_descriptive$comp$POPs_boxplot_comp_cases
```

```{r fig.height=12, fig.width=12}
#| echo: false
#| message: false
#| warning: false
#| paged-print: false
corrplot(results_descriptive$comp$POPs_heatmap_cases, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))
```

# POPs and ALS occurence

## Danish cohort

### Effects of subject characteristics on ALS occurence

```{r}
results_POPs_ALS_occurrence$main$covar |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Main analysis

#### Quartile models

In the copollutant models, the pollutant of interest is quartile while the others are s() in a GAM model

```{r fig.height=9, fig.width=9}
results_POPs_ALS_occurrence$main$plot_quart
```

```{r}
#| column: screen-inset-right
extra_rows <- results_POPs_ALS_occurrence$main$results_quart |>
  distinct(variable) |> 
  mutate(
    quartiles = "Quartile 1",
    "OR_base" = '-', "95%CI_base" = '-', "p.value_base" = '', "p.value_heterogeneity_base" = '', "p.value_trend_base" = '',
    "OR_adjusted" = '-', "95%CI_adjusted" = '-', "p.value_adjusted" = '', , "p.value_heterogeneity_adjusted" = '', "p.value_trend_adjusted" = '',
    "OR_copollutant" = '-', "95%CI_copollutant" ='-', "p.value_copollutant" = '', , "p.value_heterogeneity_copollutant" = '', "p.value_trend_copollutant" = '')

table_S2 <- 
  results_POPs_ALS_occurrence$main$main_results |>
  select(-lower_CI, -upper_CI, - p.value_raw) |>
  filter(model %in% c('base_quart', 'adjusted_quart', 'copollutant_quart')) |>
  pivot_wider(
    names_from = model,  
    values_from = c(OR, `95%CI`, p.value, p.value_heterogeneity, p.value_trend)) |>
  select(variable, 
         quartiles = df,
         contains("base_quart"), 
         contains("adjusted_quart"), 
         contains("copollutant_quart")) 
colnames(table_S2) <- gsub('_quart', '', colnames(table_S2))

table_S2 <- table_S2 |>
  mutate_if(is.numeric, as.character) |>
  bind_rows(extra_rows) |>
  mutate(
    variable = gsub("OCP_", "", variable), 
    variable = fct_relevel(variable, "PCB_4", "PCB_DL", "PCB_NDL",  "HCB", "ΣDDT", "β_HCH", "Σchlordane", "ΣPBDE" ),
    variable = fct_recode(variable, 
                          "Most prevalent PCBs" = "PCB_4",
                          "Dioxin-like PCBs" = "PCB_DL",
                          "Non-dioxin-like PCBs" = "PCB_NDL",
                          "β-HCH" = "β_HCH")) |>
  arrange(variable, quartiles) |>
  group_by(variable) |>
  mutate(p.value_heterogeneity_base = ifelse(quartiles == 'Quartile 1', p.value_heterogeneity_base[quartiles == 'Quartile 2'], ''), 
         p.value_trend_base = ifelse(quartiles == 'Quartile 1', p.value_trend_base[quartiles == 'Quartile 2'], ''),
         p.value_heterogeneity_adjusted = ifelse(quartiles == 'Quartile 1', p.value_heterogeneity_adjusted[quartiles == 'Quartile 2'], ''), 
         p.value_trend_adjusted = ifelse(quartiles == 'Quartile 1', p.value_trend_adjusted[quartiles == 'Quartile 2'], ''), 
         p.value_heterogeneity_copollutant = ifelse(quartiles == 'Quartile 1', p.value_heterogeneity_copollutant[quartiles == 'Quartile 2'], ''), 
         p.value_trend_copollutant = ifelse(quartiles == 'Quartile 1', p.value_trend_copollutant[quartiles == 'Quartile 2'], '')) |>
  ungroup() |>
  rename('OR' = 'OR_base', '95% CI' = '95%CI_base', 'p-value' = 'p.value_base', 
         'Heterogeneity test' = 'p.value_heterogeneity_base', 'Trend test' = 'p.value_trend_base',
         'OR ' = 'OR_adjusted', '95% CI ' = '95%CI_adjusted', 'p-value ' = 'p.value_adjusted', 
         'Heterogeneity test ' = 'p.value_heterogeneity_adjusted', 'Trend test ' = 'p.value_trend_adjusted',
         ' OR ' = 'OR_copollutant', ' 95% CI ' = '95%CI_copollutant', ' p-value ' = 'p.value_copollutant',
         ' Heterogeneity test ' = 'p.value_heterogeneity_copollutant', ' Trend test ' = 'p.value_trend_copollutant')
table_S2 <- table_S2 |> flextable() |>
  add_footer_lines(
    "1POPs were summed as follows: Dioxin-like PCBs corresponds to PCBs 118 and 156; non-dioxin-like PCBs corresponds to PCBs 28, 52, 74, 99, 101, 138, 153, 170, 180, 183, 187; most prevalent PCBs corresponds to PCBs 118, 138, 153, 180; ΣPBDE corresponds to PBDEs 47, 99, 153; ΣDDT corresponds to p,p’-DDT and p,p’-DDE and finally Σchlordane corresponds to trans-nonanchlor and oxychlordane.
  2All models are matched for sex and age. Adjusted models further account for smoking, BMI, serum total cholesterol, marital status, and education. Co-pollutant models further include all chemicals, where a GAM is used in which the chemical of interest is categorized into quartiles, while the other pollutants are included as smooth terms.
  3Estimated risk of ALS when exposures to POP are at quartiles 2, 3, and 4, compared to quartile 1.
  4CI: Confidence interval.
  5Heterogeneity tests in outcome value across POP quartiles, matched on age and sex.
  6Trend tests using continuous variables whose values corresponded to the quartile specific median POP levels, matched on age and sex.
  7Heterogeneity tests in outcome value across POP quartiles, matched on age and sex, and adjusted for smoking, BMI, serum total cholesterol, marital status, and education.
  8Trend tests using continuous variables whose values corresponded to the quartile specific median POP levels, matched on age and sex, and adjusted for smoking, BMI, serum total cholesterol, marital status, and education.") |>
  add_header(
    "variable" = "Exposures", "quartiles" = "Quartiles",
    "OR" = "Base Model", "95% CI" = "Base Model", "p-value" = "Base Model", "Heterogeneity test" = 'Base Model', "Trend test" = 'Base Model',
    "OR " = "Adjusted Model", "95% CI " = "Adjusted Model", "p-value " = "Adjusted Model", "Heterogeneity test " = 'Adjusted Model', "Trend test " = 'Adjusted Model',
    " OR " = "Co-pollutant model", " 95% CI " = "Co-pollutant model", " p-value " = "Co-pollutant model", " Heterogeneity test " = 'Co-pollutant model', " Trend test " = 'Co-pollutant model') |>
  merge_h(part = "header") |>
  merge_v(j = "variable") |>
  theme_vanilla() |>
  bold(j = "variable", part = "body") |>
  align(align = "center", part = "all") |>
  align(j = "variable", align = "left", part = "all") |> 
  merge_at(j = "variable", part = "header") |>
  merge_at(j = "quartiles", part = "header") |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all")|>
  set_table_properties(align = "left")

table_S2
rm(extra_rows, table_S2)
```

#### q-gcomp analysis

```{r, fig.height=3}
results_POPs_ALS_occurrence$main$POPs_ALS_qgcomp_table_danish |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
results_POPs_ALS_occurrence$main$POPs_ALS_qgcomp_figure_danish
```

#### GAM models

Relationship between POPs exposure at baseline time and ALS occurrence estimated using a generalized additive model (gam) from the Danish Diet, Cancer and Health study cohort established between December 1993 and May 1997. The solid blue line represents the predicted ALS occurrence by POPs exposure given that all other covariates are at their respective means. The light blue represents the 95% confidence intervals. All models were adjusted for sex and age. Adjusted models further account for serum cholesterol, BMI, smoking status, marital status and education. The effective degrees of freedom (EDF) for the POPs smooth terms is a summary statistic of a gam that reflects the degree of nonlinearity (edf=1: linear, edf=2: quadratic, edf\>2: more complex than quadratic).

```{r, fig.height=3.5, fig.width=9}
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$PCB_DL, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$PCB_DL, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$PCB_DL, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$PCB_NDL, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$PCB_NDL, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$PCB_NDL, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$PCB_4, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$PCB_4,  
  wrap_elements(grid::nullGrob()),
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$OCP_HCB, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$OCP_HCB, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$OCP_HCB, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$ΣDDT, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$ΣDDT, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$ΣDDT, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$OCP_β_HCH, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$OCP_β_HCH, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$OCP_β_HCH, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$Σchlordane, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$Σchlordane, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$Σchlordane, 
  nrow = 1)
wrap_plots(
  results_POPs_ALS_occurrence$main$plot_base_gam$ΣPBDE, 
  results_POPs_ALS_occurrence$main$plot_adjusted_gam$ΣPBDE, 
  results_POPs_ALS_occurrence$main$plot_copollutant_gam$ΣPBDE, 
  nrow = 1)
```

### Sensitivity analysis total lipid instead of total cholesterol

```{r fig.height=8, fig.width=9}

results_POPs_ALS_occurrence$main$plot_quart_sensi_lipid
```

```{r fig.height=9, fig.width=9}
#| column: screen-inset-right

results_POPs_ALS_occurrence$main$table_quart_sensi_lipid

```

### Sensitivity analysis about diabetes

```{r fig.height=8, fig.width=9}

results_POPs_ALS_occurrence$main$plot_quart_sensi_diabetes
```

```{r fig.height=9, fig.width=9}
#| column: screen-inset-right

results_POPs_ALS_occurrence$main$table_quart_sensi_diabetes

```

### Sensitivity analysis POPs not summed

#### quartiles

```{r,  fig.height=15, fig.width=10}
results_POPs_ALS_occurrence$sensitivity_not_summed$plot_quart_sensi_not_summed
```

#### GAMs

```{r, fig.height=6, fig.width=15}
#| column: screen-inset-right
wrap_plots(results_POPs_ALS_occurrence$sensitivity_not_summed$plot_adjusted_gam_not_summed[1:8], ncol = 4)
wrap_plots(results_POPs_ALS_occurrence$sensitivity_not_summed$plot_adjusted_gam_not_summed[9:16], ncol = 4)
wrap_plots(results_POPs_ALS_occurrence$sensitivity_not_summed$plot_adjusted_gam_not_summed[17:22], ncol = 4)
```

## Metanalysis (quartiles)

```{r fig.height=7, fig.width=9}
results_POPs_ALS_occurrence$metanalysis$plot_metanalysis_quart
```

```{r}
#| column: screen-inset-right
extra_rows <- results_POPs_ALS_occurrence$metanalysis$metanalysis_quart |>
  distinct(explanatory) |> 
  mutate(
    term = "Quartile 1",
      "OR_base" = '-', "95%CI_base" = '-', "p-value_base" = '', "p-value_heterogeneity_base" = "", 
      "OR_adjusted" = '-', "95%CI_adjusted" = '-', "p-value_adjusted" = '', "p-value_heterogeneity_adjusted" = "", 
    "OR_copollutant" = '-', "95%CI_copollutant" = '-', "p-value_copollutant" = '', "p-value_heterogeneity_copollutant" = "")
    

table_S4 <- results_POPs_ALS_occurrence$metanalysis$metanalysis_quart |>
  select(model, explanatory, term, OR, `95%CI`, "p-value", 'p-value_heterogeneity') |>
  pivot_wider(
    names_from = model,  
    values_from = c("OR", "95%CI", "p-value", 'p-value_heterogeneity')) |>
  bind_rows(extra_rows) |>
  select(Exposures = explanatory, Quartiles = term,
         contains("base"), 
         contains("adjusted"), 
         contains("copollutant")) |>
  mutate(
    Exposures = gsub("OCP_", "", Exposures), 
    Exposures = fct_relevel(Exposures, "PCB_4", "PCB_DL", "PCB_NDL", "HCB", "ΣDDT", "β_HCH", "Σchlordane"),
    Exposures = fct_recode(Exposures, 
                          "Most prevalent PCBs" = "PCB_4",
                          "Dioxin-like PCBs" = "PCB_DL",
                          "Non-dioxin-like PCBs" = "PCB_NDL",
                          "β-HCH" = "β_HCH")) |>
  arrange(Exposures, Quartiles) 

table_S4 <- table_S4 |> 
  rename('OR' = 'OR_base', '95% CI' = '95%CI_base', 'p-value' = 'p-value_base', "Hetero-geneity test" = "p-value_heterogeneity_base",
         'OR ' = 'OR_adjusted', '95% CI ' = '95%CI_adjusted', 'p-value ' = 'p-value_adjusted', "Hetero-geneity test " = "p-value_heterogeneity_adjusted",
         ' OR ' = 'OR_copollutant', ' 95% CI ' = '95%CI_copollutant', ' p-value ' = 'p-value_copollutant', " Hetero-geneity test " = "p-value_heterogeneity_copollutant") |>
  flextable() |>
  add_footer_lines(
  "1POPs were summed as follows: Dioxin-like PCBs corresponds to PCBs 118 and 156; non-dioxin-like PCBs corresponds to PCBs 28, 52, 74, 99, 101, 138, 153, 170, 180, 183, 187; most prevalent PCBs corresponds to PCBs 118, 138, 153, 180; ΣDDT corresponds to p,p’-DDT and p,p’-DDE and finally Σchlordane corresponds to trans-nonanchlor and oxychlordane.
  2Base and adjusted models are conditional logistic regressions matched for sex and age, and for municipality and serum thawing history (Finnish cohorts). Adjusted models further account for smoking, BMI, serum total cholesterol, marital status, and education (except for FMC cohort for which education data was not available). Co-pollutant models further include all chemicals, where the chemical of interest is categorized into quartiles, while the others are treated as smooth terms in GAMs.
  3Estimated risk of ALS when exposures to POP are at quartiles 2, 3, and 4, compared to quartile 1.
  4CI: Confidence interval.
  5P-value of between-study heterogeneity (Cochran’s Q test).") |>
  add_header(
    "Exposures" = "Exposures", "Quartiles" = "Quartiles",
    "OR" = "Base model", "95% CI" = "Base model", "p-value" = "Base model", "Hetero-geneity test" = "Base model",
    "OR " = "Adjusted model", "95% CI " = "Adjusted model", "p-value " = "Adjusted model", "Hetero-geneity test " = "Adjusted model",
    " OR " = "Copollutant model", " 95% CI " = "Copollutant model", " p-value " = "Copollutant model", " Hetero-geneity test " = "Copollutant model") |>
  merge_h(part = "header") |>
  merge_v(j = "Exposures") |>
  theme_vanilla() |>
  bold(j = "Exposures", part = "body") |>
  align(align = "center", part = "all") |>
  align(j = "Exposures", align = "left", part = "all") |> 
  merge_at(j = "Exposures", part = "header") |>
  merge_at(j = "Quartiles", part = "header") |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all")|>
  set_table_properties(align = "left")
table_S4
rm(extra_rows, table_S4)
```

# POPs ans ALS survival

## Subjects characteristics (among cases only)

```{r}
#| echo: false
#| message: false
#| warning: false
#| paged-print: false
results_descriptive$comp$covar_comp_cases |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

## Danish cohort

### Effects of subject characteristics on ALS survival

```{r}
results_POPs_ALS_survival$main_analysis$covar_danish |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Effects of POPs on ALS survival

#### Continuous coding (standardized)

```{r}
results_POPs_ALS_survival$main_analysis$POPs_sd_ALS_table_danish  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.height=4}
results_POPs_ALS_survival$main_analysis$POPs_sd_ALS_figure_danish
```

#### Quartiles

```{r}
#| column: screen-inset-right
results_POPs_ALS_survival$main_analysis$POPs_quart_ALS_table_danish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r fig.height=9, fig.width=9}
results_POPs_ALS_survival$main_analysis$POPs_quart_ALS_figure_danish
```

#### Cox-gam

```{r fig.height=11, fig.width=8}

wrap_plots(list(results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$PCB_4, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_4, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$PCB_DL, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_DL, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$PCB_NDL, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_NDL, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$OCP_HCB, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$OCP_HCB), 
           ncol = 2)

wrap_plots(list(results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$ΣDDT, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$ΣDDT, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$OCP_β_HCH, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$OCP_β_HCH, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$Σchlordane, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$Σchlordane, 
                
                results_POPs_ALS_survival$main_analysis$plot_base_cox_gam_danish$ΣPBDE, 
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$ΣPBDE), 
           ncol = 2)

```

#### Q-gcomp analysis

Including all the pollutants

```{r, fig.height=3}
results_POPs_ALS_survival$main_analysis$qgcomp$POPs_ALS_qgcomp_table_danish  |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
results_POPs_ALS_survival$main_analysis$qgcomp$POPs_ALS_qgcomp_figure_danish
```

Selecting the pollutants with a positive weight and adjusting for the others

```{r, fig.height=2}
results_POPs_ALS_survival$main_analysis$qgcomp$POPs_ALS_qgcomp_positive_table_danish  |>
  flextable() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
results_POPs_ALS_survival$main_analysis$qgcomp$POPs_ALS_qgcomp_positive_figure_danish
```

## Sensitivity analyses 1

Mixture models with POPs pre-selection

### Model forcing the covariates to be included

Using LASSO

```{r}
plot(results_POPs_ALS_survival$sensi1$sensi1_lasso_sd_danish)  
```

Using ridge

```{r}
plot(results_POPs_ALS_survival$sensi1$sensi1_ridge_sd_danish)  
```

Using elastic net

```{r}
plot(results_POPs_ALS_survival$sensi1$sensi1_elastic_net_sd_danish)
```

These plot the cross-validation curve (red dotted line) along with upper and lower standard deviation curves along the 𝜆sequence (error bars). Two special values along the 𝜆sequence are indicated by the vertical dotted lines.

lambda.min is the value of 𝜆that gives minimum mean cross-validated error

coefficients using lambda.min Using LASSO

```{r}
coef(results_POPs_ALS_survival$sensi1$sensi1_lasso_sd_danish, s = "lambda.min")  
```

Using ridge

```{r}
coef(results_POPs_ALS_survival$sensi1$sensi1_ridge_sd_danish, s = "lambda.min")  
```

Using elastic net

```{r}
coef(results_POPs_ALS_survival$sensi1$sensi1_elastic_net_sd_danish, s = "lambda.min")
```

while lambda.1se is the value of 𝜆that gives the most regularized model such that the cross-validated error is within one standard error of the minimum.

coefficients using lambda.1se Using LASSO

```{r}
coef(results_POPs_ALS_survival$sensi1$sensi1_lasso_sd_danish, s = "lambda.1se")  
```

Using ridge

```{r}
coef(results_POPs_ALS_survival$sensi1$sensi1_ridge_sd_danish, s = "lambda.1se")  
```

Using elastic net

```{r}
coef(results_POPs_ALS_survival$sensi1$sensi1_elastic_net_sd_danish, s = "lambda.1se")
```

### New co-pollutant model with selected POPs using elastic net (to get unbiaised coefficients and 95%CI and p-values)

#### Continuous coding

```{r, fig.height=2}
results_POPs_ALS_survival$sensi1$POPs_sd_ALS_figure_sensi1_danish
results_POPs_ALS_survival$sensi1$POPs_sd_ALS_table_sensi1_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

#### Quartile coding

```{r, fig.height=3}
results_POPs_ALS_survival$sensi1$POPs_quart_ALS_figure_sensi1_danish
results_POPs_ALS_survival$sensi1$POPs_quart_ALS_table_sensi1_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

### ERS model with the POPs selestec with elastic net

#### Continuous coding (standardized)

```{r, fig.height=2}
results_POPs_ALS_survival$sensi1$POPs_sd_ALS_figure_sensi1_ERS_danish
results_POPs_ALS_survival$sensi1$POPs_sd_ALS_table_sensi1_ERS_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

#### Quartile coding

```{r, fig.height=3}
results_POPs_ALS_survival$sensi1$POPs_quart_ALS_figure_sensi1_ERS_danish
results_POPs_ALS_survival$sensi1$POPs_quart_ALS_table_sensi1_ERS_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

## Sensitivity analyses 2

### Model forcing the covariates to be included

coefficients using lambda.min (lambda.min is the value of 𝜆that gives minimum mean cross-validated error)

Using LASSO

```{r}
coef(results_POPs_ALS_survival$sensi2$sensi2_lasso_sd_danish, s = "lambda.min")  
```

Using elastic net

```{r}
coef(results_POPs_ALS_survival$sensi2$sensi2_elastic_net_sd_danish, s = "lambda.min")
```

### New co-pollutant model with selected POPs using elastic net (to get unbiaised coefficients and 95%CI and p-values)

#### Continuous coding (standardized)

```{r, fig.height=3}
results_POPs_ALS_survival$sensi2$POPs_sd_ALS_figure_sensi2_danish
results_POPs_ALS_survival$sensi2$POPs_sd_ALS_table_sensi2_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

#### Quartile coding

```{r}
results_POPs_ALS_survival$sensi2$POPs_quart_ALS_figure_sensi2_danish
results_POPs_ALS_survival$sensi2$POPs_quart_ALS_table_sensi2_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

### ERS model with the POPs selestec with elastic net

#### Continuous coding (standardized)

```{r, fig.height=2}
results_POPs_ALS_survival$sensi2$POPs_sd_ALS_figure_sensi2_ERS_danish
results_POPs_ALS_survival$sensi2$POPs_sd_ALS_table_sensi2_ERS_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

#### Quartile coding

```{r, fig.height=3, fig.width=8}
results_POPs_ALS_survival$sensi2$POPs_quart_ALS_figure_sensi2_ERS_danish
results_POPs_ALS_survival$sensi2$POPs_quart_ALS_table_sensi2_ERS_danish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

## Sensitivity analyses 3

```{r, fig.height=3, fig.width=8}
results_POPs_ALS_survival$sensi3$sensi3_figure
```

```{r}
results_POPs_ALS_survival$sensi3$sensi3_table
```

## Sensitivity analysis 4

### Cox models (sd)

```{r, fig.height=4}
results_POPs_ALS_survival$sensi4$POPs_sd_ALS_table_danish_sensi_4 |>
  flextable::set_table_properties(align = "left", layout = "autofit")
results_POPs_ALS_survival$sensi4$POPs_sd_ALS_figure_danish_sensi_4
```

### Cox-GAMs models

```{r fig.height=11, fig.width=8}

wrap_plots(list(results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_4,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$PCB_4,
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_DL,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$PCB_DL,
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$PCB_NDL,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$PCB_NDL,
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$OCP_HCB,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$OCP_HCB), 
           ncol = 2)

wrap_plots(list(
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$ΣDDT,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$ΣDDT, 
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$OCP_β_HCH,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$OCP_β_HCH,
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$Σchlordane,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$Σchlordane,
                
                results_POPs_ALS_survival$main_analysis$plot_adjusted_cox_gam_danish$ΣPBDE,
                results_POPs_ALS_survival$sensi4$plot_adjusted_cox_gam_danish_sensi_4$ΣPBDE), 
           ncol = 2)

```

# Fatty acids and ALS occurence

## Danish cohort

### Effects of subject characterist on ALS occurence in the Danish cohort

```{r}
results_fattyacids_ALS_occurrence$danish$covar_als_table_danish  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### Effects of fatty acids on ALS occurence in the Danish cohort

#### Continuous coding (standardized)

```{r}
results_fattyacids_ALS_occurrence$danish$fattyacids_sd_als_table_danish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r}
results_fattyacids_ALS_occurrence$danish$fattyacids_sd_als_figure_danish
```

#### Quartiles

```{r}
results_fattyacids_ALS_occurrence$danish$fattyacids_quart_als_table_danish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.height=12}
results_fattyacids_ALS_occurrence$danish$fattyacids_quart_als_figure_danish
```

## Finnish cohorts

### Effects of fatty acids on ALS occurence in the Finnish cohorts

#### Continuous coding (standardized)

```{r}
results_fattyacids_ALS_occurrence$finnish$fattyacids_sd_als_table_finnish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r}
results_fattyacids_ALS_occurrence$finnish$fattyacids_sd_als_figure_finnish
```

#### Quartiles

```{r}
results_fattyacids_ALS_occurrence$finnish$fattyacids_quart_als_table_finnish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.height=12}
results_fattyacids_ALS_occurrence$finnish$fattyacids_quart_als_figure_finnish
```

### Sensitivity Ca

#### Continuous coding (standardized)

```{r}
results_fattyacids_ALS_occurrence$sensi_ca$fattyacids_sd_als_table_sensi_ca_finnish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
results_fattyacids_ALS_occurrence$sensi_ca$fattyacids_sd_als_figure_sensi_ca_finnish
```

#### Quartiles

```{r}
results_fattyacids_ALS_occurrence$sensi_ca$fattyacids_quart_als_table_sensi_ca_finnish |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

```{r, fig.height=13}
results_fattyacids_ALS_occurrence$sensi_ca$fattyacids_quart_als_figure_sensi_ca_finnish
```

# Fatty acids ans ALS survival

## Danish cohort

### Effects of covariates on ALS survival

```{r}
results_fattyacids_ALS_survival$danish$covar_danish |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

### Continuous coding (standardized)

```{r}
results_fattyacids_ALS_survival$danish$fattyacids_sd_als_table_danish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
results_fattyacids_ALS_survival$danish$fattyacids_sd_als_figure_danish
```

### Quartiles

```{r}
results_fattyacids_ALS_survival$danish$fattyacids_quart_als_table_danish |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.height=12}
results_fattyacids_ALS_survival$danish$fattyacids_quart_als_figure_danish
```

```{r fig.height=12, fig.width=12}
#| eval: false
#| column: screen-inset-right
#| include: false
arrange_ggsurvplots(
  results_fattyacids_ALS_survival$danish$survival_plots_danish[1:4],
  ncol = 2,              
  nrow = 2)
```

# NfL and ALS risk

## Main analysis

### Continuous coding (standardized)

```{r, fig.height=2}
results_proteomic_ALS_occurrence$NfL_results$NfL_sd_ALS_figure_sensi_1 
```

```{r}
results_proteomic_ALS_occurrence$NfL_results$NfL_sd_ALS_table_sensi_1 |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

### Quartiles

```{r}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$NfL_results$NfL_quart_ALS_table_sensi_1 |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

### GAMs

```{r, fig.height=4}
wrap_plots(results_proteomic_ALS_occurrence$sensi_1$plot_base_gam_sensi_1$signif$proteomic_neuro_explo_NEFL_sensi_1,  
           results_proteomic_ALS_occurrence$sensi_1$plot_adjusted_gam_sensi_1$signif$proteomic_neuro_explo_NEFL_sensi_1)
```

## Sensitivity analysis: effect of follow_up duration

### Continuous (standardized)

```{r, fig.height=3}
results_proteomic_ALS_occurrence$NfL_results$NfL_sd_ALS_figure_sensi_follow_up_base_adj
```

```{r}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$NfL_results$NfL_sd_ALS_table_sensi_follow_up_base_adj |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

### GAMs

```{r, fig.height=7, fig.width=9}
wrap_plots(results_proteomic_ALS_occurrence$sensi_1$plot_adjusted_gam_sensi_1$signif$proteomic_neuro_explo_NEFL_sensi_1, 
           results_proteomic_ALS_occurrence$sensi_2$plot_adjusted_gam_sensi_2$signif$proteomic_neuro_explo_NEFL, 
           results_proteomic_ALS_occurrence$sensi_1_3$plot_adjusted_gam_sensi_1_3$signif$proteomic_neuro_explo_NEFL_sensi_1_3, 
           results_proteomic_ALS_occurrence$sensi_1_3_4$plot_adjusted_gam_sensi_1_3_4$signif$proteomic_neuro_explo_NEFL_sensi_1_3_4, 
           results_proteomic_ALS_occurrence$sensi_1_3_5$plot_adjusted_gam_sensi_1_3_5$not_signif$proteomic_neuro_explo_NEFL_sensi_1_3_5)
```

## Sensitivity analysis: effect of sex

### Continuous (standardized)

```{r}
results_proteomic_ALS_occurrence$NfL_results$NfL_sd_ALS_table_sensi_sex_adj 
```

### GAMs

```{r, fig.height=4, fig.width=9}
wrap_plots(results_proteomic_ALS_occurrence$sensi_1$plot_adjusted_gam_sensi_1$signif$proteomic_neuro_explo_NEFL_sensi_1, 
           results_proteomic_ALS_occurrence$sensi_1_7$plot_adjusted_gam_sensi_1_7_female$signif$proteomic_neuro_explo_NEFL_sensi_1_7_female, 
           results_proteomic_ALS_occurrence$sensi_1_7$plot_adjusted_gam_sensi_1_7_male$signif$proteomic_neuro_explo_NEFL_sensi_1_7_male, 
           ncol = 3)
```

## Additionnal analysis: NfL over time

The locally estimated scatterplot smoothing (LOESS) curve illustrates the changes in the ratio of NEFL levels in each case–control pair/triplet over time. Each point corresponds to 1 case–control pair/triplet.

```{r}
results_proteomic_ALS_occurrence$NfL_results$NfL_over_time$figure_NfL_over_time
```

After removing the NEFL outlier

```{r}
results_proteomic_ALS_occurrence$NfL_results$NfL_over_time$figure_NfL_over_time_sensi_1
```

## Additional analysis: AUC

```{r, fig.height=8, fig.width=18}
results_proteomic_ALS_occurrence$NfL_results$AUC_NfL$AUC_figure_adjusted_pattern +
  theme(legend.direction = "vertical", legend.position = "right")
results_proteomic_ALS_occurrence$NfL_results$AUC_NfL$AUC_figure_unadjusted_pattern +
  theme(legend.direction = "vertical", legend.position = "right")
```

## Additional analyses: removing individuals with other neurological conditions 
```{r}
results_proteomic_ALS_occurrence$NfL_results$other_disease$table_without_other_diseases
```


# NfL and ALS survival

## Main analyses

### Continuous coding (standardized)

```{r}
results_proteomic_ALS_survival$NfL_results$NfL_sd_ALS_table |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

### Quartiles

```{r}
#| column: screen-inset-right
results_proteomic_ALS_survival$NfL_results$NfL_quart_ALS_table 
```

### GAMs

```{r, fig.height=4}
wrap_plots(results_proteomic_ALS_survival$main$plot_base_cox_gam_danish$not_signif$NEFL, 
           results_proteomic_ALS_survival$main$plot_adjusted_cox_gam_danish$not_signif$NEFL)
```

## Sensitivity analysis: effect of follow-up duration

```{r, fig.height=4}
results_proteomic_ALS_survival$NfL_results$NfL_sd_ALS_figure_sensi_follow_up
```

```{r}
results_proteomic_ALS_survival$NfL_results$NfL_sd_ALS_table_sensi_follow_up |>
  flextable::set_table_properties(align = "left", layout = "autofit")
```

## Sensitivity analysis: effect of sex

```{r}
results_proteomic_ALS_survival$NfL_results$NfL_sd_ALS_table_sensi_sex
```

# Proteomic and ALS occurrence

## Effects of subject characterist on ALS occurence in the Danish cohort

```{r}
results_proteomic_ALS_occurrence$main$covar |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

## Main analysis

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r, fig.width=15, fig.height=6}
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_table |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_base_figure +  theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_adjusted_figure + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

Conditional logistic regressions (quartiles), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$main$proteomic_quart_ALS_table  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

Generalized additive models modeling proteins as smooth terms (thin plate splines), base models adjusted on sex and age at baseline. Adjusted models further account for smoking and body mass index

```{r, fig.width=9, fig.height=60}

plots_lists <- list(
  sensi_1_base = results_proteomic_ALS_occurrence$main$plot_base_gam$signif,
  sensi_1_adj  = results_proteomic_ALS_occurrence$main$plot_adjusted_gam$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(plots_lists, all_proteins, plot_grid)


```

## Sensitivity 1: Removing NEFL outlier

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r, fig.width=15, fig.height=6}
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_table_sensi_1 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_base_figure + labs(title = "Main analysis - Base model") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + labs(title = "NEFL outlier removed - Base model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
  
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_adjusted_figure + labs(title = "Main analysis - adjusted model") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + labs(title = "NEFL outlier removed - Adjusted model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

Conditional logistic regressions (quartiles), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_quart_ALS_table_sensi_1 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

Generalized additive models modeling proteins as smooth terms (thin plate splines), base models adjusted on sex and age at baseline. Adjusted models further account for smoking and body mass index

```{r, fig.width=10, fig.height=5}
wrap_plots(results_proteomic_ALS_occurrence$main$plot_base_gam$signif$proteomic_neuro_explo_NEFL,
           results_proteomic_ALS_occurrence$sensi_1$plot_base_gam_sensi_1$signif$proteomic_neuro_explo_NEFL_sensi_1, 
           ncol = 2)
```

```{r, fig.width=10, fig.height=5}
wrap_plots(results_proteomic_ALS_occurrence$main$plot_adjusted_gam$signif$proteomic_neuro_explo_NEFL,
           results_proteomic_ALS_occurrence$sensi_1$plot_adjusted_gam_sensi_1$signif$proteomic_neuro_explo_NEFL_sensi_1, 
           ncol = 2)
```

## Sensitivity 2: Filtering to cases and their controls with follow-up \< 5 years

### Logistic regressions

```{r}

results_proteomic_ALS_occurrence$sensi_2$proteomic_sd_ALS_table_sensi_2 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r fig.height=6, fig.width=15}
#| column: screen-inset-right

results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_base_figure + labs(title = "Main analysis - Base model") + theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_2$proteomic_sd_ALS_base_figure_sensi_2 + labs(title = "Filtered to follow-up < 5 years - Base model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)

results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_adjusted_figure + labs(title = "Main analysis - Adjusted model") + theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_2$proteomic_sd_ALS_adjusted_figure_sensi_2 + labs(title = "Filtered to follow-up < 5 years - Adjusted model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

```{r}
#| column: screen-inset-right

results_proteomic_ALS_occurrence$sensi_3$proteomic_quart_ALS_table_sensi_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=60, fig.width=12}
#| column: screen-inset-right

plots_lists <- list(
  sensi_1_base = results_proteomic_ALS_occurrence$sensi_1$plot_base_gam_sensi_1$signif,
  sensi_1_adj  = results_proteomic_ALS_occurrence$sensi_1$plot_adjusted_gam_sensi_1$signif,
  sensi_2_base = results_proteomic_ALS_occurrence$sensi_2$plot_base_gam_sensi_2$signif,
  sensi_2_adj  = results_proteomic_ALS_occurrence$sensi_2$plot_adjusted_gam_sensi_2$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 4) 

rm(plots_lists, all_proteins, plot_grid)

```

## Sensitivity 3: Filtering to cases and their controls with follow-up \> 5 years

### Logistic regressions

```{r}
results_proteomic_ALS_occurrence$sensi_3$proteomic_sd_ALS_table_sensi_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r fig.height=6, fig.width=15}
#| column: screen-inset-right

results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_base_figure + labs(title = "Main analysis - Base model") + theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_3$proteomic_sd_ALS_base_figure_sensi_3 + labs(title = "Filtered to follow-up > 5 years - Base model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)

results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_adjusted_figure + labs(title = "Main analysis - Adjusted model") + theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_3$proteomic_sd_ALS_adjusted_figure_sensi_3 + labs(title = "Filtered to follow-up > 5years - Adjusted model") + theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

```{r}
#| column: screen-inset-right

results_proteomic_ALS_occurrence$sensi_3$proteomic_quart_ALS_table_sensi_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=60, fig.width=12}
#| column: screen-inset-right

plots_lists <- list(
  sensi_1_base = results_proteomic_ALS_occurrence$sensi_1$plot_base_gam_sensi_1$signif,
  sensi_1_adj  = results_proteomic_ALS_occurrence$sensi_1$plot_adjusted_gam_sensi_1$signif,
  sensi_3_base = results_proteomic_ALS_occurrence$sensi_3$plot_base_gam_sensi_3$signif,
  sensi_3_adj  = results_proteomic_ALS_occurrence$sensi_3$plot_adjusted_gam_sensi_3$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 4) 

rm(plots_lists, all_proteins, plot_grid)

```

## Sensitivity 1 + 3: Removing NEFL outlier + filtering cases and their controls with follow-up \> 5 years

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r, fig.width=15, fig.height=6}

results_proteomic_ALS_occurrence$sensi_1_3$proteomic_sd_ALS_table_sensi_1_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Base model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3$proteomic_sd_ALS_base_figure_sensi_1_3 + 
  labs(title = "Filtered to follow-up>5years\n+ NEFL outlier excluded \n Base model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
  
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Adjusted model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3$proteomic_sd_ALS_adjusted_figure_sensi_1_3 + 
  labs(title = "Filtered to follow-up>5years\n+ NEFL outlier excluded \n Adjusted model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

Conditional logistic regressions (quartiles), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r}
#| column: screen-inset-right

results_proteomic_ALS_occurrence$sensi_1_3$proteomic_quart_ALS_table_sensi_1_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

Generalized additive models modeling proteins as smooth terms (thin plate splines), base models adjusted on sex and age at baseline. Adjusted models further account for smoking and body mass index

```{r fig.width=12, fig.height=60}
#| column: screen-inset-right

plots_lists <- list(
  sensi_1_base = results_proteomic_ALS_occurrence$sensi_1$plot_base_gam_sensi_1$signif,
  sensi_1_adj  = results_proteomic_ALS_occurrence$sensi_1$plot_adjusted_gam_sensi_1$signif,
  sensi_1_3_base = results_proteomic_ALS_occurrence$sensi_1_3$plot_base_gam_sensi_1_3$signif,
  sensi_1_3_adj  = results_proteomic_ALS_occurrence$sensi_1_3$plot_adjusted_gam_sensi_1_3$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 4) 

rm(plots_lists, all_proteins, plot_grid)

```

## Sensitivity 1 + 3 + 4: Removing NEFL outlier + filtering cases and their controls with follow-up \> 5 years + filtering follow-up \< 50%

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r, fig.width=15, fig.height=6}

results_proteomic_ALS_occurrence$sensi_1_3_4$proteomic_sd_ALS_table_sensi_1_3_4 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Base model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3_4$proteomic_sd_ALS_base_figure_sensi_1_3_4 + 
  labs(title = "NEFL outlier excluded\n+Filtered to follow-up>5years\n+Filtered to follow-up<50%\nBase model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
  
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Adjusted model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3_4$proteomic_sd_ALS_adjusted_figure_sensi_1_3_4 + 
  labs(title = "NEFL outlier excluded\n+Filtered to follow-up>5years\n+Filtered to follow-up<50%\nAdjusted model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

Conditional logistic regressions (quartiles), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r}
#| column: screen-inset-right

results_proteomic_ALS_occurrence$sensi_1_3_4$proteomic_quart_ALS_table_sensi_1_3_4 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

Generalized additive models modeling proteins as smooth terms (thin plate splines), base models adjusted on sex and age at baseline. Adjusted models further account for smoking and body mass index

```{r fig.width=12, fig.height=60}
#| column: screen-inset-right

plots_lists <- list(
  sensi_1_base = results_proteomic_ALS_occurrence$sensi_1$plot_base_gam_sensi_1$signif,
  sensi_1_adj  = results_proteomic_ALS_occurrence$sensi_1$plot_adjusted_gam_sensi_1$signif,
  sensi_1_3_4_base = results_proteomic_ALS_occurrence$sensi_1_3_4$plot_base_gam_sensi_1_3_4$signif,
  sensi_1_3_4_adj  = results_proteomic_ALS_occurrence$sensi_1_3_4$plot_adjusted_gam_sensi_1_3_4$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 4) 

rm(plots_lists, all_proteins, plot_grid)

```

## Sensitivity 1 + 3 + 5: Removing NEFL outlier + filtering cases and their controls with follow-up \> 5 years + filtering follow-up \> 50%

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r, fig.width=15, fig.height=6}

results_proteomic_ALS_occurrence$sensi_1_3_5$proteomic_sd_ALS_table_sensi_1_3_5 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Base model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3_5$proteomic_sd_ALS_base_figure_sensi_1_3_5 + 
  labs(title = "NEFL outlier excluded\n+Filtered to follow-up>5years\n+Filtered to follow-up>50%\nBase model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
  
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Adjusted model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_3_5$proteomic_sd_ALS_adjusted_figure_sensi_1_3_5 + 
  labs(title = "NEFL outlier excluded\n+Filtered to follow-up>5years\n+Filtered to follow-up>50%\nAdjusted model") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
```

Conditional logistic regressions (quartiles), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r}
#| column: screen-inset-right

results_proteomic_ALS_occurrence$sensi_1_3_5$proteomic_quart_ALS_table_sensi_1_3_5 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

Generalized additive models modeling proteins as smooth terms (thin plate splines), base models adjusted on sex and age at baseline. Adjusted models further account for smoking and body mass index

```{r fig.width=12, fig.height=60}
#| column: screen-inset-right

plots_lists <- list(
  sensi_1_base = results_proteomic_ALS_occurrence$sensi_1$plot_base_gam_sensi_1$signif,
  sensi_1_adj  = results_proteomic_ALS_occurrence$sensi_1$plot_adjusted_gam_sensi_1$signif,
  sensi_1_3_5_base = results_proteomic_ALS_occurrence$sensi_1_3_5$plot_base_gam_sensi_1_3_5$signif,
  sensi_1_3_5_adj  = results_proteomic_ALS_occurrence$sensi_1_3_5$plot_adjusted_gam_sensi_1_3_5$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 4) 

rm(plots_lists, all_proteins, plot_grid)

```

## Sensitivity 6: stratifiyng the analysis in tertiles of follow-up duration

Data not shown because of merging sensi 1 and sensi 6

```{r fig.height=12, fig.width=15}
#| eval: false
#| include: false

### Logistic regressions
# Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_base_figure + labs(title = "Main analysis - full sample size") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_base_figure_sensi_6_T1 + labs(title = "Filtered to follow-up T1 (<11 years)") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_base_figure_sensi_6_T2 + labs(title = "Filtered to follow-up T2 (11-16 years)") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_base_figure_sensi_6_T3 + labs(title = "Filtered to follow-up T3 (>16 years)") + theme(legend.position = "bottom") 
  
results_proteomic_ALS_occurrence$main$proteomic_sd_ALS_adjusted_figure + labs(title = "Main analysis - full sample size") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_adjusted_figure_sensi_6_T1 + labs(title = "Filtered to follow-up T1 (<11 years)") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_adjusted_figure_sensi_6_T2 + labs(title = "Filtered to follow-up T2 (11-16 years)") + theme(legend.position = "none") +
  results_proteomic_ALS_occurrence$sensi_6$proteomic_sd_ALS_adjusted_figure_sensi_6_T3 + labs(title = "Filtered to follow-up T3 (>16 years)") + theme(legend.position = "bottom") 
```

## Sensitivity 1 + 6: Removing NEFL outlier + stratifiyng the analysis in tertiles of follow-up duration

```{r}
results_proteomic_ALS_occurrence$sensi_6$sensi_6_table_follow_up
```

```{r, fig.height=3, fig.width=4}
results_proteomic_ALS_occurrence$sensi_6$sensi_6_densityplot_follow_up
```

### Logistic regressions

Conditional logistic regressions (continuous standardized), matched on age at baseline and sex (further adjusted for smoking and body mass index)

```{r fig.width=15, fig.height=12}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + labs(title = "NEFL outlier removed \n Base model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_base_figure_sensi_1_6_T1 + labs(title = "Filtered to follow-up T1 (<11 years)\n+ NEFL outlier excluded \n Base model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_base_figure_sensi_1_6_T2 + labs(title = "Filtered to follow-up T2 (11-16 years)\n+ NEFL outlier excluded \n Base model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_base_figure_sensi_1_6_T3 + labs(title = "Filtered to follow-up T3 (>16 years)\n+ NEFL outlier excluded \n Base model") + theme(legend.position = "bottom")  
  
  
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + labs(title = "NEFL outlier removed \n Adjusted model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_adjusted_figure_sensi_1_6_T1 + labs(title = "Filtered to follow-up T1 (<11 years)\n+ NEFL outlier excluded \n Adjusted model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_adjusted_figure_sensi_1_6_T2 + labs(title = "Filtered to follow-up T2 (11-16 years)\n+ NEFL outlier excluded \n Adjusted model") + theme(legend.position = "none") +
results_proteomic_ALS_occurrence$sensi_1_6$proteomic_sd_ALS_adjusted_figure_sensi_1_6_T3 + labs(title = "Filtered to follow-up T3 (>16 years)\n+ NEFL outlier excluded \n Adjusted model") + theme(legend.position = "bottom")  
```

NEFL is not significantly associated with ALS occurence when the follow-up duration is \>16 years. We don't see many associations in the follow-up group T1: I suspect that it is because of high heterogeneity in follow-up duration among this sub-group (0-11 years). --\> then I trust more the sensi 2 results where we just remove the cases (and their controls) with follow-up \<5 years.

## Sensitivity 1 + 7: Removing NEFL outlier + stratifiyng the analysis on sex

### Continuous coding (standardized)

```{r}
results_proteomic_ALS_occurrence$sensi_1_7$proteomic_sd_ALS_table_sensi_1_7
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_base_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Base model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_7$proteomic_sd_ALS_base_figure_sensi_1_7  + 
  labs(title = "NEFL outlier excluded\n+Filtered on sex") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)
  
results_proteomic_ALS_occurrence$sensi_1$proteomic_sd_ALS_adjusted_figure_sensi_1 + 
  labs(title = "NEFL outlier removed \n Adjusted model") + 
  theme(legend.position = "none") + 
  results_proteomic_ALS_occurrence$sensi_1_7$proteomic_sd_ALS_adjusted_figure_sensi_1_7 + 
  labs(title = "NEFL outlier excluded\n+Filtered on sex") + 
  theme(legend.position = "bottom") +
  plot_layout(ncol = 2)

```

### Quartiles

```{r}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$sensi_1_7$proteomic_quart_ALS_table_sensi_1_7
```

### GAMs

```{r fig.width=9, fig.height=70}
#| column: screen-inset-right

plots_lists <- list(
  sensi_1_adj  = results_proteomic_ALS_occurrence$sensi_1$plot_adjusted_gam_sensi_1$signif,
  sensi_1_7_female = results_proteomic_ALS_occurrence$sensi_1_7$plot_base_gam_sensi_1_7_female$signif,
  sensi_1_7_male  = results_proteomic_ALS_occurrence$sensi_1_7$plot_base_gam_sensi_1_7_male$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 3) 

rm(plots_lists, all_proteins, plot_grid)


```

## Additionnal analysis 1

```{r, fig.height=3, fig.width=3}
shapiro.test(results_proteomic_ALS_occurrence$additional_analysis_1$OR_distribution$OR_log)

results_proteomic_ALS_occurrence$additional_analysis_1$boxplot_OR
results_proteomic_ALS_occurrence$additional_analysis_1$densityplot_OR
```

## Additionnal analysis 2

Testing the distribution of the p-values depending on OR\>=1 to the distribution of the p-values depending on OR\<1 using a Fasano-Franceschini test

The Fasano–Franceschini test is a multivariate extension of the Kolmogorov–Smirnov test. While the classical KS test only compares one-dimensional distributions, the Fasano–Franceschini test can compare two-dimensional or multidimensional distributions.

It works by evaluating, for every point of one dataset, how the cumulative distribution differs from the other dataset in the four quadrants around that point. The test statistic D represents the largest difference between the empirical multivariate distributions of the two datasets. A large value of D indicates that the two datasets occupy different regions of the multidimensional space (e.g., different shapes, locations, or dispersions).

The null hypothesis (H₀) is: H₀: The two datasets come from the same multivariate distribution. A small p-value indicates a significant deviation between the two distributions.

### all proteins

```{r, fig.height=3, fig.width=9}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_all_figure
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_all_results

results_proteomic_ALS_occurrence$additional_analysis_2$jackknife_all_results |> 
  select(d_stat_percent) |> 
  mutate(d_stat_percent = round(d_stat_percent, 1)) |>
  tbl_summary()


```

Because the p-value is 0.00587, which is well below the standard 0.05 threshold, we reject the null hypothesis that the two sets of results (OR ≥ 1 vs. OR \< 1) come from the same distribution.

This means that the two groups of proteins, those associated with increased risk and those associated with decreased risk, show significantly different joint distributions of odds ratios and p-values. Therefore, the asymmetric patterns we observe are unlikely to be due to random fluctuation caused by multiple testing. The results support the idea that the proteins associated with increased disease risk and those associated with decreased disease risk represent genuinely different biological signals, rather than artifacts of high-dimensional multiple testing.

### Immune response

```{r, fig.height=3, fig.width=9}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_immune_figure
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_immune_results

results_proteomic_ALS_occurrence$additional_analysis_2$jackknife_immune_results |> 
  select(d_stat_percent) |> 
  mutate(d_stat_percent = round(d_stat_percent, 1)) |>
  tbl_summary()
```

### Metbolism

```{r, fig.height=3, fig.width=9}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_metabolism_figure
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_metabolism_results

results_proteomic_ALS_occurrence$additional_analysis_2$jackknife_metabolism_results |> 
  select(d_stat_percent) |> 
  mutate(d_stat_percent = round(d_stat_percent, 1)) |>
  tbl_summary()

```

### Neuro explo

```{r, fig.height=3, fig.width=9}
#| column: screen-inset-right
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_neuro_figure
results_proteomic_ALS_occurrence$additional_analysis_2$additional_analysis_2_neuro_results

results_proteomic_ALS_occurrence$additional_analysis_2$jackknife_neuro_results |> 
  select(d_stat_percent) |> 
  mutate(d_stat_percent = round(d_stat_percent, 1)) |>
  tbl_summary()
```

## Additional analysis 3 - superlearner

### Test 1 a : method AUC, many algorithms, no pre screening, default hyperparameters

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_a)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_a) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1a)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

Leeklasso is doing the best job --\> means that the relations are probably linear and for only a few proteins (because this algorithm has a internal prescreening)

Trees methodologies are not doing a great job but there are supposed to handle correctly high dimension: maybe not a lot of interactions between the proteins? But should investigate with tunned parameters

Algorithms such as SL.bayesglm, SL.stepAIC, SL.ksvm, SL.gam, SL.polymars and SL.earth should do better wit pre screening

### Defining the best hyperparameter combinaison for each algorithm

#### SL.glmnet

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_glmnet)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_glmnet) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.glmnet)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.ranger

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_ranger)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_ranger) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.ranger)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.xgboost

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_xgboost)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_xgboost) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.xgboost)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.ksvm

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_ksvm)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_ksvm) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.ksvm)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.earth

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_earth)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_earth) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.earth)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

### Test 1 b : method AUC, many algorithms, pre screening, optimized hyperparameters

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_b_opt_20)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_b_opt_20) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1b top 20)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_b_opt_10)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_b_opt_10) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1b top 10)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

### Test 1 c : method AUC, selected algorithms, pre screening, optimized hyperparameters

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_20)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_20) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1c top 20)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_10)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_10) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1c top 10)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### Weights of the algorithms

```{r}
# Calcul des poids moyens en pourcentage
all_weights <- lapply(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_10$AllSL, function(x) x$coef)
weights_matrix <- do.call(rbind, all_weights)
colnames(weights_matrix) <- results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_10$libraryNames

df_weights <- data.frame(
  Algorithms = colnames(weights_matrix),
  Mean_weights = colMeans(weights_matrix) * 100) |> 
  dplyr::arrange(desc(Mean_weights))

df_weights |>
  flextable() |>
  set_header_labels(Mean_weights = "Mean weights (%)") |>
  colformat_double(j = "Mean_weights", digits = 1) |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 10, part = "all") |>
  autofit()
```

#### Selected proteins by leeklasso

```{r}
# Extraction des protéines et coefficients pour Leekasso
leek_results_list <- lapply(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_10$AllSL, function(fold_fit) {
  obj <- fold_fit$fitLibrary$SL.leekasso_All$object
  if(is.null(obj)) return(NULL)
  
  coefs <- coef(obj)
  coefs <- coefs[names(coefs) != "(Intercept)"]
  
  data.frame(
    Protein = names(coefs),
    Log_Odds = as.numeric(coefs)
  )
})

df_leek_final <- do.call(rbind, leek_results_list) |>
  dplyr::group_by(Protein) |>
  dplyr::summarise(
    Consistency = n() / 10 * 100,
    Mean_OR = exp(mean(Log_Odds)) ) |>
  dplyr::arrange(desc(Consistency))


df_leek_final |>
  flextable() |>
  set_header_labels(
    Protein = "Selected proteins",
    Consistency = "Consistency across folds (%)", 
    Mean_OR = "Mean ORs across folds") |>
  colformat_double(j = "Mean_OR", digits = 2) |>
  colformat_double(j = "Consistency", digits = 1) |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 10, part = "all") |>
  autofit()
```

#### Selected proteins by GLMNet

```{r}
# Extraction des protéines selectionnées par GLMNet dans chaque pli
glm_results_list <- lapply(results_proteomic_ALS_occurrence_SL$test_1$sl_fit_CV_1_c_opt_10$AllSL, function(fold_fit) {
  fit_glmnet <- fold_fit$fitLibrary$SL.glmnet.opt_All$object
  coefs <- as.matrix(coef(fit_glmnet, s = "lambda.min"))
  
  data.frame(
    Protein = rownames(coefs),
    Log_Odds = as.numeric(coefs)
  ) |>
    dplyr::filter(Protein != "(Intercept)")
})

df_glm_final <- do.call(rbind, glm_results_list) |>
  dplyr::group_by(Protein) |>
  dplyr::summarise(
    Consistency = sum(Log_Odds != 0) / 10 * 100,
    Mean_OR = exp(mean(Log_Odds[Log_Odds != 0]))) |>
  dplyr::filter(Consistency > 0) |>
  dplyr::arrange(desc(Consistency))

df_glm_final |>
  flextable() |>
  set_header_labels(
    Protein = "Proteine",
    Consistency = "Consistency across folds (%)", 
    Mean_OR = "Mean ORs across folds") |>
  colformat_double(j = "Mean_OR", digits = 2) |>
  colformat_double(j = "Consistency", digits = 1) |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 10, part = "all") |>
  autofit()
```

APEX1: DNA repair enzyme (Apurinic/Apyrimidinic Endonuclease 1)

NPTXR: Neuronal Pentraxin Receptor (involved in synaptic remodeling and glutamatergic signaling)

NfL: Marker of neuroaxonal damage

ARG1 (Arginase-1): Marker of alternative (M2) macrophage/microglial polarization

CD33: Myeloid cell surface antigen (Microglial marker)

PIK3AP1 (Phosphoinositide 3-kinase adapter protein 1): link between B-cell receptors and the PI3K signaling pathway

PSIP1 (PC4 and SFRS1-interacting protein 1): transcriptional coactivator that promotes cell survival (protecting cells against oxidative stress)

GLRX (Glutaredoxin-1): small redox enzyme that uses glutathione to reduce oxidized proteins (disulfides) (maintaining redox homeostasis)

IL-32 (Interleukin-32): Pro-inflammatory cytokine

NPM1 (Nucleophosmin): multifunctional "chaperone" protein found mainly in the nucleolus. it is involved in ribosome biogenesis, DNA repair, and protein folding.

GHRL (Ghrelin): metabolic peptide produced mainly in the stomach that regulates appetite and energy balance.

CD302: C-type lectin receptor involved in cell adhesion and migration, primarily expressed on myeloid cells like macrophages and dendritic cells.

PADI2: Protein-arginine deiminase type-2

### Test 2: + time as a predictor

### Defining the best hyperparameter combinaison for each algorithm

#### SL.glmnet

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_glmnet)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_glmnet) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.glmnet)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.ranger

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_ranger)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_ranger) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.ranger)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.xgboost

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_xgboost)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_xgboost) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.xgboost)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.ksvm

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_ksvm)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_ksvm) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.ksvm)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

#### SL.earth

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_earth)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_earth) + 
  theme_minimal() + 
  labs(title = "Model Performance (Specific to SL.earth)", 
       subtitle = "ALS detection",
       x = "Hyperparameter combinaisons", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

### Test 2 b : method AUC, many algorithms, pre screening, optimized hyperparameters

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_b_opt_20)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_b_opt_20) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1b top 20)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_b_opt_10)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_b_opt_10) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1b top 10)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

### Test 2 c : method AUC, selected algorithms, pre screening, optimized hyperparameters

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_20)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_20) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1c top 20)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

```{r}
summary(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_10)$Table |>
  as.data.frame() |>
  flextable() |> 
  colformat_double(digits = 2) |>
  font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left") |>
  autofit()

plot(results_proteomic_ALS_occurrence_SL$test_2$sl_fit_CV_2_c_opt_10) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1c top 10)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

# Proteomic and years to ALS

## Cox models among cases 

Event = ALS, time to event = from baseline to ALS diagnostic (cases)

### continuous standardized

```{r, fig.width=15, fig.height=5}
results_proteomic_ALS_occurrence_y_to_als$main_results_cox$proteomic_sd_ALS_table_cox |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right

wrap_plots(results_proteomic_ALS_occurrence_y_to_als$main_results_cox$proteomic_sd_ALS_base_figure_cox, 
           results_proteomic_ALS_occurrence_y_to_als$main_results_cox$proteomic_sd_ALS_adjusted_figure_cox, 
           ncol = 2)
```

### quartiles

```{r}
#| column: screen-inset-right
results_proteomic_ALS_occurrence_y_to_als$main_results_cox$proteomic_quart_ALS_table_cox |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=20, fig.width=9}
#| eval: false
#| include: false
plots_lists <- list(
  base  = results_proteomic_ALS_occurrence_y_to_als$main_results_cox$plot_base_cox_gam_danish$signif, 
  adj  = results_proteomic_ALS_occurrence_y_to_als$main_results_cox$plot_adjusted_cox_gam_danish$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(plots_lists, all_proteins, plot_grid)

```

## Fixed-horizon logistic analyses

### Proteome wide associations (logistic regressions)

#### Continuous coding (standardized)

```{r}
#| column: screen-inset-right
results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$proteome_wide$proteomic_sd_ALS_table_y_10_20
```

```{r}
results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$proteome_wide$proteomic_sd_ALS_adjusted_figure_y_10ans
```

```{r}
results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$proteome_wide$proteomic_sd_ALS_adjusted_figure_y_20ans
```

#### Quartiles

```{r}
#| column: screen-inset-right
results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$proteome_wide$proteomic_quart_ALS_table_y_10_20
```

### Machine learning (Xboost)

#### 10 years horizon

Number of optimal trees

```{r}
cat("---> Optimal number of trees (10 years horizon) :", results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$best_nrounds_10, "\n")
```

Maximum AUC (Cross-validated)

```{r}
cat("---> Maximum AUC (cross validated, 10 years horizon) :", max(results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$cv_10ans$evaluation_log$test_auc_mean), "\n")
```

Importance top 20 variables (10 years horizon)

```{r}
cat("\nTop 20 predictors - 10 years horizon :\n")
head(results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$importance_10, 20) |>
    flextable() |>
  merge_h(part = "header") |>
  theme_vanilla() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all")
```

```{r}
xgb.plot.importance(
  importance_matrix = results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$importance_10[1:20, ], 
  main = "Importance of the top 20 predictors (10 years horizon)")
```

#### 20 years horizon

Number of optimal trees

```{r}
cat("---> Optimal number of trees (20 years horizon) :", results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$best_nrounds_20, "\n")
```

Maximum AUC (Cross-validated)

```{r}
cat("---> Maximum AUC (cross validated, 20 years horizon) :", max(results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$cv_20ans$evaluation_log$test_auc_mean), "\n")
```

Importance top 20 variables (20 year horizon)

```{r}
cat("\nTop 20 predictors - 20 years horizon :\n")
head(results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$importance_20, 20) |>
    flextable() |>
  merge_h(part = "header") |>
  theme_vanilla() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all")
```

```{r}
xgb.plot.importance(
  importance_matrix = results_proteomic_ALS_occurrence_y_to_als$fixed_horizon_logistic_regression_analysis$machine_learning_xgboost$importance_20[1:20, ], 
  main = "Importance of the top 20 predictors (20 years horizon)")
```

## Logistic models including time
### Proteome wide associations 
```{r}
results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$proteome_wide$proteomic_interaction_table
```

### Machine learning (XGboost)
```{r}
summary(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_fit)
plot(results_proteomic_ALS_occurrence_y_to_als$logistic_models_including_time$machine_learning_xgboost$cv_sl_fit)
```


# Proteomic and ALS survival

## Effects of subject characteristics on ALS survival

```{r}
results_proteomic_ALS_survival$main$covar |>
  as_flex_table() |>
  flextable::font(fontname = "Calibri", part = "all") |> 
  fontsize(size = 10, part = "all") |>
  padding(padding.top = 0, padding.bottom = 0, part = "all") |>
  set_table_properties(align = "left")
```

## Main analyses

### continuous standardized

```{r, fig.width=15, fig.height=5}
results_proteomic_ALS_survival$main$proteomic_sd_ALS_table |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right

wrap_plots(results_proteomic_ALS_survival$main$proteomic_sd_ALS_base_figure, 
           results_proteomic_ALS_survival$main$proteomic_sd_ALS_adjusted_figure, 
           ncol = 2)
```

### quartiles

```{r}
#| column: screen-inset-right
results_proteomic_ALS_survival$main$proteomic_quart_ALS_table  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=55, fig.width=9}
plots_lists <- list(
  main_base  = results_proteomic_ALS_survival$main$plot_base_cox_gam_danish$signif, 
  main_adj  = results_proteomic_ALS_survival$main$plot_adjusted_cox_gam_danish$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(plots_lists, all_proteins, plot_grid)

```

## Sensi 1 + 3 : Removing NEFL outlier + filtering cases with follow-up \> 5 years

### continuous standardized

```{r, fig.width=15, fig.height=5}
results_proteomic_ALS_survival$sensi_1_3$proteomic_sd_ALS_table_sensi_1_3 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
wrap_plots(results_proteomic_ALS_survival$sensi_1_3$proteomic_sd_ALS_base_figure_sensi_1_3, 
           results_proteomic_ALS_survival$sensi_1_3$proteomic_sd_ALS_adjusted_figure_sensi_1_3, 
           ncol = 2)

```

### quartiles

```{r}
#| column: screen-inset-right
results_proteomic_ALS_survival$sensi_1_3$proteomic_quart_ALS_table_sensi_1_3  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=60, fig.width=9}
plots_lists <- list(
  main = results_proteomic_ALS_survival$main$plot_adjusted_cox_gam_danish$signif,
  sensi_1_3 = results_proteomic_ALS_survival$sensi_1_3$plot_adjusted_cox_gam_danish_sensi_1_3$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(plots_lists, all_proteins, plot_grid)
```

## Sensi 1 + 3 + 4 : Removing NEFL outlier + filtering cases with follow-up \> 5 years + filtering follow-up \< 50%

### continuous standardized

```{r, fig.width=9, fig.height=3}
#| column: screen-inset-right

results_proteomic_ALS_survival$sensi_1_3_4$proteomic_sd_ALS_table_sensi_1_3_4 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
wrap_plots(results_proteomic_ALS_survival$sensi_1_3_4$proteomic_sd_ALS_base_figure_sensi_1_3_4,
           results_proteomic_ALS_survival$sensi_1_3_4$proteomic_sd_ALS_adjusted_figure_sensi_1_3_4, 
           ncol = 2)

```

### quartiles

```{r}
#| column: screen-inset-right
results_proteomic_ALS_survival$sensi_1_3_4$proteomic_quart_ALS_table_sensi_1_3_4  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=100, fig.width=9}
plots_lists <- list(
  main = results_proteomic_ALS_survival$main$plot_adjusted_cox_gam_danish$signif,
  sensi_1_3 = results_proteomic_ALS_survival$sensi_1_3_4$plot_adjusted_cox_gam_danish_sensi_1_3_4$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(plots_lists, all_proteins, plot_grid)
```

## Sensi 1 + 3 + 5 : Removing NEFL outlier + filtering cases with follow-up \> 5 years + filtering follow-up \> 50%

### continuous standardized

```{r, fig.width=9, fig.height=3}
#| column: screen-inset-right

results_proteomic_ALS_survival$sensi_1_3_5$proteomic_sd_ALS_table_sensi_1_3_5 |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
wrap_plots(results_proteomic_ALS_survival$sensi_1_3_5$proteomic_sd_ALS_base_figure_sensi_1_3_5,
           results_proteomic_ALS_survival$sensi_1_3_5$proteomic_sd_ALS_adjusted_figure_sensi_1_3_5, 
           ncol = 2)

```

### quartiles

```{r}
#| column: screen-inset-right
results_proteomic_ALS_survival$sensi_1_3_5$proteomic_quart_ALS_table_sensi_1_3_5  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=80, fig.width=9}

plots_lists <- list(
  main = results_proteomic_ALS_survival$main$plot_adjusted_cox_gam_danish$signif,
  sensi_1_3_5 = results_proteomic_ALS_survival$sensi_1_3_5$plot_adjusted_cox_gam_danish_sensi_1_3_5$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(plots_lists, all_proteins, plot_grid)

```

## Sensi 1 + 7 females : Removing NEFL outlier + filtering female cases

### continuous standardized

```{r, fig.width=9, fig.height=3}
#| column: screen-inset-right

results_proteomic_ALS_survival$sensi_1_7_female$proteomic_sd_ALS_table_sensi_1_7_female |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
wrap_plots(results_proteomic_ALS_survival$sensi_1_7_female$proteomic_sd_ALS_base_figure_sensi_1_7_female,
           results_proteomic_ALS_survival$sensi_1_7_female$proteomic_sd_ALS_adjusted_figure_sensi_1_7_female, 
           ncol = 2)

```

### quartiles

```{r}
#| column: screen-inset-right
results_proteomic_ALS_survival$sensi_1_7_female$proteomic_quart_ALS_table_sensi_1_7_female  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=80, fig.width=9}
plots_lists <- list(
  main = results_proteomic_ALS_survival$main$plot_adjusted_cox_gam_danish$signif,
  sensi_1_7_female = results_proteomic_ALS_survival$sensi_1_7_female$plot_adjusted_cox_gam_danish_sensi_1_7_female$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(plots_lists, all_proteins, plot_grid)
```

## Sensi 1 + 7 males : Removing NEFL outlier + male cases

### continuous standardized

```{r, fig.width=9, fig.height=3}
#| column: screen-inset-right

results_proteomic_ALS_survival$sensi_1_7_male$proteomic_sd_ALS_table_sensi_1_7_male |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

```{r, fig.width=15, fig.height=6}
#| column: screen-inset-right
wrap_plots(results_proteomic_ALS_survival$sensi_1_7_male$proteomic_sd_ALS_base_figure_sensi_1_7_male,
           results_proteomic_ALS_survival$sensi_1_7_male$proteomic_sd_ALS_adjusted_figure_sensi_1_7_male, 
           ncol = 2)

```

### quartiles

```{r}
#| column: screen-inset-right

results_proteomic_ALS_survival$sensi_1_7_male$proteomic_quart_ALS_table_sensi_1_7_male  |>
  set_table_properties(width = 1, layout = "autofit", align = "left")
```

### GAMs

```{r fig.height=80, fig.width=9}
plots_lists <- list(
  main = results_proteomic_ALS_survival$main$plot_adjusted_cox_gam_danish$signif,
  sensi_1_7_male = results_proteomic_ALS_survival$sensi_1_7_male$plot_adjusted_cox_gam_danish_sensi_1_7_male$signif)

all_proteins <- sort(unique(unlist(lapply(plots_lists, names))))

plot_grid <- lapply(all_proteins, function(prot) {
  lapply(plots_lists, function(lst) {
    if (prot %in% names(lst)) {
      lst[[prot]]
    } else {
      wrap_elements(grid::nullGrob())
    }
  })
})

wrap_plots(
  unlist(plot_grid, recursive = FALSE),
  ncol = 2) 

rm(plots_lists, all_proteins, plot_grid)

```
