---
title: "Fatty acids and ALS"
format: html
editor: visual
toc: true
toc_depth: 4
theme:
  light: flatly
  dark: darkly
toc-location: left
---

```{r}
#| message: false
#| warning: false
#| include: false
source("~/Documents/POP_ALS_2025_02_03/1_codes/lipids.R")
```

# Data description
## Comparison finnish / danish
### Covariates
```{r}
#| echo: false
#| message: false
#| warning: false
bdd %>%
  select("study", "sex", "marital_status_2cat", "smoking_2cat", "alcohol", "education", 
         "bmi", "cholesterol", "baseline_age", "diagnosis_age") %>%
  tbl_summary(by = "study") %>%
  bold_labels() %>%
  add_p()
```

### POPs
```{r}
#| echo: false
#| message: false
#| warning: false
bdd |> 
  select(sample, als, study, all_of(POPs), all_of(POPs_group)) |>
  pivot_longer(cols = c(-sample, -als, -study), names_to = "POPs", values_to = "Values") |>
  ggplot() +
  aes(x = Values, y = POPs, fill = study) +
  geom_boxplot() +
  scale_fill_hue(direction = 1) +
  labs(
    x = "Serum concentration (/)",
    y = "POPs",
    fill = "Cohorts") +
  theme_lucid()

bdd |>
  filter(als == 0) |>
  select(study, all_of(POPs), all_of(POPs_group)) |>
  tbl_summary(by = "study") |>
  bold_labels() |>
  add_n()
```

### Fatty acids
```{r}
#| echo: false
#| message: false
#| warning: false
bdd |> 
  select(sample, als, study, all_of(fattyacids), all_of(explanatory)) |>
  pivot_longer(cols = c(-sample, -als, -study), names_to = "fattyacids", values_to = "Values") |>
  ggplot() +
  aes(x = Values, y = fattyacids, fill = study) +
  geom_boxplot() +
  scale_fill_hue(direction = 1) +
  labs(
    x = "Serum concentration (/)",
    y = "Fatty acids",
    fill = "Cohorts") +
  theme_lucid()

bdd |>
  filter(als == 0) |>
  select(study, all_of(fattyacids), all_of(explanatory)) |>
  tbl_summary(by = "study") |>
  bold_labels() |>
  add_n()
```

# Descriptif Danish cohort
## Fatty acids
### Density plots 
```{r, fig.height=10, fig.width=10}
#| echo: false
#| message: false
#| warning: false
#| column: screen-inset-right

bdd_danish |> 
  select(sample, all_of(fattyacids)) |>
  pivot_longer(cols = -sample, names_to = "fattyacids", values_to = "values") |>
  ggplot() +
  aes(x = values) +
  geom_density(fill = "#112446") +
  theme_minimal() +
  facet_wrap(~fattyacids, scales = "free", ncol = 4)
```

### Boxplots
```{r, fig.height=10, fig.width=10}
#| echo: false
#| message: false
#| warning: false
#| column: screen-inset-right

bdd_danish |> 
  select(sample, all_of(fattyacids)) |>
  pivot_longer(cols = -sample, names_to = "fattyacids", values_to = "values") |>
  ggplot() +
  aes(x = "", y = values) +
  geom_boxplot(shape = "circle", fill = "#112446") +
  coord_flip() +
  theme_bw() +
  theme(axis.title = element_blank()) +
  facet_wrap(~fattyacids, scales = "free", ncol = 3L, nrow = 14L)

bdd_danish |> select(sample, all_of(fattyacids)) |>
  pivot_longer(cols = -sample, names_to = "fattyacids", values_to = "values") |>
  ggplot() +
  aes(x = values, y = fattyacids) +
  geom_boxplot(fill = "#112446") +
  theme_minimal()
```

### Distribution tables 
```{r, fig.height=10, fig.width=10}
#| echo: false
#| message: false
#| warning: false
#| column: screen-inset-right

descrip_num(data = bdd_danish, vars = fattyacids)
```

### Heatmap of correlation
```{r, fig.height=10, fig.width=10}
#| echo: false
#| message: false
#| warning: false
#| column: screen-inset-right

heatmap_fattyacids <- bdd_danish |> select(all_of(fattyacids))
heatmap_fattyacids <- cor(heatmap_fattyacids, 
                    use = "pairwise.complete.obs", 
                    method = "pearson")

corrplot(heatmap_fattyacids, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))

heatmap_cor_pairwise(fattyacids, POPs, decimal = 1, data = bdd_danish)
```

### Calcium (indicator of quality)
Why are there two cholesterol variables?
Why don't we have a calcium variable for the danish data 
```{r}
#| echo: false
#| message: false
#| warning: false
boxplot(data = bdd_danish, vars = c("fS_Trigly", "fS_Kol", "cholesterol"))
descrip_num(data = bdd_danish, vars = c("fS_Trigly", "fS_Kol", "cholesterol"))
```

# Descriptif Finnish cohort
## Fatty acids
### Density plots 
```{r, fig.height=10, fig.width=10}
#| echo: false
#| message: false
#| warning: false
#| column: screen-inset-right

bdd_finnish |> 
  select(sample, all_of(fattyacids)) |>
  pivot_longer(cols = -sample, names_to = "fattyacids", values_to = "values") |>
  ggplot() +
  aes(x = values) +
  geom_density(fill = "#112446") +
  theme_minimal() +
  facet_wrap(~fattyacids, scales = "free", ncol = 4)
```

### Boxplots
```{r, fig.height=10, fig.width=10}
#| echo: false
#| message: false
#| warning: false
#| column: screen-inset-right

bdd_finnish |> 
  select(sample, all_of(fattyacids)) |>
  pivot_longer(cols = -sample, names_to = "fattyacids", values_to = "values") |>
  ggplot() +
  aes(x = "", y = values) +
  geom_boxplot(shape = "circle", fill = "#112446") +
  coord_flip() +
  theme_bw() +
  theme(axis.title = element_blank()) +
  facet_wrap(~fattyacids, scales = "free", ncol = 3L, nrow = 14L)

bdd_finnish |> select(sample, all_of(fattyacids)) |>
  pivot_longer(cols = -sample, names_to = "fattyacids", values_to = "values") |>
  ggplot() +
  aes(x = values, y = fattyacids) +
  geom_boxplot(fill = "#112446") +
  theme_minimal()
```

### Distribution tables 
```{r, fig.height=10, fig.width=10}
#| echo: false
#| message: false
#| warning: false
#| column: screen-inset-right

descrip_num(data = bdd_finnish, vars = fattyacids)
```

### Heatmap of correlation
```{r, fig.height=10, fig.width=10}
#| echo: false
#| message: false
#| warning: false
#| column: screen-inset-right

heatmap_fattyacids <- bdd_finnish |> select(all_of(fattyacids))
heatmap_fattyacids <- cor(heatmap_fattyacids, 
                    use = "pairwise.complete.obs", 
                    method = "pearson")

corrplot(heatmap_fattyacids, 
         method = 'color', 
         type = "lower", 
         tl.col = 'black', 
         tl.srt = 45, 
         addCoef.col = "black",
         number.cex = 0.8,
         number.digits = 1,
         tl.cex = 1,
         col = rev(COL2(diverging = "RdYlBu")))

heatmap_cor_pairwise(fattyacids, POPs_finnish, decimal = 1, data = bdd_finnish)
```

### Calcium (indicator of quality)
Why do we have 3 cholesterol variables in the finnish data ?
Regarding the samples with Ca2+<2: run a sensitivity analysis without them
```{r}
#| echo: false
#| message: false
#| warning: false
boxplot(data = bdd_finnish, vars = c("S_Ca", "fS_Trigly", "fS_Kol", "chol", "cholesterol"))
descrip_num(data = bdd_finnish, vars = c("S_Ca", "fS_Trigly", "fS_Kol", "chol", "cholesterol"))
```

