---
title: "Super learner"
format: html
editor: visual
toc: true
toc_depth: 4
theme:
  light: flatly
  dark: darkly
toc-location: left
execute:
  echo: false
  message: false
  warning: false
---

```{r}
source("~/Documents/POP_ALS_2025_02_03/1_codes/0_functions.R")
#results_proteomic_ALS_occurrence_SL <- readRDS("~/Documents/POP_ALS_2025_02_03/2_output/results_proteomic_ALS_occurrence_SL.rds")

results_proteomic_ALS_occurrence_SL <- readRDS("~/Desktop/results_proteomic_ALS_occurrence_SL.rds")
```

# Test 1 a : method AUC, many algorithms, no pre screening, default hyperparameters

```{r}
summary(results_proteomic_ALS_occurrence_SL$sl_fit_CV_1_a)
plot(results_proteomic_ALS_occurrence_SL$sl_fit_CV_1_a) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1a)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

Leeklasso is doing the best job --\> means that the relations are probably linear and for only a few proteins (because this algorithm has a internal prescreening)

Trees methodologies are not doing a great job but there are supposed to handle correctly high dimension: maybe not a lot of interactions between the proteins? But should investigate with tunned parameters

Algorithms such as SL.bayesglm, SL.stepAIC, SL.ksvm, SL.gam, SL.polymars and SL.earth should do better wit pre screening

# Test 1 b : method AUC, many algorithms, pre screening when necessary, default hyperparameters

```{r}
summary(results_proteomic_ALS_occurrence_SL$sl_fit_CV_1_b)
plot(results_proteomic_ALS_occurrence_SL$sl_fit_CV_1_b) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1b)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

Leeklasso results don't change because there was already a pre screening in step 1a

SL.gam goes from 0.518 to 0.577 between step 1 a and step 1b. the pre screening was necessary for the algorithm.

SL.earth goes from 0.547 to 0.584 between step 1 a and step 1b. the pre screening was necessary for the algorithm.

Tree methodologies (SL.ranger and SL.xgboost) need tunned hyperparameters

# Test 1 c : method AUC, many algorithms, pre screening when necessary, tunned hyperparameters

```{r}
summary(results_proteomic_ALS_occurrence_SL$sl_fit_CV_1_c)
plot(results_proteomic_ALS_occurrence_SL$sl_fit_CV_1_c) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1c)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

Finally super Learner (0.603) is higher than all other algorithms except leeklasso.

Lasso on all proteins is doing a better job than Elastic net with pre-screening top 20 proteins

SL.earth.tuned with degree = 2 is doing a better job indicating that there are some interactions

SL.ranger.tuned (0.56) and SL.xgb.tuned (0.53) are still not doing a really good job

# Test 1 d : method AUC, selected algorithms, pre screening when necessary, tunned hyperparameters

```{r}
summary(results_proteomic_ALS_occurrence_SL$sl_fit_CV_1_d)
plot(results_proteomic_ALS_occurrence_SL$sl_fit_CV_1_d) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1d)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

Super Learner is higher than in the previous steps but still lower than leeklasso

Leeklasso is doing the best job --\> probably jus a few porteins that are meaninful for ALS risk

SL.earth.tuned et SL.gam on 50 top proteins instead of 20 (step 1c) did not improve the results

Relaxed lasso is not doing a best job than a classic lasso

# Test 1 e : method AUC, selected algorithms, pre screening when necessary, tunned hyperparameters

## Main results

```{r}
summary(results_proteomic_ALS_occurrence_SL$sl_fit_CV_1_e)
plot(results_proteomic_ALS_occurrence_SL$sl_fit_CV_1_e) + 
  theme_minimal() + 
  labs(title = "Model Performance (Step 1e)", 
       subtitle = "ALS detection",
       x = "Algorithm", y = "Cross-Validated AUC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")
```

## Weights of the algorithms

```{r}
results_proteomic_ALS_occurrence_SL$weights_1_e[order(-results_proteomic_ALS_occurrence_SL$weights_1_e$Weight), ]
```

The diverse weight distribution confirms that no single model captured all the signal:

SL.leekasso (35.0%): The primary driver. Confirms a sparse, linear relationship for the most significant proteins.

SL.mean (20.8%): Suggests the signal is subtle and prevents the model from over-interpreting noise.

SL.glmnet (19.2%): Provides robust penalized linear modeling across all 276 proteins.

SL.earth.tuned (17.7%): Captures non-linear thresholds and potential protein interactions that linear models miss.

SL.gam (7.2%): Adds a small layer of flexibility through additive splines.

## Selected proteins by lasso

```{r}
results_proteomic_ALS_occurrence_SL$selected_proteins_lasso_1_e |>
  mutate(
    OR = exp(Coefficient)
  ) %>%
  arrange(desc(OR))
```

NfL: Marker of neuroaxonal damage

IL-32 (Interleukin-32): Pro-inflammatory cytokine

APEX1: DNA repair enzyme (Apurinic/Apyrimidinic Endonuclease 1)

CD33: Myeloid cell surface antigen (Microglial marker)

PADI2: Protein-arginine deiminase type-2

NPTXR: Neuronal Pentraxin Receptor (involved in synaptic remodeling and glutamatergic signaling)

ARG1 (Arginase-1): Marker of alternative (M2) macrophage/microglial polarization
